---
title: "Visualizations"
output:
  word_document: default
  pdf_document: default
  html_document: default
date: "2026-04-10"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,       # hides code
  message = FALSE,    # suppresses messages
  warning = FALSE     # suppresses warnings
)
```

# 1. Load and clean data 

```{r}

#load data 

load("data/CLSA_methylation.rds")

library(dplyr)
library(ggplot2)
library(tidyverse)
library(flextable)
library (haven)
library(table1)
library(stargazer)
library(htmltools)

library(table1)
library(Hmisc)
library(htmltools)
library(webshot2)
library(tidyverse)


#not going to include these: 
#"ED_HIGH_COM", "WEA_SVNGSVL_MCQ","temp_1yr_com" LGTNLTYY_01_COM "ENV_HMPRB_NOI_MCQ","pm25_5yrs_com",   "PM25DALYY_02_COM", "no2_5yrs_com", 
 # "Locrds_length_200m_com",   "Phwy_length_200m_com", 

#create subset 
subset_data <- CLSA_methylation[, c(
  "entity_id.x", 
  
 #demographics
  "AGE_NMBR_COM", "SEX_ASK_COM", "WGHTS_PROV_COM", "Race7",
  
  ##############
 #Socioeconomic variables
 ##############
 
  "INC_TOT_COM", 
  
 
  ##############
 #Aging outcome variables
 ##############
 
  # aging outcomes
  "GrimAge_EAA", "PhenoAge_EAA", "DunedinPACE", "DNAmIC",
  
 
 ##############
 #Environmental variables
 ##############
 
  #pollution 
   "PM25DALYY_01_COM",  "O3CHGYY_01_COM", "no2_1yr_com", "so2_1yr_com",
 
 #condensation 
 "ENV_HMPRB_CON_MCQ",
 
 #NDVI
 "GRLANYY_26_COM", 
  
  ##Tree canopy and %veg 
  "grtcc10_01_COM",  "grtcc10_04_COM", #tree canopy coverage 
  
  "GRUMPYY_01_COM",    "GRUMPYY_04_COM", #%vegetative greenness 
  
  
  # Lifestyle / behavioral
  "PA2_DSCR2_MCQ", "ALC_FREQ_COM", "ICQ_SMOKE_COM", #PASE scale: PASE Score - Final (in sharepoint dictionary)
  
  #cell types
  "CD4nv", "CD8nv",   "Mono", "Treg", "CD4mem", "CD8mem", "NK", "Bmem", "Bnv"
)]


#check structure of variables 
str(subset_data)





```


## Recode missing values to N/A

```{r}
# Recode NAs
subset_data$no2_1yr_com[subset_data$no2_1yr_com %in% c(-8888, -9999)] <- NA
subset_data$so2_1yr_com[subset_data$so2_1yr_com %in% c(-8888, -9999)] <- NA
subset_data$GRLANYY_26_COM[subset_data$GRLANYY_26_COM %in% c(-8888, -9999)] <- NA
subset_data$ENV_HMPRB_CON_MCQ[subset_data$ENV_HMPRB_CON_MCQ %in% c(-8888, -9999)] <- NA
subset_data$PM25DALYY_01_COM[subset_data$PM25DALYY_01_COM %in% c(-8888, -9999)] <- NA
subset_data$O3CHGYY_01_COM[subset_data$O3CHGYY_01_COM %in% c(-8888, -9999)] <- NA
subset_data$grtcc10_01_COM[subset_data$grtcc10_01_COM %in% c(-8888, -9999)] <- NA
subset_data$grtcc10_04_COM[subset_data$grtcc10_04_COM %in% c(-8888, -9999)] <- NA
subset_data$GRUMPYY_01_COM[subset_data$GRUMPYY_01_COM %in% c(-8888, -9999)] <- NA
subset_data$GRUMPYY_04_COM[subset_data$GRUMPYY_04_COM %in% c(-8888, -9999)] <- NA

subset_data$PA2_DSCR2_MCQ[subset_data$PA2_DSCR2_MCQ %in% c(-888)] <- NA
subset_data$ICQ_SMOKE_COM[subset_data$ICQ_SMOKE_COM %in% c(-8)] <- NA
subset_data$ALC_FREQ_COM[subset_data$ALC_FREQ_COM %in% c(77, 98, 99)] <- NA


colSums(is.na(subset_data))#check number of NAs

```




## Factor and label variables

```{r}
library(gtsummary)
library(dplyr)
library(flextable)

# ============= Recode, Factor, and Label Variables =============

# Sex
subset_data <- subset_data %>%
  mutate(SEX_ASK_COM = factor(SEX_ASK_COM,
                              levels = c("M", "F"),
                              labels = c("Male", "Female")))

# Province
subset_data <- subset_data %>%
  mutate(WGHTS_PROV_COM = factor(WGHTS_PROV_COM,
                                 levels = 1:10,
                                 labels = c("Alberta", "British Columbia", "Manitoba",
                                            "New Brunswick", "Newfoundland and Labrador",
                                            "Nova Scotia", "Ontario", "Prince Edward Island",
                                            "Quebec", "Saskatchewan")))

# Race/Ethnicity
subset_data <- subset_data %>%
  mutate(Race7 = factor(Race7,
                        levels = c("White", "South Asian", "Black", "Asian",
                                   "Arab", "Latin American", "Other")))

# Household Income — reverse code so 1=highest, 5=lowest
# Original: 1=<20K, 2=20-50K, 3=50-100K, 4=100-150K, 5=150K+, 8/9=Don't know/Refused
cat("=== Income BEFORE reverse coding ===\n")
table(subset_data$INC_TOT_COM, useNA = "ifany")

subset_data <- subset_data %>%
  mutate(
    INC_TOT_COM = factor(
      case_when(
        INC_TOT_COM == 1 ~ 5,
        INC_TOT_COM == 2 ~ 4,
        INC_TOT_COM == 3 ~ 3,
        INC_TOT_COM == 4 ~ 2,
        INC_TOT_COM == 5 ~ 1,
        INC_TOT_COM %in% c(8, 9) ~ 6,
        TRUE ~ NA_real_
      ),
      levels = c(1, 2, 3, 4, 5, 6),
      labels = c("150K+", "100K–150K", "50K–100K", "20K–50K", "<20K",
                 "Don't know/No answer/Refused")
    )
  )

cat("=== Income AFTER reverse coding ===\n")
table(subset_data$INC_TOT_COM, useNA = "ifany")

# Housing condensation
subset_data <- subset_data %>%
  mutate(ENV_HMPRB_CON_MCQ = factor(ENV_HMPRB_CON_MCQ,
                                    levels = c(0, 1),
                                    labels = c("No", "Yes")))

# Alcohol
subset_data$ALC_FREQ_COM[subset_data$ALC_FREQ_COM %in% c(-8888, -9999, 77, 98, 99)] <- NA
subset_data <- subset_data %>%
  mutate(
    ALC_FREQ_COM = factor(
      ALC_FREQ_COM,
      levels = c(96, 7, 6, 5, 4, 3, 2, 1),
      labels = c("Never",
                 "Less than once a month",
                 "About once a month",
                 "2-3 times a month",
                 "Once a week",
                 "2-3 times a week",
                 "4-5 times a week",
                 "Almost every day (incl. 6 times a week)"),
      ordered = TRUE)
  )
table(subset_data$ALC_FREQ_COM)

# Smoking
subset_data$ICQ_SMOKE_COM[subset_data$ICQ_SMOKE_COM %in% c(-8888, -9999)] <- NA
subset_data <- subset_data %>%
  mutate(ICQ_SMOKE_COM = factor(ICQ_SMOKE_COM,
                                levels = c(1, 2, 3),
                                labels = c("Current smoker",
                                           "Never smoker",
                                           "Former smoker")))
table(subset_data$ICQ_SMOKE_COM)

# ============= Variable Labels =============
var_labels <- list(
  AGE_NMBR_COM      ~ "Age (years)",
  SEX_ASK_COM       ~ "Sex",
  WGHTS_PROV_COM    ~ "Province",
  Race7             ~ "Self-reported ethnicity",
  INC_TOT_COM       ~ "Household income",
  GrimAge_EAA       ~ "GrimAge EAA",
  PhenoAge_EAA      ~ "PhenoAge EAA",
  DunedinPACE       ~ "DunedinPACE",
  DNAmIC            ~ "DNAmIC",
  PM25DALYY_01_COM  ~ "PM2.5, 1-year average (µg/m³)",
  no2_1yr_com       ~ "NO₂, 1-year average (µg/m³)",
  so2_1yr_com       ~ "SO₂, 1-year average (µg/m³)",
  O3CHGYY_01_COM    ~ "Ozone, 1-year average (µg/m³)",
  GRLANYY_26_COM    ~ "NDVI",
  ENV_HMPRB_CON_MCQ ~ "Housing condensation",
  grtcc10_01_COM    ~ "Tree canopy cover (2010)",
  grtcc10_04_COM    ~ "Tree canopy cover, mean within 500m (2010)",
  GRUMPYY_01_COM    ~ "Vegetative greenness, annual % (postal code)",
  GRUMPYY_04_COM    ~ "Vegetative greenness, mean within 500m",
  ICQ_SMOKE_COM     ~ "Smoking",
  ALC_FREQ_COM      ~ "Alcohol use",
  PA2_DSCR2_MCQ     ~ "PASE scale score final"
)
```










## Missingness 

```{r}
#install.packages("DataExplorer")
library(DataExplorer)
library(mice)
library(ggplot2)
plot_missing(subset_data)
md.pattern(subset_data)


colSums(is.na(subset_data)) #how many NAs for each variable 

p <- plot_missing(subset_data)
ggsave("Figures/missing_plot.png", plot = p, width = 12, height = 7, dpi = 600)



missing_table <- data.frame(
  Variable = c(
    "AGE_NMBR_COM", "SEX_ASK_COM", "WGHTS_PROV_COM", "Race7",
    "INC_TOT_COM", "WEA_SVNGSVL_MCQ",
    "GrimAge_EAA", "PhenoAge_EAA", "DunedinPACE", "DNAmIC",
    "grtcc10_01_COM", "grtcc10_04_COM",
    "GRUMPYY_01_COM", "GRUMPYY_04_COM",
    "GRLANYY_26_COM", "ENV_HMPRB_CON_MCQ",
    "PM25DALYY_01_COM", "O3CHGYY_01_COM", "SO2OMIYY_COM",
    "CD4nv", "CD8nv", "Mono", "Treg", "CD4mem", "CD8mem", "NK", "Bmem", "Bnv"
  ),
  Description = c(
    "Age", "Sex", "Province", "Self-reported ethnicity",
    "Household income", "Household savings",
    "GrimAge", "PhenoAge", "DunedinPACE", "DNAmIC",
    "Tree canopy cover at postal code (2010)",
    "Tree canopy cover within 500m (2010)",
    "Vegetative greenness at postal code (annual %)",
    "Vegetative greenness within 500m (mean annual %)",
    "NDVI (vegetative greenness, 500m buffer)", "Housing condensation",
    "PM2.5 concentration (ug/m³, annual avg)", "Ozone (ppb, annual avg)", "SO2 concentration (ppb, 3-yr avg)",
    "Naïve CD4+ T cells", "Naïve CD8+ T cells", "Monocytes", "Regulatory T cells",
    "Memory CD4+ T cells", "Memory CD8+ T cells", "Natural Killer cells", "Memory B cells", "Naïve B cells"
  )
)

vars <- missing_table$Variable
missing_table$N_Missing <- sapply(vars, function(v) sum(is.na(subset_data[[v]])))
missing_table$Pct_Missing <- round(missing_table$N_Missing / nrow(subset_data) * 100, 1)

knitr::kable(missing_table,
             col.names = c("Variable", "Description", "N Missing", "% Missing"),
             caption = "Table X: Missing data by variable")


library(kableExtra)

missing_table_formatted <- knitr::kable(missing_table,
               col.names = c("Variable", "Description", "N Missing", "% Missing"),
               caption = "Table X: Missing data by variable",
               format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

save_kable(missing_table_formatted, "Figures/missing_table.png", zoom = 2)

```




## Interaction btw age and income 

```{r}

#check structure 
str(subset_data$AGE_NMBR_COM)
str(subset_data$INC_TOT_COM)

model <- aov(AGE_NMBR_COM ~ INC_TOT_COM, data = subset_data)
summary(model)

```



#2. Descriptive stats

## Table 1

```{r}
table1 <- subset_data %>%
  mutate(INC_TOT_COM = recode(INC_TOT_COM,
    "Don't know/No answer/Refused" = "DK/NA/R"
  )) %>%
  select(
    # Demographics
    AGE_NMBR_COM, SEX_ASK_COM, WGHTS_PROV_COM, Race7,
    # Income
    INC_TOT_COM,
    # Air pollution and housing
    PM25DALYY_01_COM, O3CHGYY_01_COM, no2_1yr_com, so2_1yr_com, 
    GRLANYY_26_COM, ENV_HMPRB_CON_MCQ,
    # Tree canopy cover
    grtcc10_01_COM, grtcc10_04_COM,
    # Vegetative greenness
    GRUMPYY_01_COM, GRUMPYY_04_COM
  ) %>%
  tbl_summary(
    by = INC_TOT_COM,
    label = list(
      AGE_NMBR_COM ~ "Age (years)",
      SEX_ASK_COM ~ "Sex",
      WGHTS_PROV_COM ~ "Province",
      Race7 ~ "Self-reported ethnicity",
      PM25DALYY_01_COM ~ "PM2.5, 1-year average (µg/m³)",
      O3CHGYY_01_COM ~ "Ozone, 1-year average (ppb)",
      no2_1yr_com ~ "NO₂, 1-year average (ppb)",
      so2_1yr_com ~ "SO₂, 1-year average (ppb)",
      GRLANYY_26_COM ~ "NDVI",
      ENV_HMPRB_CON_MCQ ~ "Housing condensation",
      grtcc10_01_COM ~ "TCC at postal code",
      grtcc10_04_COM ~ "TCC mean within 500m",
      GRUMPYY_01_COM ~ "VGF at postal code",
      GRUMPYY_04_COM ~ "VGF mean within 500m"
    ),
    type = list(
      SEX_ASK_COM ~ "categorical",
      Race7 ~ "categorical",
      ENV_HMPRB_CON_MCQ ~ "categorical"
    ),
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "no"
  ) %>%
  add_overall(last = TRUE, col_label = "**Total** (n = {N})") %>%
  modify_header(all_stat_cols() ~ "**{level}** (n = {n})") %>%
  add_n(
    statistic = "{N_miss} ({p_miss}%)",
    col_label = "**Missing**"
  ) %>%
  modify_spanning_header(all_stat_cols() ~ "**Household Income**") %>%
 modify_footnote(all_stat_cols() ~ NA) %>%
  as_flex_table() %>%
  add_footer_lines("TCC: percent tree canopy cover of woody vegetation taller than 5 metres. VGF: vegetative greenness fraction, capturing all green vegetation regardless of height and vegetation density. NDVI: Normalized Difference Vegetation Index. DK/NA/R: Don't know, no answer, refused")

save_as_image(table1, path = "Figures/table1.png", res = 300)

```


# 3.0 Linear Regressions
## INCOME PREDICTING AGING (stratified by age)


```{r}

# recode income as numeric, highest to lowest
subset_data$income_trend <- as.numeric(factor(subset_data$INC_TOT_COM,
  levels = c("150K+", "100-150k", "50-100K", "20-50K", "<20K"),
  ordered = TRUE))

# stratify by age group
subset_data$age_group <- ifelse(subset_data$AGE_NMBR_COM < 65, "45-64", "65-85")
subset_data_firstgrp  <- subset_data[subset_data$age_group == "45-64", ]
subset_data_secondgrp <- subset_data[subset_data$age_group == "65-85", ]

library(broom)

dir.create("regression_csvs", showWarnings = FALSE)

# function to run model and extract scaled results
run_trend_model <- function(outcome, data_45, data_65, steps = 4) {
  
  formula <- as.formula(paste0(outcome, " ~ income_trend + SEX_ASK_COM + AGE_NMBR_COM + I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7 + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv"))
  
  # 45-64
  model_45   <- lm(formula, data = data_45)
  results_45 <- tidy(model_45, conf.int = TRUE, conf.level = 0.95)
  write.csv(results_45, file.path("regression_csvs", paste0(outcome, "_trend_results_45to64.csv")), row.names = FALSE)
  
  # 65-85
  model_65   <- lm(formula, data = data_65)
  results_65 <- tidy(model_65, conf.int = TRUE, conf.level = 0.95)
  write.csv(results_65, file.path("regression_csvs", paste0(outcome, "_trend_results_65to85.csv")), row.names = FALSE)
  
  # scale results
  get_scaled <- function(results, age_grp) {
    beta <- results$estimate[results$term == "income_trend"]
    se   <- results$std.error[results$term == "income_trend"]
    p    <- results$p.value[results$term == "income_trend"]
    data.frame(
      outcome   = outcome,
      age_group = age_grp,
      estimate  = beta * steps,
      conf.low  = (beta - 1.96 * se) * steps,
      conf.high = (beta + 1.96 * se) * steps,
      p.value   = p
    )
  }
  
  bind_rows(
    get_scaled(results_45, "45-64"),
    get_scaled(results_65, "65-85")
  )
}

# run for all outcomes
all_scaled <- bind_rows(
  run_trend_model("GrimAge_EAA",  subset_data_firstgrp, subset_data_secondgrp),
  run_trend_model("PhenoAge_EAA", subset_data_firstgrp, subset_data_secondgrp),
  run_trend_model("DunedinPACE",  subset_data_firstgrp, subset_data_secondgrp),
  run_trend_model("DNAmIC",       subset_data_firstgrp, subset_data_secondgrp)
)

print(all_scaled)
write.csv(all_scaled, file.path("regression_csvs", "all_trend_scaled_results.csv"), row.names = FALSE)

# format the scaled results into a nice table
table_trend <- all_scaled %>%
  mutate(
    sig_stars = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01  ~ "**",
      p.value < 0.05  ~ "*",
      TRUE            ~ ""
    ),
    Result = ifelse(
      outcome == "DNAmIC",
      paste0(formatC(estimate, format = "e", digits = 2),
             " (", formatC(conf.low, format = "e", digits = 2),
             ", ", formatC(conf.high, format = "e", digits = 2), ")", sig_stars),
      paste0(round(estimate, 2),
             " (", round(conf.low, 2),
             ", ", round(conf.high, 2), ")", sig_stars)
    )
  ) %>%
  select(age_group, outcome, Result) %>%
  pivot_wider(names_from = outcome, values_from = Result)

table_45 <- table_trend %>% filter(age_group == "45-64") %>% select(-age_group)
table_65 <- table_trend %>% filter(age_group == "65-85") %>% select(-age_group)

table_final_trend <- bind_rows(table_45, table_65)

save_kable(
  knitr::kable(table_final_trend,
               col.names = c("GrimAge EAA", "PhenoAge EAA", "DunedinPACE", "DNAmIC"),
               caption = "Table X: Association between household income (highest to lowest, per one-step increase) and epigenetic aging, stratified by age group",
               escape = FALSE) %>%
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) %>%
    pack_rows("45-64 years", 1, nrow(table_45)) %>%
    pack_rows("65-85 years", nrow(table_45) + 1, nrow(table_final_trend)) %>%
    footnote(general = "Estimates represent the difference in epigenetic age acceleration associated with moving from highest (150K+) to lowest (<20K) income (4 steps). * p < 0.05, ** p < 0.01, *** p < 0.001. Models adjusted for sex, age, age², province, ethnicity, and blood cell type proportions.Note: DNAmIC measures intrinsic capacity, where higher values indicate better health; therefore, a negative coefficient indicates that lower income is associated with reduced intrinsic capacity, consistent with accelerated aging."),
  "Figures/income_trend_table.png", zoom = 2
)

```







## INCOME PREDICTING ENVIRONMENT 

```{r}

# recode income as numeric, highest to lowest
subset_data$income_trend <- as.numeric(factor(subset_data$INC_TOT_COM,
  levels = c("150K+", "100-150k", "50-100K", "20-50K", "<20K"),
  ordered = TRUE))

# function for continuous environmental outcomes
run_env_model <- function(outcome, data) {
  
  formula <- as.formula(paste0(outcome, " ~ income_trend + SEX_ASK_COM + AGE_NMBR_COM + I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7"))
  
  model   <- lm(formula, data = data)
  results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
  write.csv(results, paste0("regression_csvs/", outcome, "_env_results.csv"), row.names = FALSE)
  
  beta <- results$estimate[results$term == "income_trend"]
  se   <- results$std.error[results$term == "income_trend"]
  p    <- results$p.value[results$term == "income_trend"]
  
  data.frame(
    outcome   = outcome,
    estimate  = beta * 4,
    conf.low  = (beta - 1.96 * se) * 4,
    conf.high = (beta + 1.96 * se) * 4,
    p.value   = p
  )
}

# run above fxn for all continuous environmental outcomes
all_env_results <- bind_rows(
  run_env_model("PM25DALYY_01_COM", subset_data),
  run_env_model("O3CHGYY_01_COM",   subset_data),
  run_env_model("no2_1yr_com",      subset_data),
  run_env_model("so2_1yr_com",      subset_data),
  run_env_model("GRLANYY_26_COM",   subset_data),
  run_env_model("grtcc10_01_COM",   subset_data),
  run_env_model("grtcc10_04_COM",   subset_data),
  run_env_model("GRUMPYY_01_COM",   subset_data),
  run_env_model("GRUMPYY_04_COM",   subset_data)
)

print(all_env_results)
write.csv(all_env_results, "regression_csvs/all_env_results.csv", row.names = FALSE)

# ---- logistic regression for housing condensation ----
formula_cond <- ENV_HMPRB_CON_MCQ ~ income_trend + SEX_ASK_COM + AGE_NMBR_COM + 
                I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7

model_cond   <- glm(formula_cond, data = subset_data, family = binomial())
results_cond <- tidy(model_cond, conf.int = TRUE, conf.level = 0.95)
write.csv(results_cond, "regression_csvs/ENV_HMPRB_CON_MCQ_env_results.csv", row.names = FALSE)

beta_cond <- results_cond$estimate[results_cond$term == "income_trend"]
se_cond   <- results_cond$std.error[results_cond$term == "income_trend"]
p_cond    <- results_cond$p.value[results_cond$term == "income_trend"]
steps <- 4

condensation_result <- data.frame(
  outcome   = "ENV_HMPRB_CON_MCQ",
  OR        = exp(beta_cond * steps),
  conf.low  = exp((beta_cond - 1.96 * se_cond) * steps),
  conf.high = exp((beta_cond + 1.96 * se_cond) * steps),
  p.value   = p_cond
)

print(condensation_result)
write.csv(condensation_result, "regression_csvs/condensation_result.csv", row.names = FALSE)

# make into a nice table 
env_table <- all_env_results %>%
  mutate(
    sig_stars = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01  ~ "**",
      p.value < 0.05  ~ "*",
      TRUE            ~ ""
    ),
    Result = paste0(round(estimate, 3),
                    " (", round(conf.low, 3),
                    ", ", round(conf.high, 3), ")", sig_stars),
    outcome = recode(outcome,
      "PM25DALYY_01_COM" = "PM2.5, 1-year average (µg/m³)",
      "O3CHGYY_01_COM"   = "Ozone, 1-year average (ppb)",
      "no2_1yr_com"      = "NO₂, 1-year average (ppb)",
      "so2_1yr_com"      = "SO₂, 1-year average (ppb)",
      "GRLANYY_26_COM"   = "Max 500m",
      "grtcc10_01_COM"   = "Tree canopy coverage % at postal code",
      "grtcc10_04_COM"   = "Tree canopy coverage % mean within 500m",
      "GRUMPYY_01_COM"   = "Vegetative greenness fraction at postal code",
      "GRUMPYY_04_COM"   = "Vegetative greenness fraction mean within 500m"
    )
  ) %>%
  select(outcome, Result)

# format condensation result (OR)
condensation_row <- condensation_result %>%
  mutate(
    sig_stars = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01  ~ "**",
      p.value < 0.05  ~ "*",
      TRUE            ~ ""
    ),
    outcome = "Yes vs No",
    Result  = paste0(round(OR, 2),
                     " (", round(conf.low, 2),
                     ", ", round(conf.high, 2), ")", sig_stars, " [OR]")
  ) %>%
  select(outcome, Result)

# combine into one table
env_table_final <- bind_rows(
  env_table[1:4, ],
  condensation_row,
  env_table[5, ],
  env_table[6:9, ]
)
save_kable(
  knitr::kable(env_table_final,
               col.names = c("Environmental Variable", "β (95% CI)"),
               caption = "Table X: Association between household income and environmental exposures (highest to lowest income, 4-step trend)",
               escape = FALSE) %>%
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) %>%
    pack_rows("Air Pollution", 1, 4) %>%
    pack_rows("Housing Condensation", 5, 5) %>%
    pack_rows("Normalized Difference Vegetation Index", 6, 6) %>%
    pack_rows("Tree Canopy Cover", 7, 8) %>%
    pack_rows("Vegetative Greenness", 9, 10) %>%
    footnote(general = "Estimates represent the difference in environmental exposure associated with moving from highest (150K+) to lowest (<20K) income (4 steps). * p < 0.05, ** p < 0.01, *** p < 0.001. Models adjusted for sex, age, age², province, and ethnicity. [OR] = odds ratio from logistic regression."),
  "Figures/env_income_table.png", zoom = 2
)

```








## ENVIRONMENT PREDICTING AGING

```{r}
library(broom)
library(tidyr)
library(dplyr)
library(kableExtra)

# ── Shared setup ──────────────────────────────────────────────────────────────

covariates <- "SEX_ASK_COM + AGE_NMBR_COM + I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7 +
               CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv"

outcomes <- c(
  GrimAge     = "GrimAge_EAA",
  PhenoAge    = "PhenoAge_EAA",
  DunedinPACE = "DunedinPACE",
  DNAmIC      = "DNAmIC"
)

run_model <- function(outcome_var, predictor, filepath) {
  model <- lm(as.formula(paste(outcome_var, "~", predictor, "+", covariates)),
              data = subset_data)
  write.csv(tidy(model, conf.int = TRUE, conf.level = 0.95), filepath, row.names = FALSE)
}

# ── Air pollution ─────────────────────────────────────────────────────────────

air_preds <- c(
  PM2.5 = "PM25DALYY_01_COM",
  Ozone = "O3CHGYY_01_COM",
  NO2   = "no2_1yr_com",
  SO2   = "so2_1yr_com"
)

for (env in names(air_preds)) {
  for (out in names(outcomes)) {
    run_model(outcomes[out], air_preds[env],
              paste0("regression_csvs/Air_", env, "_", out, ".csv"))
  }
}

# ── TCC ───────────────────────────────────────────────────────────────────────

tcc_preds <- setNames(c("grtcc10_01_COM", "grtcc10_04_COM"),
                      c("PostalCode", "500m"))

for (env in names(tcc_preds)) {
  for (out in names(outcomes)) {
    run_model(outcomes[out], tcc_preds[env],
              paste0("regression_csvs/TCC_", env, "_", out, ".csv"))
  }
}

# ── Vegetative Greenness ──────────────────────────────────────────────────────

green_preds <- setNames(c("GRUMPYY_01_COM", "GRUMPYY_04_COM"),
                        c("PostalCode", "Mean500m"))

for (env in names(green_preds)) {
  for (out in names(outcomes)) {
    run_model(outcomes[out], green_preds[env],
              paste0("regression_csvs/greenness_", env, "_", out, ".csv"))
  }
}

# ── Condensation ──────────────────────────────────────────────────────────────

for (out in names(outcomes)) {
  run_model(outcomes[out], "ENV_HMPRB_CON_MCQ",
            paste0("regression_csvs/Condensation_", out, ".csv"))
}

# ── NDVI ──────────────────────────────────────────────────────────────────────

for (out in names(outcomes)) {
  run_model(outcomes[out], "GRLANYY_26_COM",
            paste0("regression_csvs/NDVI_", out, ".csv"))
}

# ── Shared table stuff ────────────────────────────────────────────────────────

read_model <- function(file, env, outcome) {
  read.csv(file) %>% mutate(Env = env, Outcome = outcome)
}

aging_cols <- c("Exposure", "GrimAge EAA", "PhenoAge EAA", "DunedinPACE", "DNAmIC")

load_results <- function(prefix, envs, env_labels) {
  bind_rows(lapply(seq_along(envs), function(i) {
    bind_rows(lapply(names(outcomes), function(out) {
      read_model(paste0("regression_csvs/", prefix, envs[i], "_", out, ".csv"),
                 env_labels[i], out)
    }))
  }))
}

# ── Load all results ──────────────────────────────────────────────────────────

air_envs   <- c("PM2.5", "Ozone", "NO2", "SO2")
air_labels <- c("PM2.5, 1-year average (µg/m³)",
                "Ozone, 1-year average (ppb)",
                "NO₂, 1-year average (ppb)",
                "SO₂, 1-year average (ppb)")
air_terms  <- c("PM25DALYY_01_COM", "O3CHGYY_01_COM", "no2_1yr_com", "so2_1yr_com")
air_results <- load_results("Air_", air_envs, air_labels)

tcc_file_labels <- c("PostalCode", "500m")
tcc_env_labels  <- c("Tree canopy coverage, % at postal code",
                     "Tree canopy coverage, % mean within 500m")
tcc_terms       <- c("grtcc10_01_COM", "grtcc10_04_COM")
tcc_results     <- load_results("TCC_", tcc_file_labels, tcc_env_labels)

green_file_labels <- c("PostalCode", "Mean500m")
green_env_labels  <- c("Vegetative greenness fraction at postal code",
                       "Vegetative greenness fraction, mean within 500m")
green_terms       <- c("GRUMPYY_01_COM", "GRUMPYY_04_COM")
green_results     <- load_results("greenness_", green_file_labels, green_env_labels)

cond_results <- bind_rows(lapply(names(outcomes), function(out) {
  read_model(paste0("regression_csvs/Condensation_", out, ".csv"), "Yes vs No", out)
}))

ndvi_results <- bind_rows(lapply(names(outcomes), function(out) {
  read_model(paste0("regression_csvs/NDVI_", out, ".csv"), "Max NDVI within 500m", out)
}))

# ── Combine and format ────────────────────────────────────────────────────────

all_env_aging <- bind_rows(
  air_results   %>% filter(term %in% air_terms),
  cond_results  %>% filter(term == "ENV_HMPRB_CON_MCQYes"),
  ndvi_results  %>% filter(term == "GRLANYY_26_COM"),
  tcc_results   %>% filter(term %in% tcc_terms),
  green_results %>% filter(term %in% green_terms)
) %>%
  mutate(
    sig_stars = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01  ~ "**",
      p.value < 0.05  ~ "*",
      TRUE            ~ ""
    ),
    Result = ifelse(
      Outcome == "DNAmIC",
      paste0(formatC(estimate, format = "e", digits = 2),
             " (", formatC(conf.low, format = "e", digits = 2),
             ", ", formatC(conf.high, format = "e", digits = 2), ")", sig_stars),
      ifelse(
        abs(estimate) < 0.01,
        paste0(round(estimate, 4), " (", round(conf.low, 4), ", ", round(conf.high, 4), ")", sig_stars),
        paste0(round(estimate, 2), " (", round(conf.low, 2), ", ", round(conf.high, 2), ")", sig_stars)
      )
    )
  ) %>%
  select(Env, Outcome, Result) %>%
  pivot_wider(names_from = Outcome, values_from = Result)

# row counts for pack_rows
n_air   <- length(air_envs)
n_cond  <- 1
n_ndvi  <- 1
n_tcc   <- length(tcc_env_labels)
n_green <- length(green_env_labels)

save_kable(
  knitr::kable(all_env_aging,
               col.names = aging_cols,
               caption = "Table X: Association between environmental exposures and epigenetic aging",
               escape = FALSE) %>%
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) %>%
    pack_rows("Air Pollution", 1, n_air) %>%
    pack_rows("Housing Condensation", n_air + 1, n_air + n_cond) %>%
    pack_rows("Normalized Difference Vegetation Index", n_air + n_cond + 1, n_air + n_cond + n_ndvi) %>%
    pack_rows("Tree Canopy Cover", n_air + n_cond + n_ndvi + 1, n_air + n_cond + n_ndvi + n_tcc) %>%
    pack_rows("Vegetative Greenness", n_air + n_cond + n_ndvi + n_tcc + 1, n_air + n_cond + n_ndvi + n_tcc + n_green) %>%
    footnote(general = "* p < 0.05, ** p < 0.01, *** p < 0.001. Models adjusted for sex, age, age², province, ethnicity, and blood cell type proportions."),
  "Figures/env_aging_table.png", zoom = 2
)

```




































# 4.0 Plots


General Code needed for plots 

```{r}
#Custom theme for all figures

report_theme <- theme_classic(base_size = 11) +
  theme(
    legend.position = "none",
    
    plot.title = element_text(
      hjust = 0.5,
      face  = "bold",
      size  = 13,
      margin = margin(b = 8)
    ),
    
    axis.title.x = element_text(
      face = "bold",
      size = 11,
      margin = margin(t = 6)
    ),
    axis.title.y = element_text(
      face = "bold",
      size = 11,
      margin = margin(r = 6)
    ),
    
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10)
  )

```


### demographics

```{r}

subset_data <- readRDS("subset_data.rds")

# age and sex  ------------------------------------------------------

#pie chart for sex
df_sex <- subset_data %>%
  filter(!is.na(SEX_ASK_COM)) %>%
  count(SEX_ASK_COM) %>%
  mutate(
    percent = n / sum(n),
    label = paste0(SEX_ASK_COM, " (", round(percent * 100, 1), "%)")
  )

ggplot(df_sex, aes(x = 1, y = n, fill = SEX_ASK_COM)) +
  geom_col(width = 1, color = "white", linewidth = 1) +
  coord_polar(theta = "y") +
  geom_text(
    aes(label = label),
    position = position_stack(vjust = 0.5),
    size = 10,              # big, slide-friendly
    fontface = "plain"
  ) +
  scale_fill_manual(values = c("F" = "#1f77b4", "M" = "#ff7f0e")) +
  labs(title = "Sex Distribution", x = NULL, y = NULL, fill = NULL) +
  theme_void() +
  theme(
    # remove any remaining axis/radial artifacts
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    
    plot.title = element_text(
      hjust = 0.5,
      face  = "bold",
      size  = 38,
      margin = margin(b = 18)
    ),
    
    legend.position = "right",
    legend.text = element_text(size = 26),
    legend.key.size = grid::unit(1.2, "lines")
  )


#age boxplot 
subset_data %>%
  filter(!is.na(AGE_NMBR_COM)) %>%
  ggplot(aes(x = "", y = AGE_NMBR_COM)) +
  geom_boxplot(
    width = 0.25,
    fill = "white",
    color = "#2C7FB8",
    linewidth = 1.2
  ) +
  scale_y_continuous(
    breaks = seq(
      floor(min(subset_data$AGE_NMBR_COM, na.rm = TRUE) / 5) * 5,
      ceiling(max(subset_data$AGE_NMBR_COM, na.rm = TRUE) / 5) * 5,
      by = 5
    )
  ) +
  labs(
    title = "Age Distribution",
    x = NULL,
    y = "Age (years)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(
      hjust  = 0.5,
      face   = "bold",
      size   = 38,
      margin = margin(b = 18)
    ),
    axis.title.y = element_text(
      face   = "bold",
      size   = 30,
      margin = margin(r = 16)
    ),
    axis.title.x = element_blank(),
    axis.text.y  = element_text(size = 26),
    axis.text.x  = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_line(linetype = "dotted"),
    panel.grid.major.y = element_line(linetype = "dashed"),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background  = element_rect(fill = "white", color = NA)
  )


ggsave("age_distribution.png", width = 12, height = 7, dpi = 600)

# categories of cultural background:"White", "South Asian", "Black", "Asian","Arab", "Latin American", "Other"
subset_data %>%
  count(Race7) %>%
  mutate(
    percent = round(100 * n / sum(n), 1)
  )

  
```


```{r}

# ethnicity/self reported background barplot 
p_ethnicity <- subset_data %>%
  filter(!is.na(Race7)) %>%
  count(Race7, name = "n") %>%
  arrange(n) %>%
  ggplot(aes(x = n, y = Race7)) +
  geom_col(fill = "#2C7FB8", width = 0.7) +
  labs(
    title = "Self-Reported Cultural Background",
    x = "Number of Participants",
    y = NULL
  ) +
  report_theme

ggsave("Figures/ethnicity_barplot.png", plot = p_ethnicity, width = 5, height = 4, dpi = 300)
```





### Socioeconomic variables: Household Income and Wealth 

```{r}

#summary table of household income
#after reversing: 1 = $150,000 or more ; 2 = $100,000–$149,999 ; 3 = $50,000–$99,999; 4 = $20,000–$49,999; 5 = Less than $20,000; 6=  Don't know/No answer/Refused

subset_data %>%
  count(INC_TOT_COM) %>%
  mutate(
    percent = round(100 * n / sum(n), 1)
  )


# household income barplot 
p_income <- subset_data %>%
  count(income_cat, name = "n") %>%
  ggplot(aes(x = n, y = income_cat)) +
  geom_col(fill = "#2C7FB8", width = 0.7) +
  labs(
    title = "Household Income (Past 12 Months)",
    x = "Number of Participants",
    y = NULL
  ) +
  report_theme

ggsave("Figures/household_income_barplot.png", plot = p_income, width = 5, height = 4, dpi = 300)




#savings categories after reverse code are: 1=1 million or more. 2=100k to 1 million. 3=50-100k. 4=<50K. 5=don't know/no answer/refused 
subset_data %>%
  count(WEA_SVNGSVL_MCQ) %>%
  mutate(
    percent = round(100 * n / sum(n), 1)
  )

# savings/wealth barplot 
p_wealth <- subset_data %>%
  mutate(
    wealth_cat = factor(
      WEA_SVNGSVL_MCQ,
      levels = c(1, 2, 3, 4, 8, 9),
      labels = c(
        "<50K",
        "50K–100K",
        "100K to <$1M",
        "$1M+",
        "Don't know/No answer",
        "Refused"
      )
    )
  ) %>%
  count(wealth_cat, name = "n") %>%
  arrange(n) %>%
  ggplot(aes(x = n, y = wealth_cat)) +
  geom_col(fill = "#2C7FB8", width = 0.7) +
  labs(
    title = "Total Savings and Investments",
    x = "Number of Participants",
    y = NULL
  ) +
  report_theme

ggsave("Figures/savings_barplot.png", plot = p_wealth, width = 5, height = 4, dpi = 300)


```



# Plots_enviro  

## Air pollution 


### Compare PM25DALE_AYY_01 and pm25_1yr_com etc. 

PM25DALYY_01_COM
SO2OMIYY_COM
O3CHGYY_01_COM

```{r}
library(ggplot2)
library(tidyr)
library(dplyr)


# ── PM2.5 ──────────────────────────────────────────────────────────────────────
df_long <- subset_data %>%
  select(PM25DALYY_01_COM, pm25_1yr_com) %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") %>%
  filter(!is.na(value))

p_pm25_density <- ggplot(df_long, aes(x = value, fill = variable, color = variable)) +
  geom_density(alpha = 0.4) +
  labs(title = "Distribution Comparison: PM25DALYY_01_COM vs pm25_1yr_com",
       x = "PM2.5 Value", y = "Density") +
  theme_minimal()
ggsave("Figures/pm25_comparison_density.png", plot = p_pm25_density, width = 5, height = 4, dpi = 300)

p_pm25_boxplot <- ggplot(df_long, aes(x = variable, y = value, fill = variable)) +
  geom_boxplot(alpha = 0.6, outlier.shape = 21) +
  labs(title = "Boxplot Comparison: PM25DALYY_01_COM vs pm25_1yr_com",
       x = "", y = "PM2.5 Value") +
  theme_minimal()
ggsave("Figures/pm25_comparison_boxplot.png", plot = p_pm25_boxplot, width = 5, height = 4, dpi = 300)

summary(subset_data[, c("PM25DALYY_01_COM", "pm25_1yr_com")])

# ── SO2 ────────────────────────────────────────────────────────────────────────
df_long <- subset_data %>%
  select(so2_1yr_com, SO2OMIYY_COM) %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") %>%
  filter(!is.na(value))

p_so2_density <- ggplot(df_long, aes(x = value, fill = variable, color = variable)) +
  geom_density(alpha = 0.4) +
  labs(title = "Distribution Comparison: so2_1yr_com vs SO2OMIYY_COM",
       x = "SO2 Value", y = "Density") +
  theme_minimal()
ggsave("Figures/so2_comparison_density.png", plot = p_so2_density, width = 5, height = 4, dpi = 300)

p_so2_boxplot <- ggplot(df_long, aes(x = variable, y = value, fill = variable)) +
  geom_boxplot(alpha = 0.6, outlier.shape = 21) +
  labs(title = "Boxplot Comparison: so2_1yr_com vs SO2OMIYY_COM",
       x = "", y = "SO2 Value") +
  theme_minimal()
ggsave("Figures/so2_comparison_boxplot.png", plot = p_so2_boxplot, width = 5, height = 4, dpi = 300)

summary(subset_data[, c("so2_1yr_com", "SO2OMIYY_COM")])

# ── Ozone ──────────────────────────────────────────────────────────────────────
df_long <- subset_data %>%
  select(O3CHGYY_01_COM, o3_1yr_com) %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") %>%
  filter(!is.na(value))

p_o3_density <- ggplot(df_long, aes(x = value, fill = variable, color = variable)) +
  geom_density(alpha = 0.4) +
  labs(title = "Distribution Comparison: O3CHGYY_01_COM vs o3_1yr_com",
       x = "O3 Value", y = "Density") +
  theme_minimal()
ggsave("Figures/o3_comparison_density.png", plot = p_o3_density, width = 5, height = 4, dpi = 300)

p_o3_boxplot <- ggplot(df_long, aes(x = variable, y = value, fill = variable)) +
  geom_boxplot(alpha = 0.6, outlier.shape = 21) +
  labs(title = "Boxplot Comparison: O3CHGYY_01_COM vs o3_1yr_com",
       x = "", y = "O3 Value") +
  theme_minimal()
ggsave("Figures/o3_comparison_boxplot.png", plot = p_o3_boxplot, width = 5, height = 4, dpi = 300)

summary(subset_data[, c("O3CHGYY_01_COM", "o3_1yr_com")])

```



```{r}
ft <- flextable(summary_table) %>%
  add_header_lines("Table X. Air Quality Exposures") %>%
  bold(part = "header") %>%
  bold(j = "Variable") %>%
  autofit() %>%
  theme_booktabs() %>%
  fontsize(size = 10, part = "all") %>%
  fontsize(i = 1, size = 11, part = "header") %>%  # title slightly bigger
  bold(i = 1, part = "header") %>%                 # title bold
  font(fontname = "Times New Roman", part = "all") %>%
  align(i = 1, align = "center", part = "header") %>%  # title centered
  align(align = "left", part = "all") %>%
  align(j = c("Mean (SD)", "Median [Min, Max]", "Missing"), align = "center", part = "body")

save_as_image(ft, path = "Tables/Table_airquality.png", res = 300)
```

```{r}
#air pollution dashboard 
air_vars <- list(
  list(col = "pm25_1yr_com",  title = "PM2.5 (µg/m³)", file = "pm25",  binwidth = 0.5, limits = c(3, 11),  breaks = seq(3, 11, by = 1)),
  list(col = "no2_1yr_com",   title = "NO2 (ppb)",      file = "no2",   binwidth = 1,   limits = c(3, 16),  breaks = seq(3, 16, by = 2)),
  list(col = "so2_1yr_com",   title = "SO2 (ppb)",      file = "so2",   binwidth = 0.1, limits = c(0, 3.2), breaks = seq(0, 3.2, by = 0.5)),
  list(col = "o3_1yr_com",    title = "Ozone (ppb)",    file = "ozone", binwidth = 1,   limits = c(15, 33), breaks = seq(15, 33, by = 2))
)

air_plots <- list()

for (v in air_vars) {
  p <- ggplot(subset_data, aes(x = .data[[v$col]])) +
    geom_histogram(binwidth = v$binwidth, fill = "steelblue", color = "white") +
    scale_x_continuous(breaks = v$breaks, limits = v$limits) +
    labs(title = v$title, x = v$title, y = "Count") +
    report_theme +
    theme(plot.title = element_text(size = 9))
  
  air_plots[[v$col]] <- p
  ggsave(paste0("Figures/", v$file, "_histogram.png"), plot = p, width = 5, height = 4, dpi = 300)
}

dashboard_air <- wrap_plots(air_plots, ncol = 2) +
  plot_annotation(
    title = "Air Pollution Distributions",
    theme = theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 14))
  )

ggsave("Figures/air_pollution_dashboard.png", plot = dashboard_air, width = 10, height = 8, dpi = 300)

```













## Condensation/noise

```{r}
## 3.3 Condensation and noise 
vars_housing <- list(
  list(col = "ENV_HMPRB_CON_MCQ", label = "Condensation"),
  list(col = "ENV_HMPRB_NOI_MCQ", label = "Noise")
)

summary_table_housing <- lapply(vars_housing, function(v) {
  x <- subset_data[[v$col]]
  tibble(
    Variable  = v$label,
    `Yes n (%)` = paste0(sum(x == 1, na.rm=TRUE), " (", round(mean(x == 1, na.rm=TRUE)*100, 1), "%)"),
    `No n (%)`  = paste0(sum(x == 0, na.rm=TRUE), " (", round(mean(x == 0, na.rm=TRUE)*100, 1), "%)"),
    `Missing`   = paste0(sum(is.na(x)), " (", round(mean(is.na(x))*100, 1), "%)")
  )
}) %>% bind_rows()

ft_housing <- flextable(summary_table_housing) %>%
  add_header_lines("Table X. Problems with Housing") %>%
  bold(part = "header") %>%
  bold(j = "Variable") %>%
  autofit() %>%
  theme_booktabs() %>%
  fontsize(size = 10, part = "all") %>%
  fontsize(i = 1, size = 11, part = "header") %>%
  bold(i = 1, part = "header") %>%
  font(fontname = "Times New Roman", part = "all") %>%
  align(i = 1, align = "center", part = "header") %>%
  align(align = "left", part = "all") %>%
  align(j = c("Yes n (%)", "No n (%)", "Missing"), align = "center", part = "body")

save_as_image(ft_housing, path = "Tables/Table_housing.png", res = 300)
```


```{r}

#Problems with condensation
condensation_bar <- ggplot(subset_data, aes(x = factor(ENV_HMPRB_CON_MCQ))) +
  geom_bar(fill = "steelblue", color = "white") +
  geom_text(
    aes(label = paste0(round(after_stat(count/sum(count)*100), 1), "%")),
    stat = "count",
    vjust = -0.5,
    size = 4
  ) +
  scale_x_discrete(labels = c("0" = "No", "1" = "Yes")) +
  labs(
    title = "Problems with Condensation in Home",
    x = "Condensation",
    y = "Count"
  ) +
  report_theme +
  theme(plot.margin = margin(t = 20, r = 10, b = 10, l = 10)) +
  coord_cartesian(clip = "off")

ggsave("Figures/condensation_bar.png", plot = condensation_bar, width = 5, height = 4, dpi = 300)




#Problems with noise

noise_bar <- ggplot(subset_data, aes(x = factor(ENV_HMPRB_NOI_MCQ))) +
  geom_bar(fill = "steelblue", color = "white") +
  geom_text(
    aes(label = paste0(round(after_stat(count/sum(count)*100), 1), "%")),
    stat = "count",
    vjust = -0.5,
    size = 4
  ) +
  scale_x_discrete(labels = c("0" = "No", "1" = "Yes")) +
  labs(
    title = "Problems with Noise",
    x = "Noise",
    y = "Count"
  ) +
  report_theme +
  theme(plot.margin = margin(t = 20, r = 10, b = 10, l = 10)) +
  coord_cartesian(clip = "off")

ggsave("Figures/noise_bar.png", plot = noise_bar, width = 5, height = 4, dpi = 300)



```












## Tree Canopy Coverage (TCC)
```{r}
# Table 1 - Tree Canopy Coverage
vars_canopy <- list(
  list(col = "grtcc10_01_COM", label = "Mean % of pixel covered by vegetation >5m"),
  list(col = "grtcc10_02_COM", label = "Mean % of pixel covered by vegetation >5m, within 100m"),
  list(col = "grtcc10_03_COM", label = "Mean % of pixel covered by vegetation >5m, within 250m"),
  list(col = "grtcc10_04_COM", label = "Mean % of pixel covered by vegetation >5m, within 500m"),
  list(col = "grtcc10_05_COM", label = "Mean % of pixel covered by vegetation >5m, within 1km")
)

summary_table_canopy <- lapply(vars_canopy, function(v) {
  x <- subset_data[[v$col]]
  tibble(
    Variable            = v$label,
    `Mean (SD)`         = paste0(round(mean(x, na.rm=TRUE), 2), " (", round(sd(x, na.rm=TRUE), 2), ")"),
    `Median [Min, Max]` = paste0(round(median(x, na.rm=TRUE), 2), " [", round(min(x, na.rm=TRUE), 2), ", ", round(max(x, na.rm=TRUE), 2), "]"),
    `Missing`           = paste0(sum(is.na(x)), " (", round(mean(is.na(x))*100, 1), "%)")
  )
}) %>% bind_rows()

ft_canopy <- flextable(summary_table_canopy) %>%
  add_header_lines("Table X. Tree Canopy Coverage") %>%
  bold(part = "header") %>%
  bold(j = "Variable") %>%
  autofit() %>%
  theme_booktabs() %>%
  fontsize(size = 10, part = "all") %>%
  fontsize(i = 1, size = 11, part = "header") %>%
  flextable::bold(i = 1, part = "header") %>%
  font(fontname = "Times New Roman", part = "all") %>%
  align(i = 1, align = "center", part = "header") %>%
  align(align = "left", part = "all") %>%
  align(j = c("Mean (SD)", "Median [Min, Max]", "Missing"), align = "center", part = "body")

save_as_image(ft_canopy, path = "Tables/Table_treecanopy.png", res = 300)

```



```{r}
# Tree canopy histogram dashboard 
canopy_vars <- list(
  list(col = "grtcc10_01_COM", title = "Estimated % Pixel Covered by Vegetation >5m",           file = "canopy_01"),
  list(col = "grtcc10_02_COM", title = "Mean % Pixel Covered by Vegetation >5m within 100m",    file = "canopy_02"),
  list(col = "grtcc10_03_COM", title = "Mean % Pixel Covered by Vegetation >5m within 250m",    file = "canopy_03"),
  list(col = "grtcc10_04_COM", title = "Mean % Pixel Covered by Vegetation >5m within 500m",    file = "canopy_04"),
  list(col = "grtcc10_05_COM", title = "Mean % Pixel Covered by Vegetation >5m within 1km",     file = "canopy_05")
)

for (v in canopy_vars) {# Save individual histograms
  p <- ggplot(subset_data, aes(x = .data[[v$col]])) +
    geom_histogram(binwidth = 5, fill = "forestgreen", color = "white") +
    scale_x_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60, 70, 80)) +
    labs(title = v$title, x = "Tree Canopy Cover (%)", y = "Count") +
    report_theme
  ggsave(paste0("Figures/", v$file, "_histogram.png"), plot = p, width = 5, height = 4, dpi = 300)
}

canopy_plots <- list()# Dashboard

for (v in canopy_vars) {
  p <- ggplot(subset_data, aes(x = .data[[v$col]])) +
    geom_histogram(binwidth = 5, fill = "forestgreen", color = "white") +
    scale_x_continuous(breaks = c(0, 20, 40, 60, 80)) +
    labs(title = v$title, x = "Tree Canopy Cover (%)", y = "Count") +
    report_theme +
    theme(plot.title = element_text(size = 8))
  
  canopy_plots[[v$col]] <- p
}

dashboard_canopy <- wrap_plots(canopy_plots, ncol = 3) +# 5 plots — arrange in 2 rows (3 top, 2 bottom)
  plot_annotation(
    title = "Tree Canopy Coverage Distributions",
    theme = theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 14))
  )

ggsave("Figures/tree_canopy_dashboard.png", plot = dashboard_canopy, width = 15, height = 9, dpi = 300)


```




## Percent vegetation 

```{r}
# Table 2 - Vegetative Greenness
vars_greenness <- list(
  list(col = "GRUMPYY_01_COM", label = "Mean of annual % vegetative greenness, At postal code"),
  list(col = "GRUMPYY_02_COM", label = "Mean of annual % vegetative greenness, within 100m"),
  list(col = "GRUMPYY_03_COM", label = "Mean of annual % vegetative greenness, within 250m"),
  list(col = "GRUMPYY_04_COM", label = "Mean of annual % vegetative greenness, within 500m"),
  list(col = "GRUMPYY_05_COM", label = "Mean of annual % vegetative greenness, within 1km"),
  list(col = "GRUMPYY_06_COM", label = "Max of annual % vegetative greenness, within 100m"),
  list(col = "GRUMPYY_07_COM", label = "Max of annual % vegetative greenness, within 250m"),
  list(col = "GRUMPYY_08_COM", label = "Max of annual % vegetative greenness, within 500m"),
  list(col = "GRUMPYY_09_COM", label = "Max of annual % vegetative greenness, within 1km")
)

summary_table_greenness <- lapply(vars_greenness, function(v) {
  x <- subset_data[[v$col]]
  tibble(
    Variable            = v$label,
    `Mean (SD)`         = paste0(round(mean(x, na.rm=TRUE), 2), " (", round(sd(x, na.rm=TRUE), 2), ")"),
    `Median [Min, Max]` = paste0(round(median(x, na.rm=TRUE), 2), " [", round(min(x, na.rm=TRUE), 2), ", ", round(max(x, na.rm=TRUE), 2), "]"),
    `Missing`           = paste0(sum(is.na(x)), " (", round(mean(is.na(x))*100, 1), "%)")
  )
}) %>% bind_rows()

ft_greenness <- flextable(summary_table_greenness) %>%
  add_header_lines("Table X. Percent Vegetative Greenness") %>%
  bold(part = "header") %>%
  bold(j = "Variable") %>%
  autofit() %>%
  theme_booktabs() %>%
  fontsize(size = 10, part = "all") %>%
  fontsize(i = 1, size = 11, part = "header") %>%
  bold(i = 1, part = "header") %>%
  font(fontname = "Times New Roman", part = "all") %>%
  align(i = 1, align = "center", part = "header") %>%
  align(align = "left", part = "all") %>%
  align(j = c("Mean (SD)", "Median [Min, Max]", "Missing"), align = "center", part = "body")

save_as_image(ft_greenness, path = "Tables/Table_percentveg.png", res = 300)
```

```{r}

#Percent veg greenness histograms (9 of them)

greenness_vars <- list(
  list(col = "GRUMPYY_01_COM", title = "Annual % Vegetative Greenness at Postal Code",        file = "greenness_01"),
  list(col = "GRUMPYY_02_COM", title = "Mean Annual % Vegetative Greenness within 100m",       file = "greenness_02"),
  list(col = "GRUMPYY_03_COM", title = "Mean Annual % Vegetative Greenness within 250m",       file = "greenness_03"),
  list(col = "GRUMPYY_04_COM", title = "Mean Annual % Vegetative Greenness within 500m",       file = "greenness_04"),
  list(col = "GRUMPYY_05_COM", title = "Mean Annual % Vegetative Greenness within 1000m",      file = "greenness_05"),
  list(col = "GRUMPYY_06_COM", title = "Max Annual % Vegetative Greenness within 100m",        file = "greenness_06"),
  list(col = "GRUMPYY_07_COM", title = "Max Annual % Vegetative Greenness within 250m",        file = "greenness_07"),
  list(col = "GRUMPYY_08_COM", title = "Max Annual % Vegetative Greenness within 500m",        file = "greenness_08"),
  list(col = "GRUMPYY_09_COM", title = "Max Annual % Vegetative Greenness within 1000m",       file = "greenness_09")
)

for (v in greenness_vars) {
  p <- ggplot(subset_data, aes(x = .data[[v$col]])) +
    geom_histogram(binwidth = 5, fill = "forestgreen", color = "white") +
    scale_x_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90)) +
    labs(title = v$title, x = "Vegetative Greenness (%)", y = "Count") +
    report_theme
  ggsave(paste0("Figures/", v$file, "_histogram.png"), plot = p, width = 5, height = 4, dpi = 300)
}

library(patchwork)

# Use your loop to store plots in a list instead of saving individually
greenness_plots <- list()

for (v in greenness_vars) {
  p <- ggplot(subset_data, aes(x = .data[[v$col]])) +
    geom_histogram(binwidth = 5, fill = "forestgreen", color = "white") +
    scale_x_continuous(breaks = c(0, 20, 40, 60, 80, 100)) +
    labs(title = v$title, x = "Greenness (%)", y = "Count") +
    report_theme +
    theme(plot.title = element_text(size = 8))  # smaller title to fit
  
  greenness_plots[[v$col]] <- p
}

# Arrange in a 3x3 grid
dashboard <- wrap_plots(greenness_plots, ncol = 3) +
  plot_annotation(
    title = "Vegetative Greenness Distributions",
    theme = theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 14))
  )

ggsave("Figures/greenness_dashboard.png", plot = dashboard, width = 15, height = 12, dpi = 300)




ndvi_500m <- ggplot(subset_data, aes(x = GRLANYY_26_COM)) +
  geom_histogram(binwidth = 5, fill = "forestgreen", color = "white") +
  scale_x_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100)) +
  labs(
    title = "Maximum Annual NDVI within 500m",
    x = "NDVI (%)",
    y = "Count"
  ) +
  report_theme

ggsave("Figures/ndvi_500m_histogram.png", plot = ndvi_500m, width = 5, height = 4, dpi = 300)
GRLANYY_26_COM
```



## NDVI 


```{r}
vars_ndvi <- list(
  list(col = "GRLANYY_26_COM", label = "Maximum annual mean NVDI within 250m")
)

summary_table_ndvi <- lapply(vars_ndvi, function(v) {
  x <- subset_data[[v$col]]
  tibble(
    Variable            = v$label,
    `Mean (SD)`         = paste0(round(mean(x, na.rm=TRUE), 2), " (", round(sd(x, na.rm=TRUE), 2), ")"),
    `Median [Min, Max]` = paste0(round(median(x, na.rm=TRUE), 2), " [", round(min(x, na.rm=TRUE), 2), ", ", round(max(x, na.rm=TRUE), 2), "]"),
    `Missing`           = paste0(sum(is.na(x)), " (", round(mean(is.na(x))*100, 1), "%)")
  )
}) %>% bind_rows()

ft_ndvi <- flextable(summary_table_ndvi) %>%
  add_header_lines("Table X. Normalized Difference Vegetation Index (NDVI)") %>%
  bold(part = "header") %>%
  bold(j = "Variable") %>%
  autofit() %>%
  theme_booktabs() %>%
  fontsize(size = 10, part = "all") %>%
  fontsize(i = 1, size = 11, part = "header") %>%
  bold(i = 1, part = "header") %>%
  font(fontname = "Times New Roman", part = "all") %>%
  align(i = 1, align = "center", part = "header") %>%
  align(align = "left", part = "all") %>%
  align(j = c("Mean (SD)", "Median [Min, Max]", "Missing"), align = "center", part = "body")

save_as_image(ft_ndvi, path = "Tables/Table_ndvi.png", res = 300)
```


```{r}
ndvi_500m <- ggplot(subset_data, aes(x = GRLANYY_26_COM)) +
  geom_histogram(binwidth = 0.02, fill = "forestgreen", color = "white") +
  scale_x_continuous(
    breaks = seq(0.60, 1.00, by = 0.05),
    limits = c(0.60, 1.00)
  ) +
  labs(
    title = "Maximum Annual NDVI within 500m",
    x = "NDVI",
    y = "Count"
  ) +
  report_theme

ggsave("Figures/ndvi_500m_histogram.png", plot = ndvi_500m, width = 5, height = 4, dpi = 300)

```









# Plots_Enviro by income (Pollution, condensation) 

## Pollution by income 
```{r}

#pollution boxplots by income group 


plot_data <- subset_data
ggplot(data = plot_data, aes(x = income_cat, y = pm25_1yr_com)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(
    title = "PM2.5 1 y average, by Income Group",
    x = "Household Income",
    y = "PM2.5 (µg/m³)"
  ) +
  report_theme

ggsave("Figures/boxplot_pm25.png", width = 5, height = 4, dpi = 300)



df %>% #check that there is significance for PM2.5 
  group_by(income_cat) %>%
  summarise(
    n = n(),
    median_pm25 = median(pm25_1yr_com, na.rm = TRUE),
    mean_pm25 = round(mean(pm25_1yr_com, na.rm = TRUE), 2)
  )
kruskal.test(pm25_1yr_com ~ income_cat, data = df)



ggplot(df_no2, aes(x = income_cat, y = no2_1yr_com)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(
    title = "NO2 1 y average, by Income Group",
    x = "Household Income",
    y = "NO2 (ppb)"
  ) +
  report_theme

ggsave("Figures/boxplot_no2.png", width = 5, height = 4, dpi = 300)



ggplot(df_so2, aes(x = income_cat, y = so2_1yr_com)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(
    title = "SO2 1 y average, by Income Group",
    x = "Household Income",
    y = "SO2 (ppb)"
  ) +
  report_theme

ggsave("Figures/boxplot_so2.png", width = 5, height = 4, dpi = 300)



ggplot(df_o3, aes(x = income_cat, y = o3_1yr_com)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(
    title = "Ozone 1 y average, by Income Group",
    x = "Household Income",
    y = "Ozone (ppb)"
  ) +
  report_theme

ggsave("Figures/boxplot_o3.png", width = 5, height = 4, dpi = 300)


ggplot(df_greenness, aes(x = income_cat, y = GRLANYY_26_COM)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(
    title = "Max NDVI within 500m, by Income Group",
    x = "Household Income",
    y = "Greenness (NDVI)"
  ) +
  report_theme

ggsave("Figures/boxplot_ndvi.png", width = 5, height = 4, dpi = 300)


```



## Pollution by income using the potentially better variables 


PM25DALYY_01_COM
SO2OMIYY_COM
O3CHGYY_01_COM


```{r}

#PM25DALYY_01_COM by Income Group 

plot_data <- subset_data

ggplot(data = plot_data, aes(x = income_cat, y = PM25DALYY_01_COM)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(
    title = "PM2.5 1 year average, by Income Group",
    x = "Household Income",
    y = "PM2.5 (satellite, model-based) (µg/m³)"
  ) +
  report_theme
ggsave("Figures/boxplot_PM25DALYY_01_COM.png", width = 5, height = 4, dpi = 300)



plot_data %>% #check that there is significance for PM2.5 
  group_by(income_cat) %>%
  summarise(
    n = n(),
    median_pm25 = median(PM25DALYY_01_COM, na.rm = TRUE),
    mean_pm25 = round(mean(PM25DALYY_01_COM, na.rm = TRUE), 2)
  )
kruskal.test(PM25DALYY_01_COM ~ income_cat, data = df)







ggplot(plot_data, aes(x = income_cat, y = SO2OMIYY_COM)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(
    title = "SO2 3 year annual average at postal code, by Income Group",
    x = "Household Income",
    y = "3 y average SO2 (satellite derived)(ppb)"
  ) +
  report_theme

ggsave("Figures/boxplot_SO2OMIYY_COM.png", width = 5, height = 4, dpi = 300)



ggplot(plot_data, aes(x = income_cat, y = O3CHGYY_01_COM)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(
    title = "O3 annual average ground-level at postal code, by Income Group",
    x = "Household Income",
    y = "Annual average O3 (model-based) (ppb)"
  ) +
  report_theme

ggsave("Figures/boxplot_O3CHGYY_01_COM.png", width = 5, height = 4, dpi = 300)




```



```{r}
air_income_plots <- list()

for (v in air_income_vars) {
  p <- subset_data %>%
    filter(income_cat %in% c("<20K", "20K–50K", "50K–100K", "100K–150K", "150K+")) %>%
    ggplot(aes(x = income_cat, y = .data[[v$col]], fill = income_cat)) +
    geom_boxplot(outlier.alpha = 0.3, width = 0.6) +
    scale_fill_brewer(palette = "Blues") +
    labs(
      title = v$title,
      x = "Household Income",
      y = v$y_lab
    ) +
    report_theme +
    theme(
      plot.title = element_text(size = 13, face = "bold"),
      axis.title = element_text(size = 11),
      axis.text = element_text(size = 9),
      legend.position = "none"
    )
  
  air_income_plots[[v$col]] <- p
}

dashboard_air_income <- wrap_plots(air_income_plots, ncol = 2) &
  theme(
    plot.margin = margin(t = 15, r = 15, b = 15, l = 15),
    plot.background = element_rect(color = "grey60", linewidth = 1, fill = "white"),
    panel.background = element_rect(fill = "white")
  )

dashboard_air_income <- dashboard_air_income +
  plot_annotation(
    title = "Air Pollution by Household Income",
    theme = theme(
      plot.title    = element_text(hjust = 0.5, face = "bold", size = 22),
      plot.subtitle = element_text(hjust = 0.5, size = 11, color = "grey40"),
      plot.caption  = element_text(size = 9, color = "grey40"),
      plot.margin   = margin(t = 20, r = 10, b = 10, l = 10),
      plot.background = element_rect(fill = "grey90", color = NA)
    )
  )

ggsave("Figures/air_pollution_income_dashboard.png", plot = dashboard_air_income, width = 16, height = 13, dpi = 300)
```





























## Condensation by income 

```{r}
#condensation by income

cond_only <- subset_data %>% #keep only those with condensation problems (54 people)
  filter(ENV_HMPRB_CON_MCQ == 1)

cond_counts <- cond_only %>%
  filter(INC_TOT_COM %in% 1:5) %>%
  mutate(income_cat = factor(INC_TOT_COM, levels = 1:5, labels = income_labels)) %>%
  count(income_cat, name = "n") %>%
  mutate(pct = n / sum(n))

cond_counts

ggplot(cond_counts, aes(x = income_cat, y = n)) +
  geom_col(fill = "#2C7FB8", width = 0.6) +
  geom_text(
    aes(label = paste0(n, " (", percent(pct, accuracy = 0.1), ")")),
    vjust = -0.35,
    size = 3.2  # scaled to base_size 11 theme
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Homes with Condensation, by Income Group",
    x = "Household Income",
    y = "Homes with Condensation"
  ) +
  report_theme

ggsave("Figures/barplot_condensation.png", width = 5, height = 4, dpi = 300)





ggplot(df, aes(x = income_cat, y = pm25_1yr_com)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(
    title = "PM2.5 1 y average, by Income Group",
    x = "Household Income",
    y = "PM2.5 (µg/m³)"
  ) +
  report_theme

ggsave("Figures/boxplot_pm25.png", width = 5, height = 4, dpi = 300)

```








## TCC by income
```{r}

#tree canopy by income, boxplots 
p_veg1 <- ggplot(df, aes(x = income_cat, y = grtcc10_01_COM)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(title = "Vegetation Cover (>5m) by Income Group",
       x = "Household Income", y = "% Pixel Covered by Vegetation >5m") +
  report_theme
ggsave("Figures/boxplot_vegetation_01.png", plot = p_veg1, width = 5, height = 4, dpi = 300)


p_veg2 <- ggplot(df, aes(x = income_cat, y = grtcc10_02_COM)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(title = "Vegetation Cover (>5m) by Income, 100m",
       x = "Household Income", y = "% Pixel Covered by Vegetation >5m") +
  report_theme
ggsave("Figures/boxplot_vegetation_02.png", plot = p_veg2, width = 5, height = 4, dpi = 300)


p_veg3 <- ggplot(df, aes(x = income_cat, y = grtcc10_03_COM)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(title = "Vegetation Cover (>5m) by Income, 250m",
       x = "Household Income", y = "% Pixel Covered by Vegetation >5m") +
  report_theme
ggsave("Figures/boxplot_vegetation_03.png", plot = p_veg3, width = 5, height = 4, dpi = 300)


p_veg4 <- ggplot(df, aes(x = income_cat, y = grtcc10_04_COM)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(title = "Vegetation Cover (>5m) by Income, 500m",
       x = "Household Income", y = "% Pixel Covered by Vegetation >5m") +
  report_theme
ggsave("Figures/boxplot_vegetation_04.png", plot = p_veg4, width = 5, height = 4, dpi = 300)


p_veg5 <- ggplot(df, aes(x = income_cat, y = grtcc10_05_COM)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(title = "Vegetation Cover (>5m) by Income, 1km",
       x = "Household Income", y = "% Pixel Covered by Vegetation >5m") +
  report_theme
ggsave("Figures/boxplot_vegetation_05.png", plot = p_veg5, width = 5, height = 4, dpi = 300)





df <- subset_data %>%
  filter(INC_TOT_COM %in% 1:5, !is.na(grtcc10_01_COM)) %>%
  mutate(
    income_cat = factor(INC_TOT_COM, levels = 1:5, labels = income_labels)
  )

ggplot(df, aes(x = income_cat, y = grtcc10_01_COM)) +
  geom_boxplot(
    fill = "#2C7FB8",
    outlier.alpha = 0.5,
    width = 0.6
  ) +
  labs(
    title = "Vegetation Cover (>5m) by Income Group",
    x = "Household Income",
    y = "Estimated % Pixel Covered by Vegetation >5m"
  ) +
  theme_classic(base_size = 26) +
  theme(
    legend.position = "none",
    
    plot.title = element_text(
      hjust = 0.5,
      face  = "bold",
      size  = 38,
      margin = margin(b = 18)
    ),
    
    axis.title.x = element_text(
      face = "bold",
      size = 30,
      margin = margin(t = 16)
    ),
    axis.title.y = element_text(
      face = "bold",
      size = 30,
      margin = margin(r = 16)
    ),
    
    axis.text.x = element_text(size = 26),
    axis.text.y = element_text(size = 26)
  )
```





## Veg greenness fractions by income 
There is no pattern by income group. The values are all clustered between 26–31%.
The GRUMPY variable measures all green vegetation including grass, shrubs, parks — not just tree canopy in residential areas.


```{r}

library(flextable)

df_green_summary <- df %>%
  group_by(income_cat) %>%
  summarise(
    n = n(),
    median_green = median(GRUMPYY_01_COM, na.rm = TRUE),
    mean_green = round(mean(GRUMPYY_01_COM, na.rm = TRUE), 1)
  ) %>%
  rename(
    `Income Group` = income_cat,
    `N` = n,
    `Median Greenness (%)` = median_green,
    `Mean Greenness (%)` = mean_green
  )

ft_green <- flextable(df_green_summary) %>%
  flextable::add_header_lines("Table X. Vegetation Greenness (GRUMPY) by Income Group") %>%
  flextable::bold(part = "header") %>%
  flextable::bold(j = "Income Group") %>%
  flextable::autofit() %>%
  flextable::theme_booktabs() %>%
  flextable::fontsize(size = 10, part = "all") %>%
  flextable::fontsize(i = 1, size = 11, part = "header") %>%
  flextable::bold(i = 1, part = "header") %>%
  flextable::font(fontname = "Times New Roman", part = "all") %>%
  flextable::align(i = 1, align = "center", part = "header") %>%
  flextable::align(align = "left", part = "all") %>%
  flextable::align(j = c("N", "Median Greenness (%)", "Mean Greenness (%)"), align = "center", part = "body")

save_as_image(ft_green, path = "Tables/Table_veg_green_fraction_byincome.png", res = 300)

```




## Road length (nearby) by income 
This was not a good variable to look at. 

```{r}
#median hwy length within 200 m, by income group, just the participants living near a hwy 

p_hwy_dot <- ggplot(df_hwy_summary, aes(x = income_cat, y = median_length)) +
  geom_point(size = 4, color = "#2C7FB8", aes(shape = "Median")) +
  geom_text(aes(label = paste0("n=", n)), vjust = 2, size = 3.5) +
  scale_shape_manual(name = "", values = c("Median" = 16)) +
  scale_y_continuous(limits = c(40, 105)) +
  labs(title = "Median Highway Length within 200m by Income Group",
       subtitle = "Among participants living near a highway only",
       x = "Household Income",
       y = "Median Highway Length (m)") +
  report_theme
ggsave("Figures/highway_median_dot.png", plot = p_hwy_dot, 
       width = 6, height = 4, dpi = 300)


#median road length within 200 m, by income group, just the participants living near a road 

df_local <- df %>% filter(Locrds_length_200m_com > 0)

df_local_summary <- df_local %>%
  group_by(income_cat) %>%
  summarise(
    n = n(),
    median_length = median(Locrds_length_200m_com, na.rm = TRUE)
  )

p_local_dot <- ggplot(df_local_summary, aes(x = income_cat, y = median_length)) +
  geom_point(size = 4, color = "#2C7FB8", aes(shape = "Median")) +
  geom_text(aes(label = paste0("n=", n)), vjust = 2, size = 3.5) +
  scale_shape_manual(name = "", values = c("Median" = 16)) +
  labs(title = "Median Local Road Length within 200m by Income Group",
       subtitle = "Among participants living near a local road only",
       x = "Household Income",
       y = "Median Local Road Length (m)") +
  report_theme

ggsave("Figures/localroad_median_dot.png", plot = p_local_dot,
       width = 6, height = 4, dpi = 300)

```






















































































# Plots for presentation 


```{r}
air_income_plots_pres <- list()

for (v in air_income_vars) {
  p <- subset_data %>%
    filter(income_cat %in% c("<20K", "20K–50K", "50K–100K", "100K–150K", "150K+")) %>%
    ggplot(aes(x = income_cat, y = .data[[v$col]], fill = income_cat)) +
    geom_boxplot(outlier.alpha = 0.3, width = 0.6) +
    scale_fill_brewer(palette = "Blues") +
    labs(
      title = v$title,
      x = "Household Income",
      y = v$y_lab
    ) +
    report_theme +
    report_theme +
theme(
  plot.title      = element_text(size = 22, face = "bold"),
  axis.title.x    = element_text(size = 20, face = "bold", margin = margin(t = 6)),
  axis.title.y    = element_text(size = 20, face = "bold", margin = margin(r = 6)),
  axis.text.x     = element_text(size = 18),
  axis.text.y     = element_text(size = 18),
  legend.position = "none",
  plot.margin      = margin(t = 15, r = 15, b = 15, l = 15),
  plot.background  = element_rect(color = "grey60", linewidth = 1, fill = "white"),
  panel.background = element_rect(fill = "white")
)
  
  
  
  
  air_income_plots_pres[[v$col]] <- p
}

dashboard_air_income_pres <- wrap_plots(air_income_plots_pres, ncol = 2) +
  plot_annotation(
    title = "Air Pollution by Household Income",
    theme = theme(
      plot.title      = element_text(hjust = 0.5, face = "bold", size = 28),
      plot.margin     = margin(t = 20, r = 10, b = 10, l = 10),
      plot.background = element_rect(fill = "grey90", color = NA)
    )
  )


ggsave("Figures/air_pollution_income_dashboard_PRESENTATION.png",
       plot = dashboard_air_income_pres, width = 18, height = 14, dpi = 300)





```






# TO DO LIST  
-imputation - check the diagnostic plots and make sure they are ok 
-rurun models in a separate RMD (keep the other as a sensitivity analysis). reformat tables 2,3,4 to make sure column titles are ok. 
-mediation model 




































# REFERENCE NOTES
Meeting with Keenan june 16, 2026:

Keenan: CLSA_KE 
https://github.com/StringhiniLab/CLSA_KE
TrajectoryAnalysis_ImpBio.R
lines 1015-1033 
income categories become just 1-5. but keep in mind it's not really linear, there are differences between 1-2 that are different than between 4-5. 

model: line 1045 

Trajectory_HealthOutcomes.R
mediation line # 699

use officer package in R 

-report only main findings: https://pmc.ncbi.nlm.nih.gov/articles/PMC10321643/



# OTHER STUFF_DELETE



### Categorical income 

```{r, eval = FALSE}

# Regression: Income predicting biological aging

class(subset_data$AGE_NMBR_COM)#these are integers, ok to leave alone
class(subset_data$CD4nv) 

#stratify by age 
subset_data$age_group <- ifelse(subset_data$AGE_NMBR_COM < 65, "45-64", "65-85")
subset_data_firstgrp  <- subset_data[subset_data$age_group == "45-64", ]
subset_data_secondgrp <- subset_data[subset_data$age_group == "65-85", ]
table(subset_data$INC_TOT_COM) #make sure it is reverse coded 

#GrimAge 
library(broom)
model <- lm(GrimAge_EAA ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data_firstgrp)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/GrimAge_model_results_45to64.csv", row.names = FALSE)

model <- lm(GrimAge_EAA ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data_secondgrp)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/GrimAge_model_results_65to85.csv", row.names = FALSE)

# PhenoAge
model <- lm(PhenoAge_EAA ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data_firstgrp)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/PhenoAge_model_results_45to64.csv", row.names = FALSE)

model <- lm(PhenoAge_EAA ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data_secondgrp)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/PhenoAge_model_results_65to85.csv", row.names = FALSE)

# DunedinPACE
model <- lm(DunedinPACE ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data_firstgrp)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/DunedinPACE_model_results_45to64.csv", row.names = FALSE)

model <- lm(DunedinPACE ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data_secondgrp)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/DunedinPACE_model_results_65to85.csv", row.names = FALSE)

# DNAmIC
model <- lm(DNAmIC ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data_firstgrp)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/DNAmIC_model_results_45to64.csv", row.names = FALSE)

model <- lm(DNAmIC ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data_secondgrp)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/DNAmIC_model_results_65to85.csv", row.names = FALSE)







###########
#make into a nice table 

library(tidyr)

read_model <- function(file, outcome, age_group) {
  read.csv(file) %>%
    mutate(Outcome = outcome, Age_Group = age_group)
}

all_results <- bind_rows(
  read_model("regression_csvs/GrimAge_model_results_45to64.csv",     "GrimAge EAA",  "45-64"),
  read_model("regression_csvs/GrimAge_model_results_65to85.csv",     "GrimAge EAA",  "65-85"),
  read_model("regression_csvs/PhenoAge_model_results_45to64.csv",    "PhenoAge EAA", "45-64"),
  read_model("regression_csvs/PhenoAge_model_results_65to85.csv",    "PhenoAge EAA", "65-85"),
  read_model("regression_csvs/DunedinPACE_model_results_45to64.csv", "DunedinPACE",  "45-64"),
  read_model("regression_csvs/DunedinPACE_model_results_65to85.csv", "DunedinPACE",  "65-85"),
  read_model("regression_csvs/DNAmIC_model_results_45to64.csv",      "DNAmIC",       "45-64"),
  read_model("regression_csvs/DNAmIC_model_results_65to85.csv",      "DNAmIC",       "65-85")
)

table_data <- all_results %>%
  filter(grepl("INC_TOT_COM", term)) %>%
  mutate(
    term = recode(term,
      "INC_TOT_COM2" = "100-150K",
      "INC_TOT_COM3" = "50-100K",
      "INC_TOT_COM4" = "20-50K",
      "INC_TOT_COM5" = "<20K",
      "INC_TOT_COM6" = "Don't know/Refused"
    ),
    Result = ifelse(
      Outcome == "DNAmIC",
      paste0(formatC(estimate, format = "e", digits = 2),
             " (", formatC(conf.low, format = "e", digits = 2),
             ", ", formatC(conf.high, format = "e", digits = 2), ")"),
      ifelse(
        abs(estimate) < 0.01,
        paste0(round(estimate, 4), " (", round(conf.low, 4), ", ", round(conf.high, 4), ")"),
        paste0(round(estimate, 2), " (", round(conf.low, 2), ", ", round(conf.high, 2), ")")
      )
    ),
    sig = p.value < 0.05,
    p.value = ifelse(round(p.value, 3) == 0, "<0.001", as.character(round(p.value, 3))),
    Result = ifelse(sig, paste0("**", Result, "**"), Result),
    Result_with_p = paste0(Result, ", p=", p.value)
  ) %>%
  select(Outcome, Age_Group, term, Result_with_p)

table_wide <- table_data %>%
  pivot_wider(names_from = Outcome, values_from = Result_with_p)

table_45 <- table_wide %>% filter(Age_Group == "45-64") %>% select(-Age_Group)
table_65 <- table_wide %>% filter(Age_Group == "65-85") %>% select(-Age_Group)

table_final <- bind_rows(table_45, table_65)

write.csv(table_final, "all_model_results_summary.csv", row.names = FALSE)

save_kable(
  knitr::kable(table_final,
               col.names = c("Income Group", "GrimAge EAA", "PhenoAge EAA", "DunedinPACE", "DNAmIC"),
               caption = "Table X: Association between household income and epigenetic aging, stratified by age group",
               escape = TRUE) %>%
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) %>%
    pack_rows("45-64 years", 1, nrow(table_45)) %>%
    pack_rows("65-85 years", nrow(table_45) + 1, nrow(table_final)),
  "Figures/income_epigenetic_table.png", zoom = 2
)



```


### Categorical income 
income predicting environment 

```{r}
## do all the regressions first 

levels(subset_data$INC_TOT_COM)

#Income predicting PM2.5 
library(broom)
model <- lm(PM25DALYY_01_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_PM2.5.csv", row.names = FALSE)


#income predicting ozone
model <- lm(O3CHGYY_01_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_ozone.csv", row.names = FALSE)



#Income predicting NO2 
model <- lm(no2_1yr_com ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_NO2.csv", row.names = FALSE)



#Income predicting SO2 
model <- lm(so2_1yr_com ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_SO2.csv", row.names = FALSE)



#Income predicting condensation
subset_data$condensation_binary <- as.numeric(subset_data$ENV_HMPRB_CON_MCQ == "Yes") #first make condensation numeric b/c it is outcome now. 
model <- lm(condensation_binary ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_condensation.csv", row.names = FALSE)




############################
#TCC


# Tree canopy cover at postal code
model <- lm(grtcc10_01_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
               I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
             + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
             data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_TreeCanopyPostalCode.csv", row.names = FALSE)

# Tree canopy cover within 100m
model <- lm(grtcc10_02_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
               I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
             + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
             data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_TreeCanopy100m.csv", row.names = FALSE)

# Tree canopy cover within 250m
model <- lm(grtcc10_03_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
               I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
             + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
             data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_TreeCanopy250m.csv", row.names = FALSE)

# Tree canopy cover within 500m
model <- lm(grtcc10_04_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
               I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
             + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
             data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_TreeCanopy500m.csv", row.names = FALSE)

# Tree canopy cover within 1km
model <- lm(grtcc10_05_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
               I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
             + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
             data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_TreeCanopy1km.csv", row.names = FALSE)




#####################
#% veg greenness 

# Vegetative greenness at postal code
model <- lm(GRUMPYY_01_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
               I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
             + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
             data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_GreennessPostalCode.csv", row.names = FALSE)

# Vegetative greenness mean within 100m
model <- lm(GRUMPYY_02_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
               I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
             + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
             data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_GreennessMean100m.csv", row.names = FALSE)

# Vegetative greenness mean within 250m
model <- lm(GRUMPYY_03_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
               I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
             + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
             data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_GreennessMean250m.csv", row.names = FALSE)

# Vegetative greenness mean within 500m
model <- lm(GRUMPYY_04_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
               I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
             + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
             data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_GreennessMean500m.csv", row.names = FALSE)

# Vegetative greenness mean within 1000m
model <- lm(GRUMPYY_05_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
               I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
             + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
             data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_GreennessMean1000m.csv", row.names = FALSE)

# Vegetative greenness max within 100m
model <- lm(GRUMPYY_06_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
               I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
             + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
             data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_GreennessMax100m.csv", row.names = FALSE)

# Vegetative greenness max within 250m
model <- lm(GRUMPYY_07_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
               I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
             + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
             data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_GreennessMax250m.csv", row.names = FALSE)

# Vegetative greenness max within 500m
model <- lm(GRUMPYY_08_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
               I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
             + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
             data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_GreennessMax500m.csv", row.names = FALSE)

# Vegetative greenness max within 1000m
model <- lm(GRUMPYY_09_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
               I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
             + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
             data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_GreennessMax1000m.csv", row.names = FALSE)


#####################
# NDVI (Max 500m)
model <- lm(GRLANYY_26_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
               I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
             + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
             data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_NDVI.csv", row.names = FALSE)






```




```{r}

#Make nice tables for regressions of income predicting environment 

levels(subset_data$INC_TOT_COM)

income_labels <- c(
  "INC_TOT_COM100\u2013150K"                = "100-150K",
  "INC_TOT_COM50\u2013100K"                 = "50-100K",
  "INC_TOT_COM20\u201350K"                  = "20-50K",
  "INC_TOT_COM<20K"                         = "<20K",
  "INC_TOT_COMDon't know/No answer/Refused" = "Don't know/Refused"
)

read_model <- function(file, outcome) {
  read.csv(file) %>% mutate(Outcome = outcome)
}

format_result <- function(data) {
  data %>%
    filter(grepl("INC_TOT_COM", term)) %>%
    mutate(
      term = recode(term, !!!income_labels),
      Result = ifelse(
        abs(estimate) < 0.01,
        paste0(round(estimate, 4), " (", round(conf.low, 4), ", ", round(conf.high, 4), ")"),
        paste0(round(estimate, 2), " (", round(conf.low, 2), ", ", round(conf.high, 2), ")")
      ),
      sig = p.value < 0.05,
      p.value = ifelse(round(p.value, 3) == 0, "<0.001", as.character(round(p.value, 3))),
      Result = ifelse(sig, paste0("**", Result, "**"), Result),
      Result_with_p = paste0(Result, ", p=", p.value)
    )
}

make_table <- function(data, col_names, caption, filename) {
  save_kable(
    knitr::kable(data, col.names = col_names, caption = caption, escape = TRUE) %>%
      kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE),
    filename, zoom = 2
  )
}

# Air pollution
bind_rows(
  read_model("regression_csvs/Income_PM2.5.csv",  "PM2.5"),
  read_model("regression_csvs/Income_ozone.csv",  "Ozone (O3)"),
  read_model("regression_csvs/Income_NO2.csv",    "NO2"),
  read_model("regression_csvs/Income_SO2.csv",    "SO2")
) %>% format_result() %>%
  select(Outcome, term, Result_with_p) %>%
  pivot_wider(names_from = Outcome, values_from = Result_with_p) %>%
  make_table(c("Income Group", "PM2.5", "Ozone (O3)", "NO2", "SO2"),
             "Table X: Association between household income and air pollution exposures",
             "Figures/income_airpollution_table.png")

# Condensation
read_model("regression_csvs/Income_condensation.csv", "Condensation") %>%
  format_result() %>%
  select(term, Result_with_p) %>%
  make_table(c("Income Group", "Condensation"),
             "Table X: Association between household income and housing condensation",
             "Figures/income_condensation_table.png")

# Tree Canopy Cover
bind_rows(
  read_model("regression_csvs/Income_TreeCanopyPostalCode.csv", "Postal code"),
  read_model("regression_csvs/Income_TreeCanopy100m.csv",       "100m"),
  read_model("regression_csvs/Income_TreeCanopy250m.csv",       "250m"),
  read_model("regression_csvs/Income_TreeCanopy500m.csv",       "500m"),
  read_model("regression_csvs/Income_TreeCanopy1km.csv",        "1km")
) %>% format_result() %>%
  select(Outcome, term, Result_with_p) %>%
  pivot_wider(names_from = Outcome, values_from = Result_with_p) %>%
  make_table(c("Income Group", "Postal code", "100m", "250m", "500m", "1km"),
             "Table X: Association between household income and tree canopy cover",
             "Figures/income_treecanopy_table.png")

# Vegetative Greenness
bind_rows(
  read_model("regression_csvs/Income_GreennessPostalCode.csv", "Postal code"),
  read_model("regression_csvs/Income_GreennessMean100m.csv",   "Mean 100m"),
  read_model("regression_csvs/Income_GreennessMean250m.csv",   "Mean 250m"),
  read_model("regression_csvs/Income_GreennessMean500m.csv",   "Mean 500m"),
  read_model("regression_csvs/Income_GreennessMean1000m.csv",  "Mean 1000m"),
  read_model("regression_csvs/Income_GreennessMax100m.csv",    "Max 100m"),
  read_model("regression_csvs/Income_GreennessMax250m.csv",    "Max 250m"),
  read_model("regression_csvs/Income_GreennessMax500m.csv",    "Max 500m"),
  read_model("regression_csvs/Income_GreennessMax1000m.csv",   "Max 1000m")
) %>% format_result() %>%
  select(Outcome, term, Result_with_p) %>%
  pivot_wider(names_from = Outcome, values_from = Result_with_p) %>%
  make_table(c("Income Group", "Postal code", "Mean 100m", "Mean 250m", "Mean 500m",
               "Mean 1000m", "Max 100m", "Max 250m", "Max 500m", "Max 1000m"),
             "Table X: Association between household income and vegetative greenness",
             "Figures/income_greenness_table.png")

# NDVI
read_model("regression_csvs/Income_NDVI.csv", "NDVI (Max 500m)") %>%
  format_result() %>%
  select(term, Result_with_p) %>%
  make_table(c("Income Group", "NDVI (Max 500m)"),
             "Table X: Association between household income and NDVI",
             "Figures/income_NDVI_table.png")
```






























```{r}


PM25DALYY_01_COM
SO2OMIYY_COM
O3CHGYY_01_COM


#pm2.5
model <- lm(GrimAge_EAA ~ PM25DALYY_01_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "pm25_grimage.csv", row.names = FALSE)

model <- lm(PhenoAge_EAA ~ PM25DALYY_01_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "pm25_phenoage.csv", row.names = FALSE)

model <- lm(DunedinPACE ~ PM25DALYY_01_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "pm25_Duned.csv", row.names = FALSE)

model <- lm(DNAmIC ~ PM25DALYY_01_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "pm25_IC.csv", row.names = FALSE)





#so2_1yr_com
model <- lm(GrimAge_EAA ~ SO2OMIYY_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "pm25_grimage.csv", row.names = FALSE)


model <- lm(PhenoAge_EAA ~ SO2OMIYY_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "pm25_phenoage.csv", row.names = FALSE)


model <- lm(DunedinPACE ~ SO2OMIYY_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "pm25_Duned.csv", row.names = FALSE)


model <- lm(DNAmIC ~ SO2OMIYY_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "pm25_IC.csv", row.names = FALSE)




#ozone
model <- lm(GrimAge_EAA ~ O3CHGYY_01_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "pm25_grimage.csv", row.names = FALSE)


model <- lm(PhenoAge_EAA ~ O3CHGYY_01_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "pm25_phenoage.csv", row.names = FALSE)


model <- lm(DunedinPACE ~ O3CHGYY_01_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "pm25_Duned.csv", row.names = FALSE)


model <- lm(DNAmIC ~ O3CHGYY_01_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "pm25_IC.csv", row.names = FALSE)


```





