---
title: "Visualizations"
output:
  word_document: default
  pdf_document: default
  html_document: default
date: "2026-04-10"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,       # hides code
  message = FALSE,    # suppresses messages
  warning = FALSE     # suppresses warnings
)
```

# 1. Load and clean data 

```{r}

#load data 

load("data/CLSA_methylation.rds")

library(dplyr)
library(ggplot2)
library(tidyverse)
library(flextable)
library (haven)
library(table1)
library(stargazer)
library(htmltools)

library(table1)
library(Hmisc)
library(htmltools)
library(webshot2)
library(tidyverse)


#not going to include these: 
#"ED_HIGH_COM", "WEA_SVNGSVL_MCQ","temp_1yr_com" LGTNLTYY_01_COM "ENV_HMPRB_NOI_MCQ","pm25_5yrs_com",   "PM25DALYY_02_COM", "no2_5yrs_com", 
 # "Locrds_length_200m_com",   "Phwy_length_200m_com", 

#create subset 
subset_data <- CLSA_methylation[, c(
  "entity_id.x", 
  
 #demographics
  "AGE_NMBR_COM", "SEX_ASK_COM", "WGHTS_PROV_COM", "Race7",
  
  ##############
 #Socioeconomic variables
 ##############
 
  "INC_TOT_COM", "WEA_SVNGSVL_MCQ",
  
 
  ##############
 #Aging outcome variables
 ##############
 
  # aging outcomes
  "GrimAge_EAA", "PhenoAge_EAA", "DunedinPACE", "DNAmIC",
  
 
 ##############
 #Environmental variables
 ##############
 
  #pollution 
   "PM25DALYY_01_COM",  "O3CHGYY_01_COM", "no2_1yr_com", "so2_1yr_com",
 
 #condensation 
 "ENV_HMPRB_CON_MCQ",
 
 #NDVI
 "GRLANYY_26_COM", 
  
  ##Tree canopy and %veg 
  "grtcc10_01_COM",   "grtcc10_02_COM",   "grtcc10_03_COM",   "grtcc10_04_COM",  "grtcc10_05_COM",  #tree canopy coverage 
  
  "GRUMPYY_01_COM",  "GRUMPYY_02_COM", "GRUMPYY_03_COM",   "GRUMPYY_04_COM",   "GRUMPYY_05_COM",   "GRUMPYY_06_COM",   "GRUMPYY_07_COM",
  "GRUMPYY_08_COM",   "GRUMPYY_09_COM", #%vegetative greenness 
  
  
  # Lifestyle / behavioral
  "PA2_DSCR2_MCQ", "ALC_FREQ_COM", "ICQ_SMOKE_COM", #PASE scale: PASE Score - Final (in sharepoint dictionary)
  
  #cell types
  "CD4nv", "CD8nv",   "Mono", "Treg", "CD4mem", "CD8mem", "NK", "Bmem", "Bnv"
)]

#check structure of variables 
str(subset_data)

```


## Recode missing values to N/A

```{r}
# Recode NAs
subset_data$no2_1yr_com[subset_data$no2_1yr_com %in% c(-8888, -9999)] <- NA
subset_data$so2_1yr_com[subset_data$so2_1yr_com %in% c(-8888, -9999)] <- NA
subset_data$GRLANYY_26_COM[subset_data$GRLANYY_26_COM %in% c(-8888, -9999)] <- NA
subset_data$ENV_HMPRB_CON_MCQ[subset_data$ENV_HMPRB_CON_MCQ %in% c(-8888, -9999)] <- NA
subset_data$PM25DALYY_01_COM[subset_data$PM25DALYY_01_COM %in% c(-8888, -9999)] <- NA
subset_data$O3CHGYY_01_COM[subset_data$O3CHGYY_01_COM %in% c(-8888, -9999)] <- NA
subset_data$grtcc10_01_COM[subset_data$grtcc10_01_COM %in% c(-8888, -9999)] <- NA
subset_data$grtcc10_02_COM[subset_data$grtcc10_02_COM %in% c(-8888, -9999)] <- NA
subset_data$grtcc10_03_COM[subset_data$grtcc10_03_COM %in% c(-8888, -9999)] <- NA
subset_data$grtcc10_04_COM[subset_data$grtcc10_04_COM %in% c(-8888, -9999)] <- NA
subset_data$grtcc10_05_COM[subset_data$grtcc10_05_COM %in% c(-8888, -9999)] <- NA
subset_data$GRUMPYY_01_COM[subset_data$GRUMPYY_01_COM %in% c(-8888, -9999)] <- NA
subset_data$GRUMPYY_02_COM[subset_data$GRUMPYY_02_COM %in% c(-8888, -9999)] <- NA
subset_data$GRUMPYY_03_COM[subset_data$GRUMPYY_03_COM %in% c(-8888, -9999)] <- NA
subset_data$GRUMPYY_04_COM[subset_data$GRUMPYY_04_COM %in% c(-8888, -9999)] <- NA
subset_data$GRUMPYY_05_COM[subset_data$GRUMPYY_05_COM %in% c(-8888, -9999)] <- NA
subset_data$GRUMPYY_06_COM[subset_data$GRUMPYY_06_COM %in% c(-8888, -9999)] <- NA
subset_data$GRUMPYY_07_COM[subset_data$GRUMPYY_07_COM %in% c(-8888, -9999)] <- NA
subset_data$GRUMPYY_08_COM[subset_data$GRUMPYY_08_COM %in% c(-8888, -9999)] <- NA
subset_data$GRUMPYY_09_COM[subset_data$GRUMPYY_09_COM %in% c(-8888, -9999)] <- NA

# Then check — should all be 0
sapply(subset_data, function(x) sum(x %in% c(-8888, -9999), na.rm = TRUE))

colSums(is.na(subset_data))#check number of NAs

```



## Reverse code Household Income and savings 

```{r}
# BEFORE reversing
cat("=== BEFORE reverse coding ===\n")
table(subset_data$INC_TOT_COM, useNA = "ifany")
# 1 = Less than $20,000; 2 = $20,000–$49,999; 3 = $50,000–$99,999
# 4 = $100,000–$149,999; 5 = $150,000 or more; 8 = Don't know/No answer; 9 = Refused

subset_data$INC_TOT_COM_RECODED <- dplyr::case_when(
  subset_data$INC_TOT_COM == 1 ~ 5,
  subset_data$INC_TOT_COM == 2 ~ 4,
  subset_data$INC_TOT_COM == 3 ~ 3,
  subset_data$INC_TOT_COM == 4 ~ 2,
  subset_data$INC_TOT_COM == 5 ~ 1,
  subset_data$INC_TOT_COM %in% c(8, 9) ~ 6,
  TRUE ~ NA_real_
)

# AFTER reversing
cat("\n=== AFTER reverse coding ===\n")
table(subset_data$INC_TOT_COM_RECODED, useNA = "ifany")
# 1 = $150,000+; 2 = $100,000–$149,999; 3 = $50,000–$99,999
# 4 = $20,000–$49,999; 5 = Less than $20,000; 6 = Don't know/No answer/Refused

subset_data$INC_TOT_COM <- subset_data$INC_TOT_COM_RECODED
cat("\nNAs in INC_TOT_COM after recoding:", sum(is.na(subset_data$INC_TOT_COM)), "\n")
subset_data <- subset_data %>% select(-INC_TOT_COM_RECODED) 

# Create income_cat factor
subset_data <- subset_data %>%
  mutate(
    income_cat = factor(
      INC_TOT_COM,
      levels = c(1, 2, 3, 4, 5, 6),
      labels = c("150K+", "100K–150K", "50K–100K", "20K–50K", "<20K", "Don't know/No answer/Refused")
    )
  )

# Final check
cat("\n=== income_cat factor (final) ===\n")
table(subset_data$income_cat, useNA = "ifany")


```




```{r}

# BEFORE reversing
cat("=== BEFORE reverse coding ===\n")
table(subset_data$WEA_SVNGSVL_MCQ, useNA = "ifany")
# 1 = Less than $50,000; 2 = $50,000–$99,999; 3 = $100,000–<$1 million
# 4 = $1 million or more; 8 = Don't know/No answer; 9 = Refused

subset_data$WEA_SVNGSVL_MCQ_RECODED <- dplyr::case_when(
  subset_data$WEA_SVNGSVL_MCQ == 1 ~ 4,
  subset_data$WEA_SVNGSVL_MCQ == 2 ~ 3,
  subset_data$WEA_SVNGSVL_MCQ == 3 ~ 2,
  subset_data$WEA_SVNGSVL_MCQ == 4 ~ 1,
  subset_data$WEA_SVNGSVL_MCQ %in% c(8, 9) ~ 5,
  TRUE ~ NA_real_
)

# AFTER reversing
cat("\n=== AFTER reverse coding ===\n")
table(subset_data$WEA_SVNGSVL_MCQ_RECODED, useNA = "ifany")
# 1 = $1 million+; 2 = $100,000–<$1 million; 3 = $50,000–$99,999
# 4 = Less than $50,000; 5 = Don't know/No answer/Refused

subset_data$WEA_SVNGSVL_MCQ <- subset_data$WEA_SVNGSVL_MCQ_RECODED

cat("\nNAs in WEA_SVNGSVL_MCQ after recoding:", sum(is.na(subset_data$WEA_SVNGSVL_MCQ)), "\n")
subset_data <- subset_data %>% select(-WEA_SVNGSVL_MCQ_RECODED)

# Create savings_cat factor
subset_data <- subset_data %>%
  mutate(
    savings_cat = factor(
      WEA_SVNGSVL_MCQ,
      levels = c(1, 2, 3, 4, 5),
      labels = c("$1M+", "$100K–<$1M", "$50K–$100K", "<$50K", "Don't know/No answer/Refused")
    )
  )

# Final check
cat("\n=== savings_cat factor (final) ===\n")
table(subset_data$savings_cat, useNA = "ifany")

```




General Code needed for plots 

```{r}
#Custom theme for all figures

report_theme <- theme_classic(base_size = 11) +
  theme(
    legend.position = "none",
    
    plot.title = element_text(
      hjust = 0.5,
      face  = "bold",
      size  = 13,
      margin = margin(b = 8)
    ),
    
    axis.title.x = element_text(
      face = "bold",
      size = 11,
      margin = margin(t = 6)
    ),
    axis.title.y = element_text(
      face = "bold",
      size = 11,
      margin = margin(r = 6)
    ),
    
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10)
  )

```




## Missingness 

```{r}
#install.packages("DataExplorer")
library(DataExplorer)
library(mice)
library(ggplot2)
plot_missing(subset_data)
md.pattern(subset_data)


colSums(is.na(subset_data)) #how many NAs for each variable 

p <- plot_missing(subset_data)
ggsave("Figures/missing_plot.png", plot = p, width = 12, height = 7, dpi = 600)



missing_table <- data.frame(
  Variable = c(
    "AGE_NMBR_COM", "SEX_ASK_COM", "WGHTS_PROV_COM", "Race7",
    "INC_TOT_COM", "WEA_SVNGSVL_MCQ",
    "GrimAge_EAA", "PhenoAge_EAA", "DunedinPACE", "DNAmIC",
    "grtcc10_01_COM", "grtcc10_02_COM", "grtcc10_03_COM", "grtcc10_04_COM", "grtcc10_05_COM",
    "GRUMPYY_01_COM", "GRUMPYY_02_COM", "GRUMPYY_03_COM", "GRUMPYY_04_COM", "GRUMPYY_05_COM",
    "GRUMPYY_06_COM", "GRUMPYY_07_COM", "GRUMPYY_08_COM", "GRUMPYY_09_COM",
    "GRLANYY_26_COM", "ENV_HMPRB_CON_MCQ",
    "PM25DALYY_01_COM", "O3CHGYY_01_COM", "SO2OMIYY_COM",
    "CD4nv", "CD8nv", "Mono", "Treg", "CD4mem", "CD8mem", "NK", "Bmem", "Bnv"
  ),
  Description = c(
    "Age", "Sex", "Province", "Self-reported ethnicity",
    "Household income", "Household savings",
    "GrimAge", "PhenoAge", "DunedinPACE", "DNAmIC",
    "Tree canopy cover at postal code (2010)",
    "Tree canopy cover within 100m (2010)",
    "Tree canopy cover within 250m (2010)",
    "Tree canopy cover within 500m (2010)",
    "Tree canopy cover within 1km (2010)",
    "Vegetative greenness at postal code (annual %)",
    "Vegetative greenness within 100m (mean annual %)",
    "Vegetative greenness within 250m (mean annual %)",
    "Vegetative greenness within 500m (mean annual %)",
    "Vegetative greenness within 1000m (mean annual %)",
    "Vegetative greenness within 100m (max annual %)",
    "Vegetative greenness within 250m (max annual %)",
    "Vegetative greenness within 500m (max annual %)",
    "Vegetative greenness within 1000m (max annual %)",
    "NDVI (vegetative greenness, 500m buffer)", "Housing condensation",
    "PM2.5 concentration (ug/m³, annual avg)", "Ozone (ppb, annual avg)", "SO2 concentration (ppb, 3-yr avg)",
    "Naïve CD4+ T cells", "Naïve CD8+ T cells", "Monocytes", "Regulatory T cells",
    "Memory CD4+ T cells", "Memory CD8+ T cells", "Natural Killer cells", "Memory B cells", "Naïve B cells"
  )
)

vars <- missing_table$Variable
missing_table$N_Missing <- sapply(vars, function(v) sum(is.na(subset_data[[v]])))
missing_table$Pct_Missing <- round(missing_table$N_Missing / nrow(subset_data) * 100, 1)

knitr::kable(missing_table,
             col.names = c("Variable", "Description", "N Missing", "% Missing"),
             caption = "Table X: Missing data by variable")


library(kableExtra)

missing_table_formatted <- knitr::kable(missing_table,
               col.names = c("Variable", "Description", "N Missing", "% Missing"),
               caption = "Table X: Missing data by variable",
               format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

save_kable(missing_table_formatted, "Figures/missing_table.png", zoom = 2)

```





##Variable name info (new enviro variables) 


Tree canopy:
grtcc10_01_COM: Estimated percent of pixel covered by vegetation greater than 5m in height (2010)
grtcc10_02_COM: Average of estimated percent of pixel covered by vegetation greater than 5m in height within 100m (2010)
grtcc10_03_COM: Average of estimated percent of pixel covered by vegetation greater than 5m in height within 250m (2010)
grtcc10_04_COM: Average of estimated percent of pixel covered by vegetation greater than 5m in height within 500m (2010)
grtcc10_05_COM: Average of estimated percent of pixel covered by vegetation greater than 5m in height within 1 km (2010)



## Factor and label variables
```{r}

#factor and label variables 

library(gtsummary)
library(dplyr)
library(flextable)

#factor categorical variables 
subset_data <- subset_data %>%
  mutate(
    SEX_ASK_COM = factor(SEX_ASK_COM,
                         levels = c("M", "F"),
                         labels = c("Male", "Female")),

    WGHTS_PROV_COM = factor(WGHTS_PROV_COM,
                            levels = 1:10,
                            labels = c("Alberta", "British Columbia", "Manitoba",
                                       "New Brunswick", "Newfoundland and Labrador",
                                       "Nova Scotia", "Ontario", "Prince Edward Island",
                                       "Quebec", "Saskatchewan")),

    Race7 = factor(Race7,
                   levels = c("White", "South Asian", "Black", "Asian",
                              "Arab", "Latin American", "Other")),

  INC_TOT_COM = factor(INC_TOT_COM,
                     levels = 1:6,
                     labels = c("150K+", "100–150K", "50–100K",
                                "20–50K", "<20K",
                                "Don't know/No answer/Refused")),

WEA_SVNGSVL_MCQ = factor(WEA_SVNGSVL_MCQ,
                          levels = 1:5,
                          labels = c("1M+", "100K–<1M",
                                     "50–100K", "<50K",
                                     "Don't know/No answer/Refused")),



  ENV_HMPRB_CON_MCQ = factor(ENV_HMPRB_CON_MCQ,# "No" must be the first level so it serves as the reference category in regression;
# coefficients will then represent the effect of having condensation (Yes) vs. not (No)
                            levels = c(0, 1),
                            labels = c("No", "Yes"))
  )

#give other variables names 

var_labels <- list(
  AGE_NMBR_COM ~ "Age (years)",
  SEX_ASK_COM ~ "Sex",
  WGHTS_PROV_COM ~ "Province",
  Race7 ~ "Self-reported ethnicity",
  INC_TOT_COM ~ "Household income",
  WEA_SVNGSVL_MCQ ~ "Household savings",
  GrimAge_EAA ~ "GrimAge EAA",
  PhenoAge_EAA ~ "PhenoAge EAA",
  DunedinPACE ~ "DunedinPACE",
  DNAmIC ~ "DNAmIC",
  PM25DALYY_01_COM ~ "PM2.5, 1-year average (µg/m³)",
  no2_1yr_com ~ "NO₂, 1-year average (µg/m³)",
  so2_1yr_com ~ "SO₂, 1-year average (µg/m³)",
  o3_1yr_com ~ "Ozone, 1-year average (µg/m³)",
  GRLANYY_26_COM ~ "NDVI",
  ENV_HMPRB_CON_MCQ ~ "Housing condensation",
  grtcc10_01_COM ~ "Tree canopy cover(2010)",
  grtcc10_02_COM ~ "Tree canopy cover, mean within 100m (2010)",
  grtcc10_03_COM ~ "Tree canopy cover, mean within 250m (2010)",
  grtcc10_04_COM ~ "Tree canopy cover, mean within 500m (2010)",
  grtcc10_05_COM ~ "Tree canopy cover, mean within 1km (2010)",
  GRUMPYY_01_COM ~ "Vegetative greenness, annual % (postal code)",
  GRUMPYY_02_COM ~ "Vegetative greenness, mean within 100m",
  GRUMPYY_03_COM ~ "Vegetative greenness, mean within 250m",
  GRUMPYY_04_COM ~ "Vegetative greenness, mean within 500m",
  GRUMPYY_05_COM ~ "Vegetative greenness, mean within 1000m",
  GRUMPYY_06_COM ~ "Vegetative greenness, max within 100m",
  GRUMPYY_07_COM ~ "Vegetative greenness, max within 250m",
  GRUMPYY_08_COM ~ "Vegetative greenness, max within 500m",
  GRUMPYY_09_COM ~ "Vegetative greenness, max within 1000m"
)

```



#2. Descriptive stats

## Tables 

```{r}
###Table1
#demographics
table_demographics <- df_table %>%
  select(AGE_NMBR_COM, SEX_ASK_COM, WGHTS_PROV_COM, Race7) %>%
  tbl_summary(
    label = list(
      AGE_NMBR_COM ~ "Age (years)",
      SEX_ASK_COM ~ "Sex",
      WGHTS_PROV_COM ~ "Province",
      Race7 ~ "Self-reported ethnicity"
    ),
    type = list(SEX_ASK_COM ~ "categorical",
                Race7 ~ "categorical"),
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "no"
  ) %>%
  add_n(
    statistic = "{N_miss} ({p_miss}%)",
    col_label = "**Missing**"
  ) %>%
  modify_table_body(
    ~.x %>% relocate(n, .after = stat_0)
  ) %>%
  as_flex_table()

save_as_image(table_demographics, path = "Figures/demographics.png", res = 300)

#income and wealth 
table_SES <- df_table %>%
  select(INC_TOT_COM, WEA_SVNGSVL_MCQ) %>%
  tbl_summary(
    label = list(
      INC_TOT_COM ~ "Household income",
      WEA_SVNGSVL_MCQ ~ "Household savings"
    ),
    type = list(INC_TOT_COM ~ "categorical",
                WEA_SVNGSVL_MCQ ~ "categorical"),
    statistic = list(all_categorical() ~ "{n} ({p}%)"),
    missing = "no"
  ) %>%
  add_n(
    statistic = "{N_miss} ({p_miss}%)",
    col_label = "**Missing**"
  ) %>%
  modify_table_body(
    ~.x %>% relocate(n, .after = stat_0)
  ) %>%
  as_flex_table()

save_as_image(table_SES, path = "Figures/SES.png", res = 300)

#aging outcomes
table_aging <- df_table %>%
  select(GrimAge_EAA, PhenoAge_EAA, DunedinPACE, DNAmIC) %>%
  tbl_summary(
    label = list(
      GrimAge_EAA ~ "GrimAge EAA",
      PhenoAge_EAA ~ "PhenoAge EAA",
      DunedinPACE ~ "DunedinPACE",
      DNAmIC ~ "DNAmIC"
    ),
    statistic = list(all_continuous() ~ "{mean} ({sd})"),
    missing = "no"
  ) %>%
  add_n(
    statistic = "{N_miss} ({p_miss}%)",
    col_label = "**Missing**"
  ) %>%
  modify_table_body(
    ~.x %>% relocate(n, .after = stat_0)
  ) %>%
  as_flex_table()

save_as_image(table_aging, path = "Figures/aging_outcomes.png", res = 300)


#air pollution 
table_pollution <- df_table %>%
  select(PM25DALYY_01_COM, O3CHGYY_01_COM, no2_1yr_com, so2_1yr_com, GRLANYY_26_COM, ENV_HMPRB_CON_MCQ) %>%
  tbl_summary(
    label = list(
      PM25DALYY_01_COM ~ "PM2.5, 1-year average (µg/m³)",
      O3CHGYY_01_COM ~ "Ozone, 1-year average (ppb)",
      no2_1yr_com ~ "NO₂, 1-year average (µg/m³)",
      so2_1yr_com ~ "SO₂, 1-year average (ppb)",
      GRLANYY_26_COM ~ "NDVI",
      ENV_HMPRB_CON_MCQ ~ "Housing condensation"
    ),
    type = list(ENV_HMPRB_CON_MCQ ~ "categorical"),
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "no"
  ) %>%
  add_n(
    statistic = "{N_miss} ({p_miss}%)",
    col_label = "**Missing**"
  ) %>%
  modify_table_body(
    ~.x %>% relocate(n, .after = stat_0)
  ) %>%
  as_flex_table()

save_as_image(table_pollution, path = "Figures/pollution.png", res = 300)



#TCC and %vegetative greenness 
#TCC and %vegetative greenness 
table_greenness <- subset_data %>%
  select(grtcc10_01_COM, grtcc10_02_COM, grtcc10_03_COM, grtcc10_04_COM, grtcc10_05_COM,
         GRUMPYY_01_COM, GRUMPYY_02_COM, GRUMPYY_03_COM, GRUMPYY_04_COM, GRUMPYY_05_COM,
         GRUMPYY_06_COM, GRUMPYY_07_COM, GRUMPYY_08_COM, GRUMPYY_09_COM) %>%
  tbl_summary(
    label = list(
      grtcc10_01_COM ~ "Tree canopy cover (2010)",
      grtcc10_02_COM ~ "Tree canopy cover, mean within 100m (2010)",
      grtcc10_03_COM ~ "Tree canopy cover, mean within 250m (2010)",
      grtcc10_04_COM ~ "Tree canopy cover, mean within 500m (2010)",
      grtcc10_05_COM ~ "Tree canopy cover, mean within 1km (2010)",
      GRUMPYY_01_COM ~ "Vegetative greenness, annual % (postal code)",
      GRUMPYY_02_COM ~ "Vegetative greenness, mean within 100m",
      GRUMPYY_03_COM ~ "Vegetative greenness, mean within 250m",
      GRUMPYY_04_COM ~ "Vegetative greenness, mean within 500m",
      GRUMPYY_05_COM ~ "Vegetative greenness, mean within 1000m",
      GRUMPYY_06_COM ~ "Vegetative greenness, max within 100m",
      GRUMPYY_07_COM ~ "Vegetative greenness, max within 250m",
      GRUMPYY_08_COM ~ "Vegetative greenness, max within 500m",
      GRUMPYY_09_COM ~ "Vegetative greenness, max within 1000m"
    ),
    statistic = list(all_continuous() ~ "{mean} ({sd})"),
    missing = "no"
  ) %>%
  add_n(
    statistic = "{N_miss} ({p_miss}%)",
    col_label = "**Missing**"
  ) %>%
  modify_table_body(
    ~.x %>% relocate(n, .after = stat_0)
  ) %>%
  as_flex_table()

save_as_image(table_greenness, path = "Figures/greenness.png", res = 300)





```



```{r}

###Table1




table_demographics <- subset_data %>%
  select(AGE_NMBR_COM, SEX_ASK_COM, WGHTS_PROV_COM, Race7) %>%
  tbl_summary(
    label = list(
      AGE_NMBR_COM ~ "Age (years)",
      SEX_ASK_COM ~ "Sex",
      WGHTS_PROV_COM ~ "Province",
      Race7 ~ "Self-reported ethnicity"
    ),
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "no"
  ) %>%
    as_flex_table() %>%
  add_header_lines("Table X: Demographic characteristics") %>%
  bold(i = 1, part = "header")

save_as_image(table_demographics, path = "Figures/demographics.png", res = 300)



#income and wealth 
table_SES <- subset_data %>%
  select(INC_TOT_COM, WEA_SVNGSVL_MCQ) %>%
  tbl_summary(
    label = list(
      INC_TOT_COM ~ "Household income",
      WEA_SVNGSVL_MCQ ~ "Household savings"
    ),
    statistic = list(all_categorical() ~ "{n} ({p}%)"),
    missing = "no"
  ) %>%
  as_flex_table() %>%
  add_header_lines("Table X: Socioeconomic characteristics") %>%
  bold(i = 1, part = "header")

save_as_image(table_SES, path = "Figures/SES.png", res = 300)


#aging outcomes:
table_aging <- subset_data %>%
  select(GrimAge_EAA, PhenoAge_EAA, DunedinPACE, DNAmIC) %>%
  tbl_summary(
    label = list(
      GrimAge_EAA ~ "GrimAge EAA",
      PhenoAge_EAA ~ "PhenoAge EAA",
      DunedinPACE ~ "DunedinPACE",
      DNAmIC ~ "DNAmIC"
    ),
    statistic = list(all_continuous() ~ "{mean} ({sd})"),
    missing = "no"
  ) %>%
 as_flex_table() %>%
  add_header_lines("Table X: Epigenetic aging outcomes") %>%
  bold(i = 1, part = "header")

save_as_image(table_aging, path = "Figures/aging_outcomes.png", res = 300)



#air pollution 
table_pollution <- subset_data %>%
  select(PM25DALYY_01_COM, O3CHGYY_01_COM, no2_1yr_com, so2_1yr_com, GRLANYY_26_COM, ENV_HMPRB_CON_MCQ) %>%
  tbl_summary(
    label = list(
      PM25DALYY_01_COM ~ "PM2.5, 1-year average (µg/m³)",
      O3CHGYY_01_COM ~ "Ozone, 1-year average (ppb)",
      no2_1yr_com ~ "NO₂, 1-year average (µg/m³)",
      so2_1yr_com ~ "SO₂, 1-year average (ppb)",
      GRLANYY_26_COM ~ "NDVI",
      ENV_HMPRB_CON_MCQ ~ "Housing condensation"
    ),
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "no"
  ) %>%
  as_flex_table() %>%
  add_header_lines("Table X: Environmental exposures") %>%
  bold(i = 1, part = "header")

save_as_image(table_pollution, path = "Figures/pollution.png", res = 300)

#TCC and %vegetative greenness 
table_greenness <- subset_data %>%
  select(grtcc10_01_COM, grtcc10_02_COM, grtcc10_03_COM, grtcc10_04_COM, grtcc10_05_COM,
         GRUMPYY_01_COM, GRUMPYY_02_COM, GRUMPYY_03_COM, GRUMPYY_04_COM, GRUMPYY_05_COM,
         GRUMPYY_06_COM, GRUMPYY_07_COM, GRUMPYY_08_COM, GRUMPYY_09_COM) %>%
  tbl_summary(
    label = list(
      grtcc10_01_COM ~ "Tree canopy cover at postal code (2010)",
      grtcc10_02_COM ~ "Tree canopy cover, mean within 100m (2010)",
      grtcc10_03_COM ~ "Tree canopy cover, mean within 250m (2010)",
      grtcc10_04_COM ~ "Tree canopy cover, mean within 500m (2010)",
      grtcc10_05_COM ~ "Tree canopy cover, mean within 1km (2010)",
      GRUMPYY_01_COM ~ "Vegetative greenness, annual % (postal code)",
      GRUMPYY_02_COM ~ "Vegetative greenness, mean within 100m",
      GRUMPYY_03_COM ~ "Vegetative greenness, mean within 250m",
      GRUMPYY_04_COM ~ "Vegetative greenness, mean within 500m",
      GRUMPYY_05_COM ~ "Vegetative greenness, mean within 1000m",
      GRUMPYY_06_COM ~ "Vegetative greenness, max within 100m",
      GRUMPYY_07_COM ~ "Vegetative greenness, max within 250m",
      GRUMPYY_08_COM ~ "Vegetative greenness, max within 500m",
      GRUMPYY_09_COM ~ "Vegetative greenness, max within 1000m"
    ),
    statistic = list(all_continuous() ~ "{mean} ({sd})"),
    missing = "no"
  ) %>%
as_flex_table() %>%
  add_header_lines("Table X: TCC and %vegetative greenness") %>%
  bold(i = 1, part = "header")

save_as_image(table_greenness, path = "Figures/greenness.png", res = 300)

```






## Plots

### demographics

```{r}

subset_data <- readRDS("subset_data.rds")

# age and sex  ------------------------------------------------------

#pie chart for sex
df_sex <- subset_data %>%
  filter(!is.na(SEX_ASK_COM)) %>%
  count(SEX_ASK_COM) %>%
  mutate(
    percent = n / sum(n),
    label = paste0(SEX_ASK_COM, " (", round(percent * 100, 1), "%)")
  )

ggplot(df_sex, aes(x = 1, y = n, fill = SEX_ASK_COM)) +
  geom_col(width = 1, color = "white", linewidth = 1) +
  coord_polar(theta = "y") +
  geom_text(
    aes(label = label),
    position = position_stack(vjust = 0.5),
    size = 10,              # big, slide-friendly
    fontface = "plain"
  ) +
  scale_fill_manual(values = c("F" = "#1f77b4", "M" = "#ff7f0e")) +
  labs(title = "Sex Distribution", x = NULL, y = NULL, fill = NULL) +
  theme_void() +
  theme(
    # remove any remaining axis/radial artifacts
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    
    plot.title = element_text(
      hjust = 0.5,
      face  = "bold",
      size  = 38,
      margin = margin(b = 18)
    ),
    
    legend.position = "right",
    legend.text = element_text(size = 26),
    legend.key.size = grid::unit(1.2, "lines")
  )


#age boxplot 
subset_data %>%
  filter(!is.na(AGE_NMBR_COM)) %>%
  ggplot(aes(x = "", y = AGE_NMBR_COM)) +
  geom_boxplot(
    width = 0.25,
    fill = "white",
    color = "#2C7FB8",
    linewidth = 1.2
  ) +
  scale_y_continuous(
    breaks = seq(
      floor(min(subset_data$AGE_NMBR_COM, na.rm = TRUE) / 5) * 5,
      ceiling(max(subset_data$AGE_NMBR_COM, na.rm = TRUE) / 5) * 5,
      by = 5
    )
  ) +
  labs(
    title = "Age Distribution",
    x = NULL,
    y = "Age (years)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(
      hjust  = 0.5,
      face   = "bold",
      size   = 38,
      margin = margin(b = 18)
    ),
    axis.title.y = element_text(
      face   = "bold",
      size   = 30,
      margin = margin(r = 16)
    ),
    axis.title.x = element_blank(),
    axis.text.y  = element_text(size = 26),
    axis.text.x  = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_line(linetype = "dotted"),
    panel.grid.major.y = element_line(linetype = "dashed"),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background  = element_rect(fill = "white", color = NA)
  )


ggsave("age_distribution.png", width = 12, height = 7, dpi = 600)

# categories of cultural background:"White", "South Asian", "Black", "Asian","Arab", "Latin American", "Other"
subset_data %>%
  count(Race7) %>%
  mutate(
    percent = round(100 * n / sum(n), 1)
  )

  
```


```{r}

# ethnicity/self reported background barplot 
p_ethnicity <- subset_data %>%
  filter(!is.na(Race7)) %>%
  count(Race7, name = "n") %>%
  arrange(n) %>%
  ggplot(aes(x = n, y = Race7)) +
  geom_col(fill = "#2C7FB8", width = 0.7) +
  labs(
    title = "Self-Reported Cultural Background",
    x = "Number of Participants",
    y = NULL
  ) +
  report_theme

ggsave("Figures/ethnicity_barplot.png", plot = p_ethnicity, width = 5, height = 4, dpi = 300)
```





### Socioeconomic variables: Household Income and Wealth 

```{r}

#summary table of household income
#after reversing: 1 = $150,000 or more ; 2 = $100,000–$149,999 ; 3 = $50,000–$99,999; 4 = $20,000–$49,999; 5 = Less than $20,000; 6=  Don't know/No answer/Refused

subset_data %>%
  count(INC_TOT_COM) %>%
  mutate(
    percent = round(100 * n / sum(n), 1)
  )


# household income barplot 
p_income <- subset_data %>%
  count(income_cat, name = "n") %>%
  ggplot(aes(x = n, y = income_cat)) +
  geom_col(fill = "#2C7FB8", width = 0.7) +
  labs(
    title = "Household Income (Past 12 Months)",
    x = "Number of Participants",
    y = NULL
  ) +
  report_theme

ggsave("Figures/household_income_barplot.png", plot = p_income, width = 5, height = 4, dpi = 300)




#savings categories after reverse code are: 1=1 million or more. 2=100k to 1 million. 3=50-100k. 4=<50K. 5=don't know/no answer/refused 
subset_data %>%
  count(WEA_SVNGSVL_MCQ) %>%
  mutate(
    percent = round(100 * n / sum(n), 1)
  )

# savings/wealth barplot 
p_wealth <- subset_data %>%
  mutate(
    wealth_cat = factor(
      WEA_SVNGSVL_MCQ,
      levels = c(1, 2, 3, 4, 8, 9),
      labels = c(
        "<50K",
        "50K–100K",
        "100K to <$1M",
        "$1M+",
        "Don't know/No answer",
        "Refused"
      )
    )
  ) %>%
  count(wealth_cat, name = "n") %>%
  arrange(n) %>%
  ggplot(aes(x = n, y = wealth_cat)) +
  geom_col(fill = "#2C7FB8", width = 0.7) +
  labs(
    title = "Total Savings and Investments",
    x = "Number of Participants",
    y = NULL
  ) +
  report_theme

ggsave("Figures/savings_barplot.png", plot = p_wealth, width = 5, height = 4, dpi = 300)


```



# 3.0 Environmental Variables 

## Air pollution 


### Compare PM25DALE_AYY_01 and pm25_1yr_com etc. 

PM25DALYY_01_COM
SO2OMIYY_COM
O3CHGYY_01_COM

```{r}
library(ggplot2)
library(tidyr)
library(dplyr)


# ── PM2.5 ──────────────────────────────────────────────────────────────────────
df_long <- subset_data %>%
  select(PM25DALYY_01_COM, pm25_1yr_com) %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") %>%
  filter(!is.na(value))

p_pm25_density <- ggplot(df_long, aes(x = value, fill = variable, color = variable)) +
  geom_density(alpha = 0.4) +
  labs(title = "Distribution Comparison: PM25DALYY_01_COM vs pm25_1yr_com",
       x = "PM2.5 Value", y = "Density") +
  theme_minimal()
ggsave("Figures/pm25_comparison_density.png", plot = p_pm25_density, width = 5, height = 4, dpi = 300)

p_pm25_boxplot <- ggplot(df_long, aes(x = variable, y = value, fill = variable)) +
  geom_boxplot(alpha = 0.6, outlier.shape = 21) +
  labs(title = "Boxplot Comparison: PM25DALYY_01_COM vs pm25_1yr_com",
       x = "", y = "PM2.5 Value") +
  theme_minimal()
ggsave("Figures/pm25_comparison_boxplot.png", plot = p_pm25_boxplot, width = 5, height = 4, dpi = 300)

summary(subset_data[, c("PM25DALYY_01_COM", "pm25_1yr_com")])

# ── SO2 ────────────────────────────────────────────────────────────────────────
df_long <- subset_data %>%
  select(so2_1yr_com, SO2OMIYY_COM) %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") %>%
  filter(!is.na(value))

p_so2_density <- ggplot(df_long, aes(x = value, fill = variable, color = variable)) +
  geom_density(alpha = 0.4) +
  labs(title = "Distribution Comparison: so2_1yr_com vs SO2OMIYY_COM",
       x = "SO2 Value", y = "Density") +
  theme_minimal()
ggsave("Figures/so2_comparison_density.png", plot = p_so2_density, width = 5, height = 4, dpi = 300)

p_so2_boxplot <- ggplot(df_long, aes(x = variable, y = value, fill = variable)) +
  geom_boxplot(alpha = 0.6, outlier.shape = 21) +
  labs(title = "Boxplot Comparison: so2_1yr_com vs SO2OMIYY_COM",
       x = "", y = "SO2 Value") +
  theme_minimal()
ggsave("Figures/so2_comparison_boxplot.png", plot = p_so2_boxplot, width = 5, height = 4, dpi = 300)

summary(subset_data[, c("so2_1yr_com", "SO2OMIYY_COM")])

# ── Ozone ──────────────────────────────────────────────────────────────────────
df_long <- subset_data %>%
  select(O3CHGYY_01_COM, o3_1yr_com) %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") %>%
  filter(!is.na(value))

p_o3_density <- ggplot(df_long, aes(x = value, fill = variable, color = variable)) +
  geom_density(alpha = 0.4) +
  labs(title = "Distribution Comparison: O3CHGYY_01_COM vs o3_1yr_com",
       x = "O3 Value", y = "Density") +
  theme_minimal()
ggsave("Figures/o3_comparison_density.png", plot = p_o3_density, width = 5, height = 4, dpi = 300)

p_o3_boxplot <- ggplot(df_long, aes(x = variable, y = value, fill = variable)) +
  geom_boxplot(alpha = 0.6, outlier.shape = 21) +
  labs(title = "Boxplot Comparison: O3CHGYY_01_COM vs o3_1yr_com",
       x = "", y = "O3 Value") +
  theme_minimal()
ggsave("Figures/o3_comparison_boxplot.png", plot = p_o3_boxplot, width = 5, height = 4, dpi = 300)

summary(subset_data[, c("O3CHGYY_01_COM", "o3_1yr_com")])

```



```{r}
ft <- flextable(summary_table) %>%
  add_header_lines("Table X. Air Quality Exposures") %>%
  bold(part = "header") %>%
  bold(j = "Variable") %>%
  autofit() %>%
  theme_booktabs() %>%
  fontsize(size = 10, part = "all") %>%
  fontsize(i = 1, size = 11, part = "header") %>%  # title slightly bigger
  bold(i = 1, part = "header") %>%                 # title bold
  font(fontname = "Times New Roman", part = "all") %>%
  align(i = 1, align = "center", part = "header") %>%  # title centered
  align(align = "left", part = "all") %>%
  align(j = c("Mean (SD)", "Median [Min, Max]", "Missing"), align = "center", part = "body")

save_as_image(ft, path = "Tables/Table_airquality.png", res = 300)
```

```{r}
#air pollution dashboard 
air_vars <- list(
  list(col = "pm25_1yr_com",  title = "PM2.5 (µg/m³)", file = "pm25",  binwidth = 0.5, limits = c(3, 11),  breaks = seq(3, 11, by = 1)),
  list(col = "no2_1yr_com",   title = "NO2 (ppb)",      file = "no2",   binwidth = 1,   limits = c(3, 16),  breaks = seq(3, 16, by = 2)),
  list(col = "so2_1yr_com",   title = "SO2 (ppb)",      file = "so2",   binwidth = 0.1, limits = c(0, 3.2), breaks = seq(0, 3.2, by = 0.5)),
  list(col = "o3_1yr_com",    title = "Ozone (ppb)",    file = "ozone", binwidth = 1,   limits = c(15, 33), breaks = seq(15, 33, by = 2))
)

air_plots <- list()

for (v in air_vars) {
  p <- ggplot(subset_data, aes(x = .data[[v$col]])) +
    geom_histogram(binwidth = v$binwidth, fill = "steelblue", color = "white") +
    scale_x_continuous(breaks = v$breaks, limits = v$limits) +
    labs(title = v$title, x = v$title, y = "Count") +
    report_theme +
    theme(plot.title = element_text(size = 9))
  
  air_plots[[v$col]] <- p
  ggsave(paste0("Figures/", v$file, "_histogram.png"), plot = p, width = 5, height = 4, dpi = 300)
}

dashboard_air <- wrap_plots(air_plots, ncol = 2) +
  plot_annotation(
    title = "Air Pollution Distributions",
    theme = theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 14))
  )

ggsave("Figures/air_pollution_dashboard.png", plot = dashboard_air, width = 10, height = 8, dpi = 300)

```













## Condensation/noise

```{r}
## 3.3 Condensation and noise 
vars_housing <- list(
  list(col = "ENV_HMPRB_CON_MCQ", label = "Condensation"),
  list(col = "ENV_HMPRB_NOI_MCQ", label = "Noise")
)

summary_table_housing <- lapply(vars_housing, function(v) {
  x <- subset_data[[v$col]]
  tibble(
    Variable  = v$label,
    `Yes n (%)` = paste0(sum(x == 1, na.rm=TRUE), " (", round(mean(x == 1, na.rm=TRUE)*100, 1), "%)"),
    `No n (%)`  = paste0(sum(x == 0, na.rm=TRUE), " (", round(mean(x == 0, na.rm=TRUE)*100, 1), "%)"),
    `Missing`   = paste0(sum(is.na(x)), " (", round(mean(is.na(x))*100, 1), "%)")
  )
}) %>% bind_rows()

ft_housing <- flextable(summary_table_housing) %>%
  add_header_lines("Table X. Problems with Housing") %>%
  bold(part = "header") %>%
  bold(j = "Variable") %>%
  autofit() %>%
  theme_booktabs() %>%
  fontsize(size = 10, part = "all") %>%
  fontsize(i = 1, size = 11, part = "header") %>%
  bold(i = 1, part = "header") %>%
  font(fontname = "Times New Roman", part = "all") %>%
  align(i = 1, align = "center", part = "header") %>%
  align(align = "left", part = "all") %>%
  align(j = c("Yes n (%)", "No n (%)", "Missing"), align = "center", part = "body")

save_as_image(ft_housing, path = "Tables/Table_housing.png", res = 300)
```


```{r}

#Problems with condensation
condensation_bar <- ggplot(subset_data, aes(x = factor(ENV_HMPRB_CON_MCQ))) +
  geom_bar(fill = "steelblue", color = "white") +
  geom_text(
    aes(label = paste0(round(after_stat(count/sum(count)*100), 1), "%")),
    stat = "count",
    vjust = -0.5,
    size = 4
  ) +
  scale_x_discrete(labels = c("0" = "No", "1" = "Yes")) +
  labs(
    title = "Problems with Condensation in Home",
    x = "Condensation",
    y = "Count"
  ) +
  report_theme +
  theme(plot.margin = margin(t = 20, r = 10, b = 10, l = 10)) +
  coord_cartesian(clip = "off")

ggsave("Figures/condensation_bar.png", plot = condensation_bar, width = 5, height = 4, dpi = 300)




#Problems with noise

noise_bar <- ggplot(subset_data, aes(x = factor(ENV_HMPRB_NOI_MCQ))) +
  geom_bar(fill = "steelblue", color = "white") +
  geom_text(
    aes(label = paste0(round(after_stat(count/sum(count)*100), 1), "%")),
    stat = "count",
    vjust = -0.5,
    size = 4
  ) +
  scale_x_discrete(labels = c("0" = "No", "1" = "Yes")) +
  labs(
    title = "Problems with Noise",
    x = "Noise",
    y = "Count"
  ) +
  report_theme +
  theme(plot.margin = margin(t = 20, r = 10, b = 10, l = 10)) +
  coord_cartesian(clip = "off")

ggsave("Figures/noise_bar.png", plot = noise_bar, width = 5, height = 4, dpi = 300)



```












## Tree Canopy Coverage (TCC)
```{r}
# Table 1 - Tree Canopy Coverage
vars_canopy <- list(
  list(col = "grtcc10_01_COM", label = "Mean % of pixel covered by vegetation >5m"),
  list(col = "grtcc10_02_COM", label = "Mean % of pixel covered by vegetation >5m, within 100m"),
  list(col = "grtcc10_03_COM", label = "Mean % of pixel covered by vegetation >5m, within 250m"),
  list(col = "grtcc10_04_COM", label = "Mean % of pixel covered by vegetation >5m, within 500m"),
  list(col = "grtcc10_05_COM", label = "Mean % of pixel covered by vegetation >5m, within 1km")
)

summary_table_canopy <- lapply(vars_canopy, function(v) {
  x <- subset_data[[v$col]]
  tibble(
    Variable            = v$label,
    `Mean (SD)`         = paste0(round(mean(x, na.rm=TRUE), 2), " (", round(sd(x, na.rm=TRUE), 2), ")"),
    `Median [Min, Max]` = paste0(round(median(x, na.rm=TRUE), 2), " [", round(min(x, na.rm=TRUE), 2), ", ", round(max(x, na.rm=TRUE), 2), "]"),
    `Missing`           = paste0(sum(is.na(x)), " (", round(mean(is.na(x))*100, 1), "%)")
  )
}) %>% bind_rows()

ft_canopy <- flextable(summary_table_canopy) %>%
  add_header_lines("Table X. Tree Canopy Coverage") %>%
  bold(part = "header") %>%
  bold(j = "Variable") %>%
  autofit() %>%
  theme_booktabs() %>%
  fontsize(size = 10, part = "all") %>%
  fontsize(i = 1, size = 11, part = "header") %>%
  flextable::bold(i = 1, part = "header") %>%
  font(fontname = "Times New Roman", part = "all") %>%
  align(i = 1, align = "center", part = "header") %>%
  align(align = "left", part = "all") %>%
  align(j = c("Mean (SD)", "Median [Min, Max]", "Missing"), align = "center", part = "body")

save_as_image(ft_canopy, path = "Tables/Table_treecanopy.png", res = 300)

```



```{r}
# Tree canopy histogram dashboard 
canopy_vars <- list(
  list(col = "grtcc10_01_COM", title = "Estimated % Pixel Covered by Vegetation >5m",           file = "canopy_01"),
  list(col = "grtcc10_02_COM", title = "Mean % Pixel Covered by Vegetation >5m within 100m",    file = "canopy_02"),
  list(col = "grtcc10_03_COM", title = "Mean % Pixel Covered by Vegetation >5m within 250m",    file = "canopy_03"),
  list(col = "grtcc10_04_COM", title = "Mean % Pixel Covered by Vegetation >5m within 500m",    file = "canopy_04"),
  list(col = "grtcc10_05_COM", title = "Mean % Pixel Covered by Vegetation >5m within 1km",     file = "canopy_05")
)

for (v in canopy_vars) {# Save individual histograms
  p <- ggplot(subset_data, aes(x = .data[[v$col]])) +
    geom_histogram(binwidth = 5, fill = "forestgreen", color = "white") +
    scale_x_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60, 70, 80)) +
    labs(title = v$title, x = "Tree Canopy Cover (%)", y = "Count") +
    report_theme
  ggsave(paste0("Figures/", v$file, "_histogram.png"), plot = p, width = 5, height = 4, dpi = 300)
}

canopy_plots <- list()# Dashboard

for (v in canopy_vars) {
  p <- ggplot(subset_data, aes(x = .data[[v$col]])) +
    geom_histogram(binwidth = 5, fill = "forestgreen", color = "white") +
    scale_x_continuous(breaks = c(0, 20, 40, 60, 80)) +
    labs(title = v$title, x = "Tree Canopy Cover (%)", y = "Count") +
    report_theme +
    theme(plot.title = element_text(size = 8))
  
  canopy_plots[[v$col]] <- p
}

dashboard_canopy <- wrap_plots(canopy_plots, ncol = 3) +# 5 plots — arrange in 2 rows (3 top, 2 bottom)
  plot_annotation(
    title = "Tree Canopy Coverage Distributions",
    theme = theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 14))
  )

ggsave("Figures/tree_canopy_dashboard.png", plot = dashboard_canopy, width = 15, height = 9, dpi = 300)


```




## Percent vegetation 

```{r}
# Table 2 - Vegetative Greenness
vars_greenness <- list(
  list(col = "GRUMPYY_01_COM", label = "Mean of annual % vegetative greenness, At postal code"),
  list(col = "GRUMPYY_02_COM", label = "Mean of annual % vegetative greenness, within 100m"),
  list(col = "GRUMPYY_03_COM", label = "Mean of annual % vegetative greenness, within 250m"),
  list(col = "GRUMPYY_04_COM", label = "Mean of annual % vegetative greenness, within 500m"),
  list(col = "GRUMPYY_05_COM", label = "Mean of annual % vegetative greenness, within 1km"),
  list(col = "GRUMPYY_06_COM", label = "Max of annual % vegetative greenness, within 100m"),
  list(col = "GRUMPYY_07_COM", label = "Max of annual % vegetative greenness, within 250m"),
  list(col = "GRUMPYY_08_COM", label = "Max of annual % vegetative greenness, within 500m"),
  list(col = "GRUMPYY_09_COM", label = "Max of annual % vegetative greenness, within 1km")
)

summary_table_greenness <- lapply(vars_greenness, function(v) {
  x <- subset_data[[v$col]]
  tibble(
    Variable            = v$label,
    `Mean (SD)`         = paste0(round(mean(x, na.rm=TRUE), 2), " (", round(sd(x, na.rm=TRUE), 2), ")"),
    `Median [Min, Max]` = paste0(round(median(x, na.rm=TRUE), 2), " [", round(min(x, na.rm=TRUE), 2), ", ", round(max(x, na.rm=TRUE), 2), "]"),
    `Missing`           = paste0(sum(is.na(x)), " (", round(mean(is.na(x))*100, 1), "%)")
  )
}) %>% bind_rows()

ft_greenness <- flextable(summary_table_greenness) %>%
  add_header_lines("Table X. Percent Vegetative Greenness") %>%
  bold(part = "header") %>%
  bold(j = "Variable") %>%
  autofit() %>%
  theme_booktabs() %>%
  fontsize(size = 10, part = "all") %>%
  fontsize(i = 1, size = 11, part = "header") %>%
  bold(i = 1, part = "header") %>%
  font(fontname = "Times New Roman", part = "all") %>%
  align(i = 1, align = "center", part = "header") %>%
  align(align = "left", part = "all") %>%
  align(j = c("Mean (SD)", "Median [Min, Max]", "Missing"), align = "center", part = "body")

save_as_image(ft_greenness, path = "Tables/Table_percentveg.png", res = 300)
```

```{r}

#Percent veg greenness histograms (9 of them)

greenness_vars <- list(
  list(col = "GRUMPYY_01_COM", title = "Annual % Vegetative Greenness at Postal Code",        file = "greenness_01"),
  list(col = "GRUMPYY_02_COM", title = "Mean Annual % Vegetative Greenness within 100m",       file = "greenness_02"),
  list(col = "GRUMPYY_03_COM", title = "Mean Annual % Vegetative Greenness within 250m",       file = "greenness_03"),
  list(col = "GRUMPYY_04_COM", title = "Mean Annual % Vegetative Greenness within 500m",       file = "greenness_04"),
  list(col = "GRUMPYY_05_COM", title = "Mean Annual % Vegetative Greenness within 1000m",      file = "greenness_05"),
  list(col = "GRUMPYY_06_COM", title = "Max Annual % Vegetative Greenness within 100m",        file = "greenness_06"),
  list(col = "GRUMPYY_07_COM", title = "Max Annual % Vegetative Greenness within 250m",        file = "greenness_07"),
  list(col = "GRUMPYY_08_COM", title = "Max Annual % Vegetative Greenness within 500m",        file = "greenness_08"),
  list(col = "GRUMPYY_09_COM", title = "Max Annual % Vegetative Greenness within 1000m",       file = "greenness_09")
)

for (v in greenness_vars) {
  p <- ggplot(subset_data, aes(x = .data[[v$col]])) +
    geom_histogram(binwidth = 5, fill = "forestgreen", color = "white") +
    scale_x_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90)) +
    labs(title = v$title, x = "Vegetative Greenness (%)", y = "Count") +
    report_theme
  ggsave(paste0("Figures/", v$file, "_histogram.png"), plot = p, width = 5, height = 4, dpi = 300)
}

library(patchwork)

# Use your loop to store plots in a list instead of saving individually
greenness_plots <- list()

for (v in greenness_vars) {
  p <- ggplot(subset_data, aes(x = .data[[v$col]])) +
    geom_histogram(binwidth = 5, fill = "forestgreen", color = "white") +
    scale_x_continuous(breaks = c(0, 20, 40, 60, 80, 100)) +
    labs(title = v$title, x = "Greenness (%)", y = "Count") +
    report_theme +
    theme(plot.title = element_text(size = 8))  # smaller title to fit
  
  greenness_plots[[v$col]] <- p
}

# Arrange in a 3x3 grid
dashboard <- wrap_plots(greenness_plots, ncol = 3) +
  plot_annotation(
    title = "Vegetative Greenness Distributions",
    theme = theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 14))
  )

ggsave("Figures/greenness_dashboard.png", plot = dashboard, width = 15, height = 12, dpi = 300)




ndvi_500m <- ggplot(subset_data, aes(x = GRLANYY_26_COM)) +
  geom_histogram(binwidth = 5, fill = "forestgreen", color = "white") +
  scale_x_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100)) +
  labs(
    title = "Maximum Annual NDVI within 500m",
    x = "NDVI (%)",
    y = "Count"
  ) +
  report_theme

ggsave("Figures/ndvi_500m_histogram.png", plot = ndvi_500m, width = 5, height = 4, dpi = 300)
GRLANYY_26_COM
```



## NDVI 


```{r}
vars_ndvi <- list(
  list(col = "GRLANYY_26_COM", label = "Maximum annual mean NVDI within 250m")
)

summary_table_ndvi <- lapply(vars_ndvi, function(v) {
  x <- subset_data[[v$col]]
  tibble(
    Variable            = v$label,
    `Mean (SD)`         = paste0(round(mean(x, na.rm=TRUE), 2), " (", round(sd(x, na.rm=TRUE), 2), ")"),
    `Median [Min, Max]` = paste0(round(median(x, na.rm=TRUE), 2), " [", round(min(x, na.rm=TRUE), 2), ", ", round(max(x, na.rm=TRUE), 2), "]"),
    `Missing`           = paste0(sum(is.na(x)), " (", round(mean(is.na(x))*100, 1), "%)")
  )
}) %>% bind_rows()

ft_ndvi <- flextable(summary_table_ndvi) %>%
  add_header_lines("Table X. Normalized Difference Vegetation Index (NDVI)") %>%
  bold(part = "header") %>%
  bold(j = "Variable") %>%
  autofit() %>%
  theme_booktabs() %>%
  fontsize(size = 10, part = "all") %>%
  fontsize(i = 1, size = 11, part = "header") %>%
  bold(i = 1, part = "header") %>%
  font(fontname = "Times New Roman", part = "all") %>%
  align(i = 1, align = "center", part = "header") %>%
  align(align = "left", part = "all") %>%
  align(j = c("Mean (SD)", "Median [Min, Max]", "Missing"), align = "center", part = "body")

save_as_image(ft_ndvi, path = "Tables/Table_ndvi.png", res = 300)
```


```{r}
ndvi_500m <- ggplot(subset_data, aes(x = GRLANYY_26_COM)) +
  geom_histogram(binwidth = 0.02, fill = "forestgreen", color = "white") +
  scale_x_continuous(
    breaks = seq(0.60, 1.00, by = 0.05),
    limits = c(0.60, 1.00)
  ) +
  labs(
    title = "Maximum Annual NDVI within 500m",
    x = "NDVI",
    y = "Count"
  ) +
  report_theme

ggsave("Figures/ndvi_500m_histogram.png", plot = ndvi_500m, width = 5, height = 4, dpi = 300)

```









# 5.0 Environmental variables by income (Pollution, condensation) 

## Pollution by income 
```{r}

#pollution boxplots by income group 


plot_data <- subset_data
ggplot(data = plot_data, aes(x = income_cat, y = pm25_1yr_com)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(
    title = "PM2.5 1 y average, by Income Group",
    x = "Household Income",
    y = "PM2.5 (µg/m³)"
  ) +
  report_theme

ggsave("Figures/boxplot_pm25.png", width = 5, height = 4, dpi = 300)



df %>% #check that there is significance for PM2.5 
  group_by(income_cat) %>%
  summarise(
    n = n(),
    median_pm25 = median(pm25_1yr_com, na.rm = TRUE),
    mean_pm25 = round(mean(pm25_1yr_com, na.rm = TRUE), 2)
  )
kruskal.test(pm25_1yr_com ~ income_cat, data = df)



ggplot(df_no2, aes(x = income_cat, y = no2_1yr_com)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(
    title = "NO2 1 y average, by Income Group",
    x = "Household Income",
    y = "NO2 (ppb)"
  ) +
  report_theme

ggsave("Figures/boxplot_no2.png", width = 5, height = 4, dpi = 300)



ggplot(df_so2, aes(x = income_cat, y = so2_1yr_com)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(
    title = "SO2 1 y average, by Income Group",
    x = "Household Income",
    y = "SO2 (ppb)"
  ) +
  report_theme

ggsave("Figures/boxplot_so2.png", width = 5, height = 4, dpi = 300)



ggplot(df_o3, aes(x = income_cat, y = o3_1yr_com)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(
    title = "Ozone 1 y average, by Income Group",
    x = "Household Income",
    y = "Ozone (ppb)"
  ) +
  report_theme

ggsave("Figures/boxplot_o3.png", width = 5, height = 4, dpi = 300)


ggplot(df_greenness, aes(x = income_cat, y = GRLANYY_26_COM)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(
    title = "Max NDVI within 500m, by Income Group",
    x = "Household Income",
    y = "Greenness (NDVI)"
  ) +
  report_theme

ggsave("Figures/boxplot_ndvi.png", width = 5, height = 4, dpi = 300)


```



## Pollution by income using the potentially better variables 


PM25DALYY_01_COM
SO2OMIYY_COM
O3CHGYY_01_COM


```{r}

#PM25DALYY_01_COM by Income Group 

plot_data <- subset_data

ggplot(data = plot_data, aes(x = income_cat, y = PM25DALYY_01_COM)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(
    title = "PM2.5 1 year average, by Income Group",
    x = "Household Income",
    y = "PM2.5 (satellite, model-based) (µg/m³)"
  ) +
  report_theme
ggsave("Figures/boxplot_PM25DALYY_01_COM.png", width = 5, height = 4, dpi = 300)



plot_data %>% #check that there is significance for PM2.5 
  group_by(income_cat) %>%
  summarise(
    n = n(),
    median_pm25 = median(PM25DALYY_01_COM, na.rm = TRUE),
    mean_pm25 = round(mean(PM25DALYY_01_COM, na.rm = TRUE), 2)
  )
kruskal.test(PM25DALYY_01_COM ~ income_cat, data = df)







ggplot(plot_data, aes(x = income_cat, y = SO2OMIYY_COM)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(
    title = "SO2 3 year annual average at postal code, by Income Group",
    x = "Household Income",
    y = "3 y average SO2 (satellite derived)(ppb)"
  ) +
  report_theme

ggsave("Figures/boxplot_SO2OMIYY_COM.png", width = 5, height = 4, dpi = 300)



ggplot(plot_data, aes(x = income_cat, y = O3CHGYY_01_COM)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(
    title = "O3 annual average ground-level at postal code, by Income Group",
    x = "Household Income",
    y = "Annual average O3 (model-based) (ppb)"
  ) +
  report_theme

ggsave("Figures/boxplot_O3CHGYY_01_COM.png", width = 5, height = 4, dpi = 300)




```


##Pollution by wealth (savings)




```{r}
# PM25DALYY_01_COM by Savings/Wealth Group
plot_data <- subset_data

ggplot(data = plot_data, aes(x = savings_cat, y = PM25DALYY_01_COM)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(
    title = "PM2.5 1 year average, by Total Savings/Investments",
    x = "Total Savings/Investments",
    y = "PM2.5 (satellite, model-based) (µg/m³)"
  ) +
  report_theme
ggsave("Figures/boxplot_PM25DALYY_01_COM_savings.png", width = 5, height = 4, dpi = 300)


ggplot(plot_data, aes(x = savings_cat, y = SO2OMIYY_COM)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(
    title = "SO2 3 year annual average, by Total Savings/Investments",
    x = "Total Savings/Investments",
    y = "3 y average SO2 (satellite derived) (ppb)"
  ) +
  report_theme
ggsave("Figures/boxplot_SO2OMIYY_COM_savings.png", width = 5, height = 4, dpi = 300)

ggplot(plot_data, aes(x = savings_cat, y = O3CHGYY_01_COM)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(
    title = "O3 annual average, by Total Savings/Investments",
    x = "Total Savings/Investments",
    y = "Annual average O3 (model-based) (ppb)"
  ) +
  report_theme
ggsave("Figures/boxplot_O3CHGYY_01_COM_savings.png", width = 5, height = 4, dpi = 300)


```



















### median dot plot for PM2.5 by income group 


```{r}
df_pm25_summary <- df %>%
  group_by(income_cat) %>%
  summarise(
    n = n(),
    median_pm25 = median(pm25_1yr_com, na.rm = TRUE)
  )

p_pm25_dot <- ggplot(df_pm25_summary, aes(x = income_cat, y = median_pm25)) +
  geom_point(size = 4, color = "#2C7FB8", aes(shape = "Median")) +
  geom_text(aes(label = paste0("n=", n)), vjust = 2, size = 3.5) +
  scale_shape_manual(name = "", values = c("Median" = 16)) +
  scale_y_continuous(limits = c(6.0, 6.8)) +
  labs(title = "Median PM2.5 Exposure by Income Group",
       subtitle = "Kruskal-Wallis p = 0.032",
       x = "Household Income",
       y = "Median PM2.5 (µg/m³)") +
  report_theme

ggsave("Figures/pm25_median_dot.png", plot = p_pm25_dot,
       width = 6, height = 4, dpi = 300)
```



```{r}
air_income_plots <- list()

for (v in air_income_vars) {
  p <- subset_data %>%
    filter(income_cat %in% c("<20K", "20K–50K", "50K–100K", "100K–150K", "150K+")) %>%
    ggplot(aes(x = income_cat, y = .data[[v$col]], fill = income_cat)) +
    geom_boxplot(outlier.alpha = 0.3, width = 0.6) +
    scale_fill_brewer(palette = "Blues") +
    labs(
      title = v$title,
      x = "Household Income",
      y = v$y_lab
    ) +
    report_theme +
    theme(
      plot.title = element_text(size = 13, face = "bold"),
      axis.title = element_text(size = 11),
      axis.text = element_text(size = 9),
      legend.position = "none"
    )
  
  air_income_plots[[v$col]] <- p
}

dashboard_air_income <- wrap_plots(air_income_plots, ncol = 2) &
  theme(
    plot.margin = margin(t = 15, r = 15, b = 15, l = 15),
    plot.background = element_rect(color = "grey60", linewidth = 1, fill = "white"),
    panel.background = element_rect(fill = "white")
  )

dashboard_air_income <- dashboard_air_income +
  plot_annotation(
    title = "Air Pollution by Household Income",
    theme = theme(
      plot.title    = element_text(hjust = 0.5, face = "bold", size = 22),
      plot.subtitle = element_text(hjust = 0.5, size = 11, color = "grey40"),
      plot.caption  = element_text(size = 9, color = "grey40"),
      plot.margin   = margin(t = 20, r = 10, b = 10, l = 10),
      plot.background = element_rect(fill = "grey90", color = NA)
    )
  )

ggsave("Figures/air_pollution_income_dashboard.png", plot = dashboard_air_income, width = 16, height = 13, dpi = 300)
```





























## Condensation by income 

```{r}
#condensation by income

cond_only <- subset_data %>% #keep only those with condensation problems (54 people)
  filter(ENV_HMPRB_CON_MCQ == 1)

cond_counts <- cond_only %>%
  filter(INC_TOT_COM %in% 1:5) %>%
  mutate(income_cat = factor(INC_TOT_COM, levels = 1:5, labels = income_labels)) %>%
  count(income_cat, name = "n") %>%
  mutate(pct = n / sum(n))

cond_counts

ggplot(cond_counts, aes(x = income_cat, y = n)) +
  geom_col(fill = "#2C7FB8", width = 0.6) +
  geom_text(
    aes(label = paste0(n, " (", percent(pct, accuracy = 0.1), ")")),
    vjust = -0.35,
    size = 3.2  # scaled to base_size 11 theme
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Homes with Condensation, by Income Group",
    x = "Household Income",
    y = "Homes with Condensation"
  ) +
  report_theme

ggsave("Figures/barplot_condensation.png", width = 5, height = 4, dpi = 300)





ggplot(df, aes(x = income_cat, y = pm25_1yr_com)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(
    title = "PM2.5 1 y average, by Income Group",
    x = "Household Income",
    y = "PM2.5 (µg/m³)"
  ) +
  report_theme

ggsave("Figures/boxplot_pm25.png", width = 5, height = 4, dpi = 300)

```








## TCC by income
```{r}

#tree canopy by income, boxplots 
p_veg1 <- ggplot(df, aes(x = income_cat, y = grtcc10_01_COM)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(title = "Vegetation Cover (>5m) by Income Group",
       x = "Household Income", y = "% Pixel Covered by Vegetation >5m") +
  report_theme
ggsave("Figures/boxplot_vegetation_01.png", plot = p_veg1, width = 5, height = 4, dpi = 300)


p_veg2 <- ggplot(df, aes(x = income_cat, y = grtcc10_02_COM)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(title = "Vegetation Cover (>5m) by Income, 100m",
       x = "Household Income", y = "% Pixel Covered by Vegetation >5m") +
  report_theme
ggsave("Figures/boxplot_vegetation_02.png", plot = p_veg2, width = 5, height = 4, dpi = 300)


p_veg3 <- ggplot(df, aes(x = income_cat, y = grtcc10_03_COM)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(title = "Vegetation Cover (>5m) by Income, 250m",
       x = "Household Income", y = "% Pixel Covered by Vegetation >5m") +
  report_theme
ggsave("Figures/boxplot_vegetation_03.png", plot = p_veg3, width = 5, height = 4, dpi = 300)


p_veg4 <- ggplot(df, aes(x = income_cat, y = grtcc10_04_COM)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(title = "Vegetation Cover (>5m) by Income, 500m",
       x = "Household Income", y = "% Pixel Covered by Vegetation >5m") +
  report_theme
ggsave("Figures/boxplot_vegetation_04.png", plot = p_veg4, width = 5, height = 4, dpi = 300)


p_veg5 <- ggplot(df, aes(x = income_cat, y = grtcc10_05_COM)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(title = "Vegetation Cover (>5m) by Income, 1km",
       x = "Household Income", y = "% Pixel Covered by Vegetation >5m") +
  report_theme
ggsave("Figures/boxplot_vegetation_05.png", plot = p_veg5, width = 5, height = 4, dpi = 300)





df <- subset_data %>%
  filter(INC_TOT_COM %in% 1:5, !is.na(grtcc10_01_COM)) %>%
  mutate(
    income_cat = factor(INC_TOT_COM, levels = 1:5, labels = income_labels)
  )

ggplot(df, aes(x = income_cat, y = grtcc10_01_COM)) +
  geom_boxplot(
    fill = "#2C7FB8",
    outlier.alpha = 0.5,
    width = 0.6
  ) +
  labs(
    title = "Vegetation Cover (>5m) by Income Group",
    x = "Household Income",
    y = "Estimated % Pixel Covered by Vegetation >5m"
  ) +
  theme_classic(base_size = 26) +
  theme(
    legend.position = "none",
    
    plot.title = element_text(
      hjust = 0.5,
      face  = "bold",
      size  = 38,
      margin = margin(b = 18)
    ),
    
    axis.title.x = element_text(
      face = "bold",
      size = 30,
      margin = margin(t = 16)
    ),
    axis.title.y = element_text(
      face = "bold",
      size = 30,
      margin = margin(r = 16)
    ),
    
    axis.text.x = element_text(size = 26),
    axis.text.y = element_text(size = 26)
  )
```





## Veg greenness fractions by income 
There is no pattern by income group. The values are all clustered between 26–31%.
The GRUMPY variable measures all green vegetation including grass, shrubs, parks — not just tree canopy in residential areas.


```{r}

library(flextable)

df_green_summary <- df %>%
  group_by(income_cat) %>%
  summarise(
    n = n(),
    median_green = median(GRUMPYY_01_COM, na.rm = TRUE),
    mean_green = round(mean(GRUMPYY_01_COM, na.rm = TRUE), 1)
  ) %>%
  rename(
    `Income Group` = income_cat,
    `N` = n,
    `Median Greenness (%)` = median_green,
    `Mean Greenness (%)` = mean_green
  )

ft_green <- flextable(df_green_summary) %>%
  flextable::add_header_lines("Table X. Vegetation Greenness (GRUMPY) by Income Group") %>%
  flextable::bold(part = "header") %>%
  flextable::bold(j = "Income Group") %>%
  flextable::autofit() %>%
  flextable::theme_booktabs() %>%
  flextable::fontsize(size = 10, part = "all") %>%
  flextable::fontsize(i = 1, size = 11, part = "header") %>%
  flextable::bold(i = 1, part = "header") %>%
  flextable::font(fontname = "Times New Roman", part = "all") %>%
  flextable::align(i = 1, align = "center", part = "header") %>%
  flextable::align(align = "left", part = "all") %>%
  flextable::align(j = c("N", "Median Greenness (%)", "Mean Greenness (%)"), align = "center", part = "body")

save_as_image(ft_green, path = "Tables/Table_veg_green_fraction_byincome.png", res = 300)

```




## Road length (nearby) by income 
This was not a good variable to look at. 

```{r}
#median hwy length within 200 m, by income group, just the participants living near a hwy 

p_hwy_dot <- ggplot(df_hwy_summary, aes(x = income_cat, y = median_length)) +
  geom_point(size = 4, color = "#2C7FB8", aes(shape = "Median")) +
  geom_text(aes(label = paste0("n=", n)), vjust = 2, size = 3.5) +
  scale_shape_manual(name = "", values = c("Median" = 16)) +
  scale_y_continuous(limits = c(40, 105)) +
  labs(title = "Median Highway Length within 200m by Income Group",
       subtitle = "Among participants living near a highway only",
       x = "Household Income",
       y = "Median Highway Length (m)") +
  report_theme
ggsave("Figures/highway_median_dot.png", plot = p_hwy_dot, 
       width = 6, height = 4, dpi = 300)


#median road length within 200 m, by income group, just the participants living near a road 

df_local <- df %>% filter(Locrds_length_200m_com > 0)

df_local_summary <- df_local %>%
  group_by(income_cat) %>%
  summarise(
    n = n(),
    median_length = median(Locrds_length_200m_com, na.rm = TRUE)
  )

p_local_dot <- ggplot(df_local_summary, aes(x = income_cat, y = median_length)) +
  geom_point(size = 4, color = "#2C7FB8", aes(shape = "Median")) +
  geom_text(aes(label = paste0("n=", n)), vjust = 2, size = 3.5) +
  scale_shape_manual(name = "", values = c("Median" = 16)) +
  labs(title = "Median Local Road Length within 200m by Income Group",
       subtitle = "Among participants living near a local road only",
       x = "Household Income",
       y = "Median Local Road Length (m)") +
  report_theme

ggsave("Figures/localroad_median_dot.png", plot = p_local_dot,
       width = 6, height = 4, dpi = 300)

```































# 6.0 Linear Regressions
## 6.1 INCOME PREDICTING AGING (stratified by age)

```{r}

# Regression: Income predicting biological aging

class(subset_data$AGE_NMBR_COM)#these are integers, ok to leave alone
class(subset_data$CD4nv) 


#stratify by age 
subset_data$age_group <- ifelse(subset_data$AGE_NMBR_COM < 65, "45-64", "65-85")
subset_data_firstgrp  <- subset_data[subset_data$age_group == "45-64", ]
subset_data_secondgrp <- subset_data[subset_data$age_group == "65-85", ]

table(subset_data$INC_TOT_COM) #make sure it is reverse coded 


#GrimAGe 
library(broom)
model <- lm(GrimAge_EAA ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data_firstgrp)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "GrimAge_model_results_45to64.csv", row.names = FALSE)

model <- lm(GrimAge_EAA ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data_secondgrp)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "GrimAge_model_results_65to85.csv", row.names = FALSE)


# PhenoAge
model <- lm(PhenoAge_EAA ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data_firstgrp)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "PhenoAge_model_results_45to64.csv", row.names = FALSE)

model <- lm(PhenoAge_EAA ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data_secondgrp)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "PhenoAge_model_results_65to85.csv", row.names = FALSE)


# DunedinPACE
model <- lm(DunedinPACE ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data_firstgrp)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "DunedinPACE_model_results_45to64.csv", row.names = FALSE)

model <- lm(DunedinPACE ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data_secondgrp)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "DunedinPACE_model_results_65to85.csv", row.names = FALSE)


# DNAmIC
model <- lm(DNAmIC ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data_firstgrp)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "DNAmIC_model_results_45to64.csv", row.names = FALSE)

model <- lm(DNAmIC ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data_secondgrp)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "DNAmIC_model_results_65to85.csv", row.names = FALSE)









###########
#make into a nice table 

library(tidyr)

read_model <- function(file, outcome, age_group) {
  read.csv(file) %>%
    mutate(Outcome = outcome, Age_Group = age_group)
}

all_results <- bind_rows(
  read_model("regression_csvs/GrimAge_model_results_45to64.csv",     "GrimAge EAA",  "45-64"),
  read_model("regression_csvs/GrimAge_model_results_65to85.csv",     "GrimAge EAA",  "65-85"),
  read_model("regression_csvs/PhenoAge_model_results_45to64.csv",    "PhenoAge EAA", "45-64"),
  read_model("regression_csvs/PhenoAge_model_results_65to85.csv",    "PhenoAge EAA", "65-85"),
  read_model("regression_csvs/DunedinPACE_model_results_45to64.csv", "DunedinPACE",  "45-64"),
  read_model("regression_csvs/DunedinPACE_model_results_65to85.csv", "DunedinPACE",  "65-85"),
  read_model("regression_csvs/DNAmIC_model_results_45to64.csv",      "DNAmIC",       "45-64"),
  read_model("regression_csvs/DNAmIC_model_results_65to85.csv",      "DNAmIC",       "65-85")
)

table_data <- all_results %>%
  filter(grepl("INC_TOT_COM", term)) %>%
  mutate(
    term = recode(term,
      "INC_TOT_COM2" = "100-150K",
      "INC_TOT_COM3" = "50-100K",
      "INC_TOT_COM4" = "20-50K",
      "INC_TOT_COM5" = "<20K",
      "INC_TOT_COM6" = "Don't know/Refused"
    ),
    Result = ifelse(
      Outcome == "DNAmIC",
      paste0(formatC(estimate, format = "e", digits = 2),
             " (", formatC(conf.low, format = "e", digits = 2),
             ", ", formatC(conf.high, format = "e", digits = 2), ")"),
      ifelse(
        abs(estimate) < 0.01,
        paste0(round(estimate, 4), " (", round(conf.low, 4), ", ", round(conf.high, 4), ")"),
        paste0(round(estimate, 2), " (", round(conf.low, 2), ", ", round(conf.high, 2), ")")
      )
    ),
    sig = p.value < 0.05,
    p.value = ifelse(round(p.value, 3) == 0, "<0.001", as.character(round(p.value, 3))),
    Result = ifelse(sig, paste0("**", Result, "**"), Result),
    Result_with_p = paste0(Result, ", p=", p.value)
  ) %>%
  select(Outcome, Age_Group, term, Result_with_p)

table_wide <- table_data %>%
  pivot_wider(names_from = Outcome, values_from = Result_with_p)

table_45 <- table_wide %>% filter(Age_Group == "45-64") %>% select(-Age_Group)
table_65 <- table_wide %>% filter(Age_Group == "65-85") %>% select(-Age_Group)

table_final <- bind_rows(table_45, table_65)

write.csv(table_final, "all_model_results_summary.csv", row.names = FALSE)

save_kable(
  knitr::kable(table_final,
               col.names = c("Income Group", "GrimAge EAA", "PhenoAge EAA", "DunedinPACE", "DNAmIC"),
               caption = "Table X: Association between household income and epigenetic aging, stratified by age group",
               escape = TRUE) %>%
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) %>%
    pack_rows("45-64 years", 1, nrow(table_45)) %>%
    pack_rows("65-85 years", nrow(table_45) + 1, nrow(table_final)),
  "Figures/income_epigenetic_table.png", zoom = 2
)



```




















## 6.2 INCOME PREDICTING ENVIRONMENT 


```{r}
## do all the regressions first 

levels(subset_data$INC_TOT_COM)

#Income predicting PM2.5 
library(broom)
model <- lm(PM25DALYY_01_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_PM2.5.csv", row.names = FALSE)


#income predicting ozone
model <- lm(O3CHGYY_01_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_ozone.csv", row.names = FALSE)



#Income predicting NO2 
model <- lm(no2_1yr_com ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_NO2.csv", row.names = FALSE)



#Income predicting SO2 
model <- lm(so2_1yr_com ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_SO2.csv", row.names = FALSE)



#Income predicting condensation
subset_data$condensation_binary <- as.numeric(subset_data$ENV_HMPRB_CON_MCQ == "Yes") #first make condensation numeric b/c it is outcome now. 
model <- lm(condensation_binary ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_condensation.csv", row.names = FALSE)




############################
#TCC


# Tree canopy cover at postal code
model <- lm(grtcc10_01_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
               I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
             + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
             data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_TreeCanopyPostalCode.csv", row.names = FALSE)

# Tree canopy cover within 100m
model <- lm(grtcc10_02_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
               I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
             + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
             data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_TreeCanopy100m.csv", row.names = FALSE)

# Tree canopy cover within 250m
model <- lm(grtcc10_03_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
               I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
             + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
             data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_TreeCanopy250m.csv", row.names = FALSE)

# Tree canopy cover within 500m
model <- lm(grtcc10_04_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
               I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
             + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
             data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_TreeCanopy500m.csv", row.names = FALSE)

# Tree canopy cover within 1km
model <- lm(grtcc10_05_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
               I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
             + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
             data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_TreeCanopy1km.csv", row.names = FALSE)




#####################
#% veg greenness 

# Vegetative greenness at postal code
model <- lm(GRUMPYY_01_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
               I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
             + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
             data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_GreennessPostalCode.csv", row.names = FALSE)

# Vegetative greenness mean within 100m
model <- lm(GRUMPYY_02_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
               I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
             + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
             data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_GreennessMean100m.csv", row.names = FALSE)

# Vegetative greenness mean within 250m
model <- lm(GRUMPYY_03_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
               I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
             + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
             data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_GreennessMean250m.csv", row.names = FALSE)

# Vegetative greenness mean within 500m
model <- lm(GRUMPYY_04_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
               I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
             + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
             data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_GreennessMean500m.csv", row.names = FALSE)

# Vegetative greenness mean within 1000m
model <- lm(GRUMPYY_05_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
               I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
             + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
             data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_GreennessMean1000m.csv", row.names = FALSE)

# Vegetative greenness max within 100m
model <- lm(GRUMPYY_06_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
               I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
             + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
             data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_GreennessMax100m.csv", row.names = FALSE)

# Vegetative greenness max within 250m
model <- lm(GRUMPYY_07_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
               I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
             + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
             data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_GreennessMax250m.csv", row.names = FALSE)

# Vegetative greenness max within 500m
model <- lm(GRUMPYY_08_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
               I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
             + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
             data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_GreennessMax500m.csv", row.names = FALSE)

# Vegetative greenness max within 1000m
model <- lm(GRUMPYY_09_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
               I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
             + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
             data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_GreennessMax1000m.csv", row.names = FALSE)


#####################
# NDVI (Max 500m)
model <- lm(GRLANYY_26_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
               I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
             + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
             data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Income_NDVI.csv", row.names = FALSE)






```




```{r}

#Make nice tables for regressions of income predicting environment 

levels(subset_data$INC_TOT_COM)

income_labels <- c(
  "INC_TOT_COM100\u2013150K"                = "100-150K",
  "INC_TOT_COM50\u2013100K"                 = "50-100K",
  "INC_TOT_COM20\u201350K"                  = "20-50K",
  "INC_TOT_COM<20K"                         = "<20K",
  "INC_TOT_COMDon't know/No answer/Refused" = "Don't know/Refused"
)

read_model <- function(file, outcome) {
  read.csv(file) %>% mutate(Outcome = outcome)
}

format_result <- function(data) {
  data %>%
    filter(grepl("INC_TOT_COM", term)) %>%
    mutate(
      term = recode(term, !!!income_labels),
      Result = ifelse(
        abs(estimate) < 0.01,
        paste0(round(estimate, 4), " (", round(conf.low, 4), ", ", round(conf.high, 4), ")"),
        paste0(round(estimate, 2), " (", round(conf.low, 2), ", ", round(conf.high, 2), ")")
      ),
      sig = p.value < 0.05,
      p.value = ifelse(round(p.value, 3) == 0, "<0.001", as.character(round(p.value, 3))),
      Result = ifelse(sig, paste0("**", Result, "**"), Result),
      Result_with_p = paste0(Result, ", p=", p.value)
    )
}

make_table <- function(data, col_names, caption, filename) {
  save_kable(
    knitr::kable(data, col.names = col_names, caption = caption, escape = TRUE) %>%
      kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE),
    filename, zoom = 2
  )
}

# Air pollution
bind_rows(
  read_model("regression_csvs/Income_PM2.5.csv",  "PM2.5"),
  read_model("regression_csvs/Income_ozone.csv",  "Ozone (O3)"),
  read_model("regression_csvs/Income_NO2.csv",    "NO2"),
  read_model("regression_csvs/Income_SO2.csv",    "SO2")
) %>% format_result() %>%
  select(Outcome, term, Result_with_p) %>%
  pivot_wider(names_from = Outcome, values_from = Result_with_p) %>%
  make_table(c("Income Group", "PM2.5", "Ozone (O3)", "NO2", "SO2"),
             "Table X: Association between household income and air pollution exposures",
             "Figures/income_airpollution_table.png")

# Condensation
read_model("regression_csvs/Income_condensation.csv", "Condensation") %>%
  format_result() %>%
  select(term, Result_with_p) %>%
  make_table(c("Income Group", "Condensation"),
             "Table X: Association between household income and housing condensation",
             "Figures/income_condensation_table.png")

# Tree Canopy Cover
bind_rows(
  read_model("regression_csvs/Income_TreeCanopyPostalCode.csv", "Postal code"),
  read_model("regression_csvs/Income_TreeCanopy100m.csv",       "100m"),
  read_model("regression_csvs/Income_TreeCanopy250m.csv",       "250m"),
  read_model("regression_csvs/Income_TreeCanopy500m.csv",       "500m"),
  read_model("regression_csvs/Income_TreeCanopy1km.csv",        "1km")
) %>% format_result() %>%
  select(Outcome, term, Result_with_p) %>%
  pivot_wider(names_from = Outcome, values_from = Result_with_p) %>%
  make_table(c("Income Group", "Postal code", "100m", "250m", "500m", "1km"),
             "Table X: Association between household income and tree canopy cover",
             "Figures/income_treecanopy_table.png")

# Vegetative Greenness
bind_rows(
  read_model("regression_csvs/Income_GreennessPostalCode.csv", "Postal code"),
  read_model("regression_csvs/Income_GreennessMean100m.csv",   "Mean 100m"),
  read_model("regression_csvs/Income_GreennessMean250m.csv",   "Mean 250m"),
  read_model("regression_csvs/Income_GreennessMean500m.csv",   "Mean 500m"),
  read_model("regression_csvs/Income_GreennessMean1000m.csv",  "Mean 1000m"),
  read_model("regression_csvs/Income_GreennessMax100m.csv",    "Max 100m"),
  read_model("regression_csvs/Income_GreennessMax250m.csv",    "Max 250m"),
  read_model("regression_csvs/Income_GreennessMax500m.csv",    "Max 500m"),
  read_model("regression_csvs/Income_GreennessMax1000m.csv",   "Max 1000m")
) %>% format_result() %>%
  select(Outcome, term, Result_with_p) %>%
  pivot_wider(names_from = Outcome, values_from = Result_with_p) %>%
  make_table(c("Income Group", "Postal code", "Mean 100m", "Mean 250m", "Mean 500m",
               "Mean 1000m", "Max 100m", "Max 250m", "Max 500m", "Max 1000m"),
             "Table X: Association between household income and vegetative greenness",
             "Figures/income_greenness_table.png")

# NDVI
read_model("regression_csvs/Income_NDVI.csv", "NDVI (Max 500m)") %>%
  format_result() %>%
  select(term, Result_with_p) %>%
  make_table(c("Income Group", "NDVI (Max 500m)"),
             "Table X: Association between household income and NDVI",
             "Figures/income_NDVI_table.png")
```









## 6.3 ENVIRONMENT PREDICTING AGING


```{r}
#Regressions 
library(broom)
library(tidyr)
library(dplyr)
library(kableExtra)

# ── Shared setup ──────────────────────────────────────────────────────────────

covariates <- "SEX_ASK_COM + AGE_NMBR_COM + I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7 +
               CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv"

outcomes <- c(
  GrimAge     = "GrimAge_EAA",
  PhenoAge    = "PhenoAge_EAA",
  DunedinPACE = "DunedinPACE",
  DNAmIC      = "DNAmIC"
)

run_model <- function(outcome_var, predictor, filepath) {
  model <- lm(as.formula(paste(outcome_var, "~", predictor, "+", covariates)),
              data = subset_data)
  write.csv(tidy(model, conf.int = TRUE, conf.level = 0.95), filepath, row.names = FALSE)
}

# ── Air pollution ─────────────────────────────────────────────────────────────

air_preds <- c(
  PM2.5 = "PM25DALYY_01_COM",
  Ozone = "O3CHGYY_01_COM",
  NO2   = "no2_1yr_com",
  SO2   = "so2_1yr_com"
)

for (env in names(air_preds)) {
  for (out in names(outcomes)) {
    run_model(outcomes[out], air_preds[env],
              paste0("regression_csvs/Air_", env, "_", out, ".csv"))
  }
}

# ── TCC ───────────────────────────────────────────────────────────────────────

tcc_preds  <- setNames(paste0("grtcc10_0", 1:5, "_COM"),
                       c("PostalCode", "100m", "250m", "500m", "1km"))

for (env in names(tcc_preds)) {
  for (out in names(outcomes)) {
    run_model(outcomes[out], tcc_preds[env],
              paste0("regression_csvs/TCC_", env, "_", out, ".csv"))
  }
}

# ── Vegetative Greenness ──────────────────────────────────────────────────────

green_preds <- setNames(paste0("GRUMPYY_0", 1:9, "_COM"),
                        c("PostalCode", "Mean100m", "Mean250m", "Mean500m", "Mean1000m",
                          "Max100m", "Max250m", "Max500m", "Max1000m"))

for (env in names(green_preds)) {
  for (out in names(outcomes)) {
    run_model(outcomes[out], green_preds[env],
              paste0("regression_csvs/greenness_", env, "_", out, ".csv"))
  }
}

# ── Condensation ──────────────────────────────────────────────────────────────

for (out in names(outcomes)) {
  run_model(outcomes[out], "ENV_HMPRB_CON_MCQ",
            paste0("regression_csvs/Condensation_", out, ".csv"))
}

# ── NDVI ──────────────────────────────────────────────────────────────────────

for (out in names(outcomes)) {
  run_model(outcomes[out], "GRLANYY_26_COM",
            paste0("regression_csvs/NDVI_", out, ".csv"))
}

```

```{r}
#nice tables 

# ── Shared stuff  ────────────────────────────────────────────────────────────

read_model <- function(file, env, outcome) {
  read.csv(file) %>% mutate(Env = env, Outcome = outcome)
}

format_and_save <- function(all_results, term_filter, col_names, caption, filename) {
  table_data <- all_results %>%
    filter(term %in% term_filter) %>%
    mutate(
      Result = ifelse(
        Outcome == "DNAmIC",
        paste0(formatC(estimate, format = "e", digits = 2),
               " (", formatC(conf.low, format = "e", digits = 2),
               ", ", formatC(conf.high, format = "e", digits = 2), ")"),
        ifelse(
          abs(estimate) < 0.01,
          paste0(round(estimate, 4), " (", round(conf.low, 4), ", ", round(conf.high, 4), ")"),
          paste0(round(estimate, 2), " (", round(conf.low, 2), ", ", round(conf.high, 2), ")")
        )
      ),
      sig     = p.value < 0.05,
      p.value = ifelse(round(p.value, 3) == 0, "<0.001", as.character(round(p.value, 3))),
      Result  = ifelse(sig, paste0("**", Result, "**"), Result),
      Result_with_p = paste0(Result, ", p=", p.value)
    ) %>%
    select(Env, Outcome, Result_with_p) %>%
    pivot_wider(names_from = Outcome, values_from = Result_with_p)

  save_kable(
    knitr::kable(table_data, col.names = col_names, caption = caption, escape = TRUE) %>%
      kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE),
    filename, zoom = 2
  )
  invisible(table_data)
}

aging_cols <- c("Exposure", "GrimAge EAA", "PhenoAge EAA", "DunedinPACE", "DNAmIC")

load_results <- function(prefix, envs, env_labels) {
  bind_rows(lapply(seq_along(envs), function(i) {
    bind_rows(lapply(names(outcomes), function(out) {
      read_model(paste0("regression_csvs/", prefix, envs[i], "_", out, ".csv"),
                 env_labels[i], out)
    }))
  }))
}

# ── Air pollution ─────────────────────────────────────────────────────────────

air_envs   <- c("PM2.5", "Ozone", "NO2", "SO2")
air_terms  <- c("PM25DALYY_01_COM", "O3CHGYY_01_COM", "no2_1yr_com", "so2_1yr_com")

air_results <- load_results("Air_", air_envs, air_envs)

format_and_save(air_results, air_terms,
  c("Air Pollution", "GrimAge EAA", "PhenoAge EAA", "DunedinPACE", "DNAmIC"),
  "Table X: Association between air pollution exposures and epigenetic aging",
  "Figures/airpollution_aging_table.png")

# ── TCC ───────────────────────────────────────────────────────────────────────

tcc_file_labels <- c("PostalCode", "100m", "250m", "500m", "1km")
tcc_env_labels  <- c("Postal Code", "100m", "250m", "500m", "1km")
tcc_terms       <- paste0("grtcc10_0", 1:5, "_COM")

tcc_results <- load_results("TCC_", tcc_file_labels, tcc_env_labels)

format_and_save(tcc_results, tcc_terms,
  c("Buffer", "GrimAge EAA", "PhenoAge EAA", "DunedinPACE", "DNAmIC"),
  "Table X: Association between tree canopy cover and epigenetic aging",
  "Figures/TCC_aging_table.png")

# ── Vegetative Greenness ──────────────────────────────────────────────────────

green_file_labels <- c("PostalCode", "Mean100m", "Mean250m", "Mean500m", "Mean1000m",
                       "Max100m", "Max250m", "Max500m", "Max1000m")
green_env_labels  <- c("Postal code", "Mean 100m", "Mean 250m", "Mean 500m", "Mean 1000m",
                       "Max 100m", "Max 250m", "Max 500m", "Max 1000m")
green_terms       <- paste0("GRUMPYY_0", 1:9, "_COM")

green_results <- load_results("greenness_", green_file_labels, green_env_labels)

format_and_save(green_results, green_terms,
  c("Buffer", "GrimAge EAA", "PhenoAge EAA", "DunedinPACE", "DNAmIC"),
  "Table X: Association between vegetative greenness and epigenetic aging",
  "Figures/greenness_aging_table.png")

# ── Condensation ──────────────────────────────────────────────────────────────

cond_results <- bind_rows(lapply(names(outcomes), function(out) {
  read_model(paste0("regression_csvs/Condensation_", out, ".csv"), "Condensation", out)
}))

format_and_save(cond_results, "ENV_HMPRB_CON_MCQYes", aging_cols,
  "Table X: Association between housing condensation and epigenetic aging",
  "Figures/condensation_aging_table.png")

# ── NDVI ──────────────────────────────────────────────────────────────────────

ndvi_results <- bind_rows(lapply(names(outcomes), function(out) {
  read_model(paste0("regression_csvs/NDVI_", out, ".csv"), "NDVI (Max 500m)", out)
}))

format_and_save(ndvi_results, "GRLANYY_26_COM", aging_cols,
  "Table X: Association between NDVI and epigenetic aging",
  "Figures/NDVI_aging_table.png")

```




















### Air pollution predicting aging 

```{r}
library(broom)

str(subset_data)

# GrimAge EAA
model <- lm(GrimAge_EAA ~ PM25DALYY_01_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Air_PM25_GrimAge.csv", row.names = FALSE)

model <- lm(GrimAge_EAA ~ O3CHGYY_01_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Air_Ozone_GrimAge.csv", row.names = FALSE)

model <- lm(GrimAge_EAA ~ no2_1yr_com + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Air_NO2_GrimAge.csv", row.names = FALSE)

model <- lm(GrimAge_EAA ~ so2_1yr_com + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Air_SO2_GrimAge.csv", row.names = FALSE)


# PhenoAge EAA
model <- lm(PhenoAge_EAA ~ PM25DALYY_01_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Air_PM25_PhenoAge.csv", row.names = FALSE)

model <- lm(PhenoAge_EAA ~ O3CHGYY_01_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Air_Ozone_PhenoAge.csv", row.names = FALSE)

model <- lm(PhenoAge_EAA ~ no2_1yr_com + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Air_NO2_PhenoAge.csv", row.names = FALSE)

model <- lm(PhenoAge_EAA ~ so2_1yr_com + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Air_SO2_PhenoAge.csv", row.names = FALSE)


# DunedinPACE
model <- lm(DunedinPACE ~ PM25DALYY_01_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Air_PM25_DunedinPACE.csv", row.names = FALSE)

model <- lm(DunedinPACE ~ O3CHGYY_01_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Air_Ozone_DunedinPACE.csv", row.names = FALSE)

model <- lm(DunedinPACE ~ no2_1yr_com + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Air_NO2_DunedinPACE.csv", row.names = FALSE)

model <- lm(DunedinPACE ~ so2_1yr_com + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Air_SO2_DunedinPACE.csv", row.names = FALSE)


# DNAmIC
model <- lm(DNAmIC ~ PM25DALYY_01_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Air_PM25_DNAmIC.csv", row.names = FALSE)

model <- lm(DNAmIC ~ O3CHGYY_01_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Air_Ozone_DNAmIC.csv", row.names = FALSE)

model <- lm(DNAmIC ~ no2_1yr_com + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Air_NO2_DNAmIC.csv", row.names = FALSE)

model <- lm(DNAmIC ~ so2_1yr_com + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Air_SO2_DNAmIC.csv", row.names = FALSE)













#####make a nice table of air pollution predicting aging 

library(tidyr)
library(dplyr)
library(kableExtra)

read_model <- function(file, env) {
  read.csv(file) %>%
    mutate(Env = env)
}

all_results <- bind_rows(
  read_model("regression_csvs/Air_PM25_GrimAge.csv",      "PM2.5"),
  read_model("regression_csvs/Air_Ozone_GrimAge.csv",     "Ozone"),
  read_model("regression_csvs/Air_NO2_GrimAge.csv",       "NO2"),
  read_model("regression_csvs/Air_SO2_GrimAge.csv",       "SO2"),
  read_model("regression_csvs/Air_PM25_PhenoAge.csv",     "PM2.5"),
  read_model("regression_csvs/Air_Ozone_PhenoAge.csv",    "Ozone"),
  read_model("regression_csvs/Air_NO2_PhenoAge.csv",      "NO2"),
  read_model("regression_csvs/Air_SO2_PhenoAge.csv",      "SO2"),
  read_model("regression_csvs/Air_PM25_DunedinPACE.csv",  "PM2.5"),
  read_model("regression_csvs/Air_Ozone_DunedinPACE.csv", "Ozone"),
  read_model("regression_csvs/Air_NO2_DunedinPACE.csv",   "NO2"),
  read_model("regression_csvs/Air_SO2_DunedinPACE.csv",   "SO2"),
  read_model("regression_csvs/Air_PM25_DNAmIC.csv",       "PM2.5"),
  read_model("regression_csvs/Air_Ozone_DNAmIC.csv",      "Ozone"),
  read_model("regression_csvs/Air_NO2_DNAmIC.csv",        "NO2"),
  read_model("regression_csvs/Air_SO2_DNAmIC.csv",        "SO2")
)

# Add outcome label based on filename pattern
all_results <- bind_rows(
  read_model("regression_csvs/Air_PM25_GrimAge.csv",      "PM2.5")    %>% mutate(Outcome = "GrimAge EAA"),
  read_model("regression_csvs/Air_Ozone_GrimAge.csv",     "Ozone")    %>% mutate(Outcome = "GrimAge EAA"),
  read_model("regression_csvs/Air_NO2_GrimAge.csv",       "NO2")      %>% mutate(Outcome = "GrimAge EAA"),
  read_model("regression_csvs/Air_SO2_GrimAge.csv",       "SO2")      %>% mutate(Outcome = "GrimAge EAA"),
  read_model("regression_csvs/Air_PM25_PhenoAge.csv",     "PM2.5")    %>% mutate(Outcome = "PhenoAge EAA"),
  read_model("regression_csvs/Air_Ozone_PhenoAge.csv",    "Ozone")    %>% mutate(Outcome = "PhenoAge EAA"),
  read_model("regression_csvs/Air_NO2_PhenoAge.csv",      "NO2")      %>% mutate(Outcome = "PhenoAge EAA"),
  read_model("regression_csvs/Air_SO2_PhenoAge.csv",      "SO2")      %>% mutate(Outcome = "PhenoAge EAA"),
  read_model("regression_csvs/Air_PM25_DunedinPACE.csv",  "PM2.5")    %>% mutate(Outcome = "DunedinPACE"),
  read_model("regression_csvs/Air_Ozone_DunedinPACE.csv", "Ozone")    %>% mutate(Outcome = "DunedinPACE"),
  read_model("regression_csvs/Air_NO2_DunedinPACE.csv",   "NO2")      %>% mutate(Outcome = "DunedinPACE"),
  read_model("regression_csvs/Air_SO2_DunedinPACE.csv",   "SO2")      %>% mutate(Outcome = "DunedinPACE"),
  read_model("regression_csvs/Air_PM25_DNAmIC.csv",       "PM2.5")    %>% mutate(Outcome = "DNAmIC"),
  read_model("regression_csvs/Air_Ozone_DNAmIC.csv",      "Ozone")    %>% mutate(Outcome = "DNAmIC"),
  read_model("regression_csvs/Air_NO2_DNAmIC.csv",        "NO2")      %>% mutate(Outcome = "DNAmIC"),
  read_model("regression_csvs/Air_SO2_DNAmIC.csv",        "SO2")      %>% mutate(Outcome = "DNAmIC")
)

# Keep only the air pollution predictor row
air_terms <- c("PM25DALYY_01_COM", "O3CHGYY_01_COM", "no2_1yr_com", "so2_1yr_com")

table_data <- all_results %>%
  filter(term %in% air_terms) %>%
  mutate(
    p.value = round(p.value, 3),
    Result = ifelse(
      Outcome == "DNAmIC",
      paste0(formatC(estimate, format = "e", digits = 2),
             " (", formatC(conf.low, format = "e", digits = 2),
             ", ", formatC(conf.high, format = "e", digits = 2), ")"),
      ifelse(
        abs(estimate) < 0.01,
        paste0(round(estimate, 4), " (", round(conf.low, 4), ", ", round(conf.high, 4), ")"),
        paste0(round(estimate, 2), " (", round(conf.low, 2), ", ", round(conf.high, 2), ")")
      )
    ),
    Result = ifelse(p.value < 0.05, paste0("**", Result, "**"), Result),
    Result_with_p = paste0(Result, ", p=", p.value)
    
  ) %>%
  select(Env, Outcome, Result_with_p)

table_wide <- table_data %>%
  pivot_wider(names_from = Outcome, values_from = Result_with_p)

write.csv(table_wide, "regression_csvs/airpollution_aging_results.csv", row.names = FALSE)

save_kable(
  knitr::kable(table_wide,
               col.names = c("Air Pollution", "GrimAge EAA", "PhenoAge EAA", "DunedinPACE", "DNAmIC"),
               caption = "Table X: Association between air pollution exposures and epigenetic aging",
               escape = TRUE) %>%
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE),
  "Figures/airpollution_aging_table.png", zoom = 2
)
  
```





```{r}

```










































### TCC predicting aging 

```{r}

# GrimAge EAA
model <- lm(GrimAge_EAA ~ grtcc10_01_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/TCC_GrimAge.csv", row.names = FALSE)

model <- lm(GrimAge_EAA ~ grtcc10_02_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/TCC_100m_GrimAge.csv", row.names = FALSE)

model <- lm(GrimAge_EAA ~ grtcc10_03_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/TCC_250m_GrimAge.csv", row.names = FALSE)

model <- lm(GrimAge_EAA ~ grtcc10_04_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/TCC_500m_GrimAge.csv", row.names = FALSE)

model <- lm(GrimAge_EAA ~ grtcc10_05_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/TCC_1km_GrimAge.csv", row.names = FALSE)


# PhenoAge EAA

model <- lm(PhenoAge_EAA ~ grtcc10_01_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/TCC_PhenoAge.csv", row.names = FALSE)


model <- lm(PhenoAge_EAA ~ grtcc10_02_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/TCC_100m_PhenoAge.csv", row.names = FALSE)

model <- lm(PhenoAge_EAA ~ grtcc10_03_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/TCC_250m_PhenoAge.csv", row.names = FALSE)

model <- lm(PhenoAge_EAA ~ grtcc10_04_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/TCC_500m_PhenoAge.csv", row.names = FALSE)

model <- lm(PhenoAge_EAA ~ grtcc10_05_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/TCC_1km_PhenoAge.csv", row.names = FALSE)


# DunedinPACE

model <- lm(DunedinPACE ~ grtcc10_01_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/TCC_DunedinPACE.csv", row.names = FALSE)


model <- lm(DunedinPACE ~ grtcc10_02_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/TCC_100m_DunedinPACE.csv", row.names = FALSE)

model <- lm(DunedinPACE ~ grtcc10_03_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/TCC_250m_DunedinPACE.csv", row.names = FALSE)

model <- lm(DunedinPACE ~ grtcc10_04_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/TCC_500m_DunedinPACE.csv", row.names = FALSE)

model <- lm(DunedinPACE ~ grtcc10_05_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/TCC_1km_DunedinPACE.csv", row.names = FALSE)


# DNAmIC

model <- lm(DNAmIC ~ grtcc10_01_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/TCC_DNAmIC.csv", row.names = FALSE)


model <- lm(DNAmIC ~ grtcc10_02_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/TCC_100m_DNAmIC.csv", row.names = FALSE)

model <- lm(DNAmIC ~ grtcc10_03_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/TCC_250m_DNAmIC.csv", row.names = FALSE)

model <- lm(DNAmIC ~ grtcc10_04_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/TCC_500m_DNAmIC.csv", row.names = FALSE)

model <- lm(DNAmIC ~ grtcc10_05_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/TCC_1km_DNAmIC.csv", row.names = FALSE)







##############
###nice table: 


library(tidyr)
library(dplyr)
library(kableExtra)

read_model <- function(file, env) {
  read.csv(file) %>%
    mutate(Env = env)
}

all_results <- bind_rows(
  read_model("regression_csvs/TCC_GrimAge.csv",          "Postal Code") %>% mutate(Outcome = "GrimAge EAA"),
  read_model("regression_csvs/TCC_100m_GrimAge.csv",     "100m")        %>% mutate(Outcome = "GrimAge EAA"),
  read_model("regression_csvs/TCC_250m_GrimAge.csv",     "250m")        %>% mutate(Outcome = "GrimAge EAA"),
  read_model("regression_csvs/TCC_500m_GrimAge.csv",     "500m")        %>% mutate(Outcome = "GrimAge EAA"),
  read_model("regression_csvs/TCC_1km_GrimAge.csv",      "1km")         %>% mutate(Outcome = "GrimAge EAA"),
  read_model("regression_csvs/TCC_PhenoAge.csv",         "Postal Code") %>% mutate(Outcome = "PhenoAge EAA"),
  read_model("regression_csvs/TCC_100m_PhenoAge.csv",    "100m")        %>% mutate(Outcome = "PhenoAge EAA"),
  read_model("regression_csvs/TCC_250m_PhenoAge.csv",    "250m")        %>% mutate(Outcome = "PhenoAge EAA"),
  read_model("regression_csvs/TCC_500m_PhenoAge.csv",    "500m")        %>% mutate(Outcome = "PhenoAge EAA"),
  read_model("regression_csvs/TCC_1km_PhenoAge.csv",     "1km")         %>% mutate(Outcome = "PhenoAge EAA"),
  read_model("regression_csvs/TCC_DunedinPACE.csv",      "Postal Code") %>% mutate(Outcome = "DunedinPACE"),
  read_model("regression_csvs/TCC_100m_DunedinPACE.csv", "100m")        %>% mutate(Outcome = "DunedinPACE"),
  read_model("regression_csvs/TCC_250m_DunedinPACE.csv", "250m")        %>% mutate(Outcome = "DunedinPACE"),
  read_model("regression_csvs/TCC_500m_DunedinPACE.csv", "500m")        %>% mutate(Outcome = "DunedinPACE"),
  read_model("regression_csvs/TCC_1km_DunedinPACE.csv",  "1km")         %>% mutate(Outcome = "DunedinPACE"),
  read_model("regression_csvs/TCC_DNAmIC.csv",           "Postal Code") %>% mutate(Outcome = "DNAmIC"),
  read_model("regression_csvs/TCC_100m_DNAmIC.csv",      "100m")        %>% mutate(Outcome = "DNAmIC"),
  read_model("regression_csvs/TCC_250m_DNAmIC.csv",      "250m")        %>% mutate(Outcome = "DNAmIC"),
  read_model("regression_csvs/TCC_500m_DNAmIC.csv",      "500m")        %>% mutate(Outcome = "DNAmIC"),
  read_model("regression_csvs/TCC_1km_DNAmIC.csv",       "1km")         %>% mutate(Outcome = "DNAmIC")
)

tcc_terms <- c("grtcc10_01_COM", "grtcc10_02_COM", "grtcc10_03_COM", "grtcc10_04_COM", "grtcc10_05_COM")




table_data <- all_results %>%
  filter(term %in% tcc_terms) %>%
  mutate(
    p.value = round(p.value, 3),
    Result = ifelse(
      Outcome == "DNAmIC",
      paste0(formatC(estimate, format = "e", digits = 2),
             " (", formatC(conf.low, format = "e", digits = 2),
             ", ", formatC(conf.high, format = "e", digits = 2), ")"),
      ifelse(
        abs(estimate) < 0.01,
        paste0(round(estimate, 4), " (", round(conf.low, 4), ", ", round(conf.high, 4), ")"),
        paste0(round(estimate, 2), " (", round(conf.low, 2), ", ", round(conf.high, 2), ")")
      )
    ),
    Result = ifelse(p.value < 0.05, paste0("**", Result, "**"), Result),
    Result_with_p = paste0(Result, " p=", p.value)
  ) %>%
  select(Env, Outcome, Result_with_p)

table_wide <- table_data %>%
  pivot_wider(names_from = Outcome, values_from = Result_with_p)




write.csv(table_wide, "regression_csvs/TCC_aging_results.csv", row.names = FALSE)

save_kable(
  knitr::kable(table_wide,
               col.names = c("Buffer", "GrimAge EAA", "PhenoAge EAA", "DunedinPACE", "DNAmIC"),
               caption = "Table X: Association between tree canopy cover and epigenetic aging",
               escape = TRUE) %>%
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE),
  "Figures/TCC_aging_table.png", zoom = 2
)
```

### %veg predicting aging 

```{r}


# GrimAge EAA
model <- lm(GrimAge_EAA ~ GRUMPYY_01_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/greenness_PostalCode_GrimAge.csv", row.names = FALSE)

model <- lm(GrimAge_EAA ~ GRUMPYY_02_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/greenness_Mean100m_GrimAge.csv", row.names = FALSE)

model <- lm(GrimAge_EAA ~ GRUMPYY_03_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/greenness_Mean250m_GrimAge.csv", row.names = FALSE)

model <- lm(GrimAge_EAA ~ GRUMPYY_04_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/greenness_Mean500m_GrimAge.csv", row.names = FALSE)

model <- lm(GrimAge_EAA ~ GRUMPYY_05_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/greenness_Mean1000m_GrimAge.csv", row.names = FALSE)

model <- lm(GrimAge_EAA ~ GRUMPYY_06_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/greenness_Max100m_GrimAge.csv", row.names = FALSE)

model <- lm(GrimAge_EAA ~ GRUMPYY_07_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/greenness_Max250m_GrimAge.csv", row.names = FALSE)

model <- lm(GrimAge_EAA ~ GRUMPYY_08_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/greenness_Max500m_GrimAge.csv", row.names = FALSE)

model <- lm(GrimAge_EAA ~ GRUMPYY_09_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/greenness_Max1000m_GrimAge.csv", row.names = FALSE)


# PhenoAge EAA
model <- lm(PhenoAge_EAA ~ GRUMPYY_01_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/greenness_PostalCode_PhenoAge.csv", row.names = FALSE)

model <- lm(PhenoAge_EAA ~ GRUMPYY_02_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/greenness_Mean100m_PhenoAge.csv", row.names = FALSE)

model <- lm(PhenoAge_EAA ~ GRUMPYY_03_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/greenness_Mean250m_PhenoAge.csv", row.names = FALSE)

model <- lm(PhenoAge_EAA ~ GRUMPYY_04_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/greenness_Mean500m_PhenoAge.csv", row.names = FALSE)

model <- lm(PhenoAge_EAA ~ GRUMPYY_05_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/greenness_Mean1000m_PhenoAge.csv", row.names = FALSE)

model <- lm(PhenoAge_EAA ~ GRUMPYY_06_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/greenness_Max100m_PhenoAge.csv", row.names = FALSE)

model <- lm(PhenoAge_EAA ~ GRUMPYY_07_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/greenness_Max250m_PhenoAge.csv", row.names = FALSE)

model <- lm(PhenoAge_EAA ~ GRUMPYY_08_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/greenness_Max500m_PhenoAge.csv", row.names = FALSE)

model <- lm(PhenoAge_EAA ~ GRUMPYY_09_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/greenness_Max1000m_PhenoAge.csv", row.names = FALSE)


# DunedinPACE
model <- lm(DunedinPACE ~ GRUMPYY_01_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/greenness_PostalCode_DunedinPACE.csv", row.names = FALSE)

model <- lm(DunedinPACE ~ GRUMPYY_02_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/greenness_Mean100m_DunedinPACE.csv", row.names = FALSE)

model <- lm(DunedinPACE ~ GRUMPYY_03_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/greenness_Mean250m_DunedinPACE.csv", row.names = FALSE)

model <- lm(DunedinPACE ~ GRUMPYY_04_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/greenness_Mean500m_DunedinPACE.csv", row.names = FALSE)

model <- lm(DunedinPACE ~ GRUMPYY_05_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/greenness_Mean1000m_DunedinPACE.csv", row.names = FALSE)

model <- lm(DunedinPACE ~ GRUMPYY_06_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/greenness_Max100m_DunedinPACE.csv", row.names = FALSE)

model <- lm(DunedinPACE ~ GRUMPYY_07_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/greenness_Max250m_DunedinPACE.csv", row.names = FALSE)

model <- lm(DunedinPACE ~ GRUMPYY_08_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/greenness_Max500m_DunedinPACE.csv", row.names = FALSE)

model <- lm(DunedinPACE ~ GRUMPYY_09_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/greenness_Max1000m_DunedinPACE.csv", row.names = FALSE)


# DNAmIC
model <- lm(DNAmIC ~ GRUMPYY_01_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/greenness_PostalCode_DNAmIC.csv", row.names = FALSE)

model <- lm(DNAmIC ~ GRUMPYY_02_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/greenness_Mean100m_DNAmIC.csv", row.names = FALSE)

model <- lm(DNAmIC ~ GRUMPYY_03_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/greenness_Mean250m_DNAmIC.csv", row.names = FALSE)

model <- lm(DNAmIC ~ GRUMPYY_04_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/greenness_Mean500m_DNAmIC.csv", row.names = FALSE)

model <- lm(DNAmIC ~ GRUMPYY_05_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/greenness_Mean1000m_DNAmIC.csv", row.names = FALSE)

model <- lm(DNAmIC ~ GRUMPYY_06_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/greenness_Max100m_DNAmIC.csv", row.names = FALSE)

model <- lm(DNAmIC ~ GRUMPYY_07_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/greenness_Max250m_DNAmIC.csv", row.names = FALSE)

model <- lm(DNAmIC ~ GRUMPYY_08_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/greenness_Max500m_DNAmIC.csv", row.names = FALSE)

model <- lm(DNAmIC ~ GRUMPYY_09_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/greenness_Max1000m_DNAmIC.csv", row.names = FALSE)


##############
### nice table:

all_results <- bind_rows(
  read_model("regression_csvs/greenness_PostalCode_GrimAge.csv",  "Postal code") %>% mutate(Outcome = "GrimAge EAA"),
  read_model("regression_csvs/greenness_Mean100m_GrimAge.csv",    "Mean 100m")   %>% mutate(Outcome = "GrimAge EAA"),
  read_model("regression_csvs/greenness_Mean250m_GrimAge.csv",    "Mean 250m")   %>% mutate(Outcome = "GrimAge EAA"),
  read_model("regression_csvs/greenness_Mean500m_GrimAge.csv",    "Mean 500m")   %>% mutate(Outcome = "GrimAge EAA"),
  read_model("regression_csvs/greenness_Mean1000m_GrimAge.csv",   "Mean 1000m")  %>% mutate(Outcome = "GrimAge EAA"),
  read_model("regression_csvs/greenness_Max100m_GrimAge.csv",     "Max 100m")    %>% mutate(Outcome = "GrimAge EAA"),
  read_model("regression_csvs/greenness_Max250m_GrimAge.csv",     "Max 250m")    %>% mutate(Outcome = "GrimAge EAA"),
  read_model("regression_csvs/greenness_Max500m_GrimAge.csv",     "Max 500m")    %>% mutate(Outcome = "GrimAge EAA"),
  read_model("regression_csvs/greenness_Max1000m_GrimAge.csv",    "Max 1000m")   %>% mutate(Outcome = "GrimAge EAA"),

  read_model("regression_csvs/greenness_PostalCode_PhenoAge.csv", "Postal code") %>% mutate(Outcome = "PhenoAge EAA"),
  read_model("regression_csvs/greenness_Mean100m_PhenoAge.csv",   "Mean 100m")   %>% mutate(Outcome = "PhenoAge EAA"),
  read_model("regression_csvs/greenness_Mean250m_PhenoAge.csv",   "Mean 250m")   %>% mutate(Outcome = "PhenoAge EAA"),
  read_model("regression_csvs/greenness_Mean500m_PhenoAge.csv",   "Mean 500m")   %>% mutate(Outcome = "PhenoAge EAA"),
  read_model("regression_csvs/greenness_Mean1000m_PhenoAge.csv",  "Mean 1000m")  %>% mutate(Outcome = "PhenoAge EAA"),
  read_model("regression_csvs/greenness_Max100m_PhenoAge.csv",    "Max 100m")    %>% mutate(Outcome = "PhenoAge EAA"),
  read_model("regression_csvs/greenness_Max250m_PhenoAge.csv",    "Max 250m")    %>% mutate(Outcome = "PhenoAge EAA"),
  read_model("regression_csvs/greenness_Max500m_PhenoAge.csv",    "Max 500m")    %>% mutate(Outcome = "PhenoAge EAA"),
  read_model("regression_csvs/greenness_Max1000m_PhenoAge.csv",   "Max 1000m")   %>% mutate(Outcome = "PhenoAge EAA"),

  read_model("regression_csvs/greenness_PostalCode_DunedinPACE.csv", "Postal code") %>% mutate(Outcome = "DunedinPACE"),
  read_model("regression_csvs/greenness_Mean100m_DunedinPACE.csv",   "Mean 100m")   %>% mutate(Outcome = "DunedinPACE"),
  read_model("regression_csvs/greenness_Mean250m_DunedinPACE.csv",   "Mean 250m")   %>% mutate(Outcome = "DunedinPACE"),
  read_model("regression_csvs/greenness_Mean500m_DunedinPACE.csv",   "Mean 500m")   %>% mutate(Outcome = "DunedinPACE"),
  read_model("regression_csvs/greenness_Mean1000m_DunedinPACE.csv",  "Mean 1000m")  %>% mutate(Outcome = "DunedinPACE"),
  read_model("regression_csvs/greenness_Max100m_DunedinPACE.csv",    "Max 100m")    %>% mutate(Outcome = "DunedinPACE"),
  read_model("regression_csvs/greenness_Max250m_DunedinPACE.csv",    "Max 250m")    %>% mutate(Outcome = "DunedinPACE"),
  read_model("regression_csvs/greenness_Max500m_DunedinPACE.csv",    "Max 500m")    %>% mutate(Outcome = "DunedinPACE"),
  read_model("regression_csvs/greenness_Max1000m_DunedinPACE.csv",   "Max 1000m")   %>% mutate(Outcome = "DunedinPACE"),

  read_model("regression_csvs/greenness_PostalCode_DNAmIC.csv", "Postal code") %>% mutate(Outcome = "DNAmIC"),
  read_model("regression_csvs/greenness_Mean100m_DNAmIC.csv",   "Mean 100m")   %>% mutate(Outcome = "DNAmIC"),
  read_model("regression_csvs/greenness_Mean250m_DNAmIC.csv",   "Mean 250m")   %>% mutate(Outcome = "DNAmIC"),
  read_model("regression_csvs/greenness_Mean500m_DNAmIC.csv",   "Mean 500m")   %>% mutate(Outcome = "DNAmIC"),
  read_model("regression_csvs/greenness_Mean1000m_DNAmIC.csv",  "Mean 1000m")  %>% mutate(Outcome = "DNAmIC"),
  read_model("regression_csvs/greenness_Max100m_DNAmIC.csv",    "Max 100m")    %>% mutate(Outcome = "DNAmIC"),
  read_model("regression_csvs/greenness_Max250m_DNAmIC.csv",    "Max 250m")    %>% mutate(Outcome = "DNAmIC"),
  read_model("regression_csvs/greenness_Max500m_DNAmIC.csv",    "Max 500m")    %>% mutate(Outcome = "DNAmIC"),
  read_model("regression_csvs/greenness_Max1000m_DNAmIC.csv",   "Max 1000m")   %>% mutate(Outcome = "DNAmIC")
)

greenness_terms <- c("GRUMPYY_01_COM", "GRUMPYY_02_COM", "GRUMPYY_03_COM", "GRUMPYY_04_COM",
                     "GRUMPYY_05_COM", "GRUMPYY_06_COM", "GRUMPYY_07_COM", "GRUMPYY_08_COM",
                     "GRUMPYY_09_COM")

table_data <- all_results %>%
  filter(term %in% greenness_terms) %>%
  mutate(
    p.value = round(p.value, 3),
    Result = ifelse(
      Outcome == "DNAmIC",
      paste0(formatC(estimate, format = "e", digits = 2),
             " (", formatC(conf.low, format = "e", digits = 2),
             ", ", formatC(conf.high, format = "e", digits = 2), ")"),
      ifelse(
        abs(estimate) < 0.01,
        paste0(round(estimate, 4), " (", round(conf.low, 4), ", ", round(conf.high, 4), ")"),
        paste0(round(estimate, 2), " (", round(conf.low, 2), ", ", round(conf.high, 2), ")")
      )
    ),
    Result = ifelse(p.value < 0.05, paste0("**", Result, "**"), Result),
    Result_with_p = paste0(Result, " p=", p.value)
  ) %>%
  select(Env, Outcome, Result_with_p) %>%
  pivot_wider(names_from = Outcome, values_from = Result_with_p)

write.csv(table_data, "regression_csvs/greenness_aging_results.csv", row.names = FALSE)

save_kable(
  knitr::kable(table_data,
               col.names = c("Buffer", "GrimAge EAA", "PhenoAge EAA", "DunedinPACE", "DNAmIC"),
               caption = "Table X: Association between vegetative greenness and epigenetic aging",
               escape = TRUE) %>%
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE),
  "Figures/greenness_aging_table.png", zoom = 2
)

```



### condensation predicting aging 

 #condensation 
 "ENV_HMPRB_CON_MCQ",
 
 #NDVI
 "GRLANYY_26_COM", 

```{r}

#condensation predicting aging 
table(subset_data$ENV_HMPRB_CON_MCQ)
levels(subset_data$ENV_HMPRB_CON_MCQ)

# GrimAge EAA
model <- lm(GrimAge_EAA ~ ENV_HMPRB_CON_MCQ + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Condensation_GrimAge.csv", row.names = FALSE)

# PhenoAge EAA
model <- lm(PhenoAge_EAA ~ ENV_HMPRB_CON_MCQ + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Condensation_PhenoAge.csv", row.names = FALSE)

# DunedinPACE
model <- lm(DunedinPACE ~ ENV_HMPRB_CON_MCQ + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Condensation_DunedinPACE.csv", row.names = FALSE)

# DNAmIC
model <- lm(DNAmIC ~ ENV_HMPRB_CON_MCQ + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/Condensation_DNAmIC.csv", row.names = FALSE)



##############
### nice table:

all_results <- bind_rows(
  read_model("regression_csvs/Condensation_GrimAge.csv",     "Condensation") %>% mutate(Outcome = "GrimAge EAA"),
  read_model("regression_csvs/Condensation_PhenoAge.csv",    "Condensation") %>% mutate(Outcome = "PhenoAge EAA"),
  read_model("regression_csvs/Condensation_DunedinPACE.csv", "Condensation") %>% mutate(Outcome = "DunedinPACE"),
  read_model("regression_csvs/Condensation_DNAmIC.csv",      "Condensation") %>% mutate(Outcome = "DNAmIC")
)

condensation_terms <- c("ENV_HMPRB_CON_MCQNo")

table_data <- all_results %>%
  filter(term %in% condensation_terms) %>%
  mutate(
    p.value = round(p.value, 3),
    Result = ifelse(
      Outcome == "DNAmIC",
      paste0(formatC(estimate, format = "e", digits = 2),
             " (", formatC(conf.low, format = "e", digits = 2),
             ", ", formatC(conf.high, format = "e", digits = 2), ")"),
      ifelse(
        abs(estimate) < 0.01,
        paste0(round(estimate, 4), " (", round(conf.low, 4), ", ", round(conf.high, 4), ")"),
        paste0(round(estimate, 2), " (", round(conf.low, 2), ", ", round(conf.high, 2), ")")
      )
    ),
    Result = ifelse(p.value < 0.05, paste0("**", Result, "**"), Result),
    Result_with_p = paste0(Result, ", p=", p.value)
  ) %>%
  select(Env, Outcome, Result_with_p) %>%
  pivot_wider(names_from = Outcome, values_from = Result_with_p)

write.csv(table_data, "regression_csvs/condensation_aging_results.csv", row.names = FALSE)

save_kable(
  knitr::kable(table_data,
               col.names = c("Exposure", "GrimAge EAA", "PhenoAge EAA", "DunedinPACE", "DNAmIC"),
               caption = "Table X: Association between housing condensation and epigenetic aging",
               escape = TRUE) %>%
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE),
  "Figures/condensation_aging_table.png", zoom = 2
)
```


### NDVI predicting aging

```{r}
# GrimAge EAA
model <- lm(GrimAge_EAA ~ GRLANYY_26_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/NDVI_GrimAge.csv", row.names = FALSE)

# PhenoAge EAA
model <- lm(PhenoAge_EAA ~ GRLANYY_26_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/NDVI_PhenoAge.csv", row.names = FALSE)

# DunedinPACE
model <- lm(DunedinPACE ~ GRLANYY_26_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/NDVI_DunedinPACE.csv", row.names = FALSE)

# DNAmIC
model <- lm(DNAmIC ~ GRLANYY_26_COM + SEX_ASK_COM + AGE_NMBR_COM +
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "regression_csvs/NDVI_DNAmIC.csv", row.names = FALSE)

##############
### nice table:

all_results <- bind_rows(
  read_model("regression_csvs/NDVI_GrimAge.csv",     "NDVI (Max 500m)") %>% mutate(Outcome = "GrimAge EAA"),
  read_model("regression_csvs/NDVI_PhenoAge.csv",    "NDVI (Max 500m)") %>% mutate(Outcome = "PhenoAge EAA"),
  read_model("regression_csvs/NDVI_DunedinPACE.csv", "NDVI (Max 500m)") %>% mutate(Outcome = "DunedinPACE"),
  read_model("regression_csvs/NDVI_DNAmIC.csv",      "NDVI (Max 500m)") %>% mutate(Outcome = "DNAmIC")
)

ndvi_terms <- c("GRLANYY_26_COM")

table_data <- all_results %>%
  filter(term %in% ndvi_terms) %>%
  mutate(
    p.value = round(p.value, 3),
    Result = ifelse(
      Outcome == "DNAmIC",
      paste0(formatC(estimate, format = "e", digits = 2),
             " (", formatC(conf.low, format = "e", digits = 2),
             ", ", formatC(conf.high, format = "e", digits = 2), ")"),
      ifelse(
        abs(estimate) < 0.01,
        paste0(round(estimate, 4), " (", round(conf.low, 4), ", ", round(conf.high, 4), ")"),
        paste0(round(estimate, 2), " (", round(conf.low, 2), ", ", round(conf.high, 2), ")")
      )
    ),
    Result = ifelse(p.value < 0.05, paste0("**", Result, "**"), Result),
    Result_with_p = paste0(Result, ", p=", p.value)
  ) %>%
  select(Env, Outcome, Result_with_p) %>%
  pivot_wider(names_from = Outcome, values_from = Result_with_p)

write.csv(table_data, "regression_csvs/NDVI_aging_results.csv", row.names = FALSE)

save_kable(
  knitr::kable(table_data,
               col.names = c("Exposure", "GrimAge EAA", "PhenoAge EAA", "DunedinPACE", "DNAmIC"),
               caption = "Table X: Association between NDVI and epigenetic aging",
               escape = TRUE) %>%
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE),
  "Figures/NDVI_aging_table.png", zoom = 2
)
```





























### 6.3.1 Enviro/aging with other air pollution variables 

```{r}


PM25DALYY_01_COM
SO2OMIYY_COM
O3CHGYY_01_COM


#pm2.5
model <- lm(GrimAge_EAA ~ PM25DALYY_01_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "pm25_grimage.csv", row.names = FALSE)

model <- lm(PhenoAge_EAA ~ PM25DALYY_01_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "pm25_phenoage.csv", row.names = FALSE)

model <- lm(DunedinPACE ~ PM25DALYY_01_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "pm25_Duned.csv", row.names = FALSE)

model <- lm(DNAmIC ~ PM25DALYY_01_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "pm25_IC.csv", row.names = FALSE)





#so2_1yr_com
model <- lm(GrimAge_EAA ~ SO2OMIYY_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "pm25_grimage.csv", row.names = FALSE)


model <- lm(PhenoAge_EAA ~ SO2OMIYY_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "pm25_phenoage.csv", row.names = FALSE)


model <- lm(DunedinPACE ~ SO2OMIYY_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "pm25_Duned.csv", row.names = FALSE)


model <- lm(DNAmIC ~ SO2OMIYY_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "pm25_IC.csv", row.names = FALSE)




#ozone
model <- lm(GrimAge_EAA ~ O3CHGYY_01_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "pm25_grimage.csv", row.names = FALSE)


model <- lm(PhenoAge_EAA ~ O3CHGYY_01_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "pm25_phenoage.csv", row.names = FALSE)


model <- lm(DunedinPACE ~ O3CHGYY_01_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "pm25_Duned.csv", row.names = FALSE)


model <- lm(DNAmIC ~ O3CHGYY_01_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "pm25_IC.csv", row.names = FALSE)


```






























# 7 To do 
-median dot plots for NO2, SO2, Ozone 
-visualizations of tree canopy and veg. 


# 8 plots for presentation 


```{r}
air_income_plots_pres <- list()

for (v in air_income_vars) {
  p <- subset_data %>%
    filter(income_cat %in% c("<20K", "20K–50K", "50K–100K", "100K–150K", "150K+")) %>%
    ggplot(aes(x = income_cat, y = .data[[v$col]], fill = income_cat)) +
    geom_boxplot(outlier.alpha = 0.3, width = 0.6) +
    scale_fill_brewer(palette = "Blues") +
    labs(
      title = v$title,
      x = "Household Income",
      y = v$y_lab
    ) +
    report_theme +
    report_theme +
theme(
  plot.title      = element_text(size = 22, face = "bold"),
  axis.title.x    = element_text(size = 20, face = "bold", margin = margin(t = 6)),
  axis.title.y    = element_text(size = 20, face = "bold", margin = margin(r = 6)),
  axis.text.x     = element_text(size = 18),
  axis.text.y     = element_text(size = 18),
  legend.position = "none",
  plot.margin      = margin(t = 15, r = 15, b = 15, l = 15),
  plot.background  = element_rect(color = "grey60", linewidth = 1, fill = "white"),
  panel.background = element_rect(fill = "white")
)
  
  
  
  
  air_income_plots_pres[[v$col]] <- p
}

dashboard_air_income_pres <- wrap_plots(air_income_plots_pres, ncol = 2) +
  plot_annotation(
    title = "Air Pollution by Household Income",
    theme = theme(
      plot.title      = element_text(hjust = 0.5, face = "bold", size = 28),
      plot.margin     = margin(t = 20, r = 10, b = 10, l = 10),
      plot.background = element_rect(fill = "grey90", color = NA)
    )
  )


ggsave("Figures/air_pollution_income_dashboard_PRESENTATION.png",
       plot = dashboard_air_income_pres, width = 18, height = 14, dpi = 300)





```






#9 Imputation 











# Not using yet 



## Table1 of All variables (Summary Stats)
```{r}

# Table1 all the variables (label and make into factor first)  -----------------------------------------------

#demographic variables age, sex, race 
subset_data$AGE_NMBR_COM <- as.numeric(subset_data$AGE_NMBR_COM) 
subset_data$SEX_ASK_COM <- factor(subset_data$SEX_ASK_COM) #F is first 
subset_data$WGHTS_PROV_COM <- factor(subset_data$WGHTS_PROV_COM) #AB is first 
subset_data$Race7 <- factor(subset_data$Race7,
                            levels = c("White", "South Asian", "Black", "Asian",
                                       "Arab", "Other")) #White is reference

label(subset_data$SEX_ASK_COM) <- "Sex (M/F)"#label categorical variables 
label(subset_data$AGE_NMBR_COM) <- "Age (years)"
label(subset_data$Race7) <- "Cultural background"

table1(~ SEX_ASK_COM + AGE_NMBR_COM + Race7, data = subset_data)

#SEC variables 
table(subset_data$INC_TOT_COM)
subset_data$INC_TOT_COM_factored <- factor(subset_data$INC_TOT_COM,#household income
                                                levels = c(1, 2, 3, 4, 5),
                                                labels = c("150K+",
                                                      "100K to 150K",
                                                      "50K–100K",
                                                      "20K-50K",
                                                      "<20K"))
                           

table(subset_data$ED_HIGH_COM)
subset_data$ED_HIGH_COM_factored <- factor(subset_data$ED_HIGH_COM,#education
                                                levels = c(1, 2, 3, 4, 5, 6),
                                                labels = c("01 = Above bachelor’s degg",
                                                           "02 = Bachelor’s deg",
                                                           "03 = Uni cert below bachelor’s",
                                                           "04 = Non-univ college diplomas",
                                                           "05 = Trade cert/diploma",
                                                           "06 = Above bachelor’s deg"))


table(subset_data$WEA_SVNGSVL_MCQ)
subset_data$WEA_SVNGSVL_MCQ_factored <- factor(subset_data$WEA_SVNGSVL_MCQ,#savings
                                                    levels = c(1, 2, 3, 4, 5),
                                                    labels = c("$1 million +", 
                                                               "100K to <$1 million",
                                                               "50K–100K",
                                                               "<50K",
                                                               "Don't know/No answer/Refused"
                                                               ))



label(subset_data$INC_TOT_COM_factored)          <- "Total Household Income (past 12 months)"
label(subset_data$ED_HIGH_COM_factored)           <- "Highest Education Obtained"
label(subset_data$WEA_SVNGSVL_MCQ_factored)       <- "Total Savings and Investments"

table1(~ INC_TOT_COM_factored + ED_HIGH_COM_factored + WEA_SVNGSVL_MCQ_factored, data = subset_data)



#environmental variables 

table(subset_data$ENV_HMPRB_CON_MCQ)
subset_data$ENV_HMPRB_CON_MCQ_factored <- factor(subset_data$ENV_HMPRB_CON_MCQ,#condensation
                                                      levels = c(0, 1),
                                                      labels = c("No", "Yes"))

table(subset_data$ENV_HMPRB_NOI_MCQ)
subset_data$ENV_HMPRB_NOI_MCQ_factored <- factor(subset_data$ENV_HMPRB_NOI_MCQ,#noise
                                                      levels = c(0, 1),
                                                      labels = c("No", "Yes"))



label(subset_data$ENV_HMPRB_CON_MCQ_factored)     <- "Home Problems: Condensation"
label(subset_data$ENV_HMPRB_NOI_MCQ_factored)     <- "Home Problems: Noise"
label(subset_data$pm25_1yr_com)          <- "PM2.5 (1-year average)"
label(subset_data$no2_1yr_com)          <- "NO2 (1-year average)"
label(subset_data$so2_1yr_com)          <- "SO2 (1-year average)"
label(subset_data$o3_1yr_com)           <- "Ozone (1-year average)"
label(subset_data$GRLANYY_26_COM)       <- "Greenness (Maximum of annual mean NDVI within 250m)"
label(subset_data$temp_1yr_com)         <- "Temperature (1-year average)"
label(subset_data$LGTNLTYY_01_COM)      <- "Nighttime Light Brightness"
label(subset_data$Locrds_length_200m_com)  <- "Length of local rds within 200m of postal code centre"
label(subset_data$Phwy_length_200m_com) <- "Length of hwy within 200m of postal code centre" 

table1(~ ENV_HMPRB_CON_MCQ_factored + ENV_HMPRB_NOI_MCQ_factored + pm25_1yr_com + no2_1yr_com + so2_1yr_com 
       + o3_1yr_com +  GRLANYY_26_COM + temp_1yr_com + LGTNLTYY_01_COM + 
         Locrds_length_200m_com + Phwy_length_200m_com, data = subset_data)



label(subset_data$grtcc10_01_COM)  <- "Mean % of pixel covered by vegetation >5m"
label(subset_data$grtcc10_02_COM)  <- "Mean % of pixel covered by vegetation >5m within 100m"
label(subset_data$grtcc10_03_COM)  <- "Mean % of pixel covered by vegetation >5m within 250m"
label(subset_data$grtcc10_04_COM)  <- "Mean % of pixel covered by vegetation >5m within 500m"
label(subset_data$grtcc10_05_COM)  <- "Mean % of pixel covered by vegetation >5m within 1km"

label(subset_data$GRUMPYY_01_COM)  <- "Annual % vegetative greenness at postal code"
label(subset_data$GRUMPYY_02_COM)  <- "Mean of annual % of vegetative greenness within 100m"
label(subset_data$GRUMPYY_03_COM)  <- "Mean of annual % of vegetative greenness within 250m" 
label(subset_data$GRUMPYY_04_COM)  <- "Mean of annual % of vegetative greenness within 500m"
label(subset_data$GRUMPYY_05_COM)  <- "Mean of annual % of vegetative greenness within 1km"
label(subset_data$GRUMPYY_06_COM)  <- "Max of annual % of vegetative greenness within 100m" 
label(subset_data$GRUMPYY_07_COM)  <- "Max of annual % of vegetative greenness within 250m" 
label(subset_data$GRUMPYY_08_COM)  <- "Max of annual % of vegetative greenness within 500m" 
label(subset_data$GRUMPYY_09_COM)  <- "Max of annual % of vegetative greenness within 1km" 



table1(~ grtcc10_01_COM + grtcc10_02_COM + grtcc10_03_COM + grtcc10_04_COM + grtcc10_05_COM 
       + GRUMPYY_01_COM +  GRUMPYY_02_COM + GRUMPYY_03_COM + GRUMPYY_04_COM + 
         GRUMPYY_05_COM + GRUMPYY_06_COM + GRUMPYY_07_COM + GRUMPYY_08_COM + GRUMPYY_09_COM, data = subset_data)



#Aging measures (outcomes)

table1(~ GrimAge_EAA + PhenoAge_EAA + DunedinPACE + DNAmIC, data = subset_data)


```




## Covariates Smoking, Alcohol, PASE

```{r}

#PASE
str(subset_data$PA2_DSCR2_MCQ) #PASE is numeric variable 

#smoke
str(subset_data$ICQ_SMOKE_COM) #smoke is integer (though it is categorical...)
table(subset_data$ICQ_SMOKE_COM) #1=Yes (I currently smoke), 2=No (I don't smoke and I never have), 3=Former (I don't smoke now but I have in the past)
subset_data$ICQ_SMOKE_COM_RECODED <- dplyr::case_when( #recode to 1=No I don't smoke and never have. 2=former. 3= yes i currently smoke 
  subset_data$ICQ_SMOKE_COM == 2 ~ 1,
  subset_data$ICQ_SMOKE_COM == 3 ~ 2,
  subset_data$ICQ_SMOKE_COM == 1 ~ 3,
  TRUE ~ NA_real_   
)
table(subset_data$ICQ_SMOKE_COM_RECODED)
subset_data$ICQ_SMOKE_COM <- subset_data$ICQ_SMOKE_COM_RECODED
sum(is.na(subset_data$ICQ_SMOKE_COM)) #number of NAs is 229


#alcohol
str(subset_data$ALC_FREQ_COM) #alcohol is integer (though it is categorical...)
table(subset_data$ALC_FREQ_COM)  

01			Almost every day (incl. 6 times a week)
02			4-5 times a week
03			2-3 times a week
04			Once a week
05			2-3 times a month
06			About once a month
07			Less than once a month
96=never. 98=don't know/no answer 

subset_data$ALC_FREQ_COM_RECODED <- dplyr::case_when( #recode to 1=Least drinking. 
  subset_data$ALC_FREQ_COM == 7 ~ 1,
  subset_data$ALC_FREQ_COM == 6 ~ 2,
  subset_data$ALC_FREQ_COM == 5 ~ 3,
  subset_data$ALC_FREQ_COM == 4 ~ 4,
  subset_data$ALC_FREQ_COM == 3 ~ 5,
  subset_data$ALC_FREQ_COM == 2 ~ 6,
  subset_data$ALC_FREQ_COM == 1 ~ 7,
  subset_data$ALC_FREQ_COM == 96 ~ 8,
  subset_data$ALC_FREQ_COM == 98 ~ NA_real_,  
  TRUE ~ NA_real_
)
table(subset_data$ALC_FREQ_COM_RECODED)
subset_data$ALC_FREQ_COM <- subset_data$ALC_FREQ_COM_RECODED
sum(is.na(subset_data$ALC_FREQ_COM)) #number of NAs is 257


#make sure the variables are appropriate types 

vars_to_check <- c(
  "AGE_NMBR_COM", "SEX_ASK_COM", "WGHTS_PROV_COM", "Race7",
  
  # Socioeconomic
  "INC_TOT_COM", "ED_HIGH_COM", "WEA_SVNGSVL_MCQ",
  
  # Outcomes
  "GrimAge_EAA", "PhenoAge_EAA", "DunedinPACE", "DNAmIC",
  
  # Environmental exposures
  "pm25_1yr_com", "no2_1yr_com", "o3_1yr_com", "GRLANYY_26_COM", "temp_1yr_com", 
  "LGTNLTYY_01_COM", "ENV_HMPRB_CON_MCQ", "ENV_HMPRB_NOI_MCQ",   
  
  # Lifestyle / behavioral
  "PA2_DSCR2_MCQ", "ALC_FREQ_COM", "ICQ_SMOKE_COM"
)
sapply(subset_data[vars_to_check], class)

```


# Delete 

## Charts: age, sex, ethnicity, income, wealth  
```{r}

#sex and age 
barplot(table(subset_data$SEX_ASK_COM),
        main = "Sex Distribution",
        col = "lightblue",
        ylab = "Count",
        las = 2,
        cex.names = 0.8)


# age and sex  ------------------------------------------------------

#pie chart for sex
df_sex <- subset_data %>%
  filter(!is.na(SEX_ASK_COM)) %>%
  count(SEX_ASK_COM) %>%
  mutate(
    percent = n / sum(n),
    label = paste0(SEX_ASK_COM, " (", round(percent * 100, 1), "%)")
  )

ggplot(df_sex, aes(x = 1, y = n, fill = SEX_ASK_COM)) +
  geom_col(width = 1, color = "white", linewidth = 1) +
  coord_polar(theta = "y") +
  geom_text(
    aes(label = label),
    position = position_stack(vjust = 0.5),
    size = 10,              # big, slide-friendly
    fontface = "plain"
  ) +
  scale_fill_manual(values = c("F" = "#1f77b4", "M" = "#ff7f0e")) +
  labs(title = "Sex Distribution", x = NULL, y = NULL, fill = NULL) +
  theme_void() +
  theme(
    # remove any remaining axis/radial artifacts
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    
    plot.title = element_text(
      hjust = 0.5,
      face  = "bold",
      size  = 38,
      margin = margin(b = 18)
    ),
    
    legend.position = "right",
    legend.text = element_text(size = 26),
    legend.key.size = grid::unit(1.2, "lines")
  )



#age boxplot 
subset_data %>%
  filter(!is.na(AGE_NMBR_COM)) %>%
  ggplot(aes(x = "", y = AGE_NMBR_COM)) +
  geom_boxplot(
    width = 0.25,
    fill = "white",
    color = "#2C7FB8",
    linewidth = 1.2
  ) +
  scale_y_continuous(
    breaks = seq(
      floor(min(subset_data$AGE_NMBR_COM, na.rm = TRUE) / 5) * 5,
      ceiling(max(subset_data$AGE_NMBR_COM, na.rm = TRUE) / 5) * 5,
      by = 5
    )
  ) +
  labs(
    title = "Age Distribution",
    x = NULL,
    y = "Age (years)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(
      hjust  = 0.5,
      face   = "bold",
      size   = 38,
      margin = margin(b = 18)
    ),
    axis.title.y = element_text(
      face   = "bold",
      size   = 30,
      margin = margin(r = 16)
    ),
    axis.title.x = element_blank(),
    axis.text.y  = element_text(size = 26),
    axis.text.x  = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_line(linetype = "dotted"),
    panel.grid.major.y = element_line(linetype = "dashed"),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background  = element_rect(fill = "white", color = NA)
  )



# ethnicity, income, wealth ---------------------------------------------------
#cultural background
#horizontal bar plots

subset_data %>%
  filter(!is.na(Race7)) %>%
  count(Race7, name = "n") %>%
  arrange(n) %>%
  ggplot(aes(x = n, y = Race7)) +
  geom_col(fill = "#2C7FB8", width = 0.7) +
  labs(
    title = "Self-Reported Cultural Background",
    x = "Number of participants",
    y = NULL
  ) +
  theme_classic(base_size = 26) +   # ⬅️ bigger global base
  theme(
    legend.position = "none",
    
    # TITLE — BIG
    plot.title = element_text(
      hjust = 0.5,
      face  = "bold",
      size  = 38,          # ⬅️ increased
      margin = margin(b = 18)
    ),
    
    # AXIS TITLES — BIGGER
    axis.title.x = element_text(
      face = "bold",
      size = 30,          # ⬅️ increased
      margin = margin(t = 16)
    ),
    axis.title.y = element_blank(),
    
    # AXIS TICK LABELS — BIGGER
    axis.text.x = element_text(size = 26),   # ⬅️ increased
    axis.text.y = element_text(size = 26)
  )




#horizontal bar chart of household income categories 
subset_data %>%
  mutate(
    income_cat = factor(
      INC_TOT_COM,  # personal/household income
      levels = c(1, 2, 3, 4, 5, 7, 8, 9),
      labels = c(
        "<20K",
        "20K–50K",
        "50K–100K",
        "100K–150K",
        "150K+",
        "Missing",
        "Don't know/No answer",
        "Refused"
      )
    )
  ) %>%
  count(income_cat, name = "n") %>%
  arrange(n) %>%
  ggplot(aes(x = n, y = income_cat)) +
  geom_col(fill = "#2C7FB8", width = 0.7) +
  labs(
    title = "Household Income (Past 12 Months)",
    x = "Number of participants",
    y = NULL
  ) +
  theme_classic(base_size = 26) +
  theme(
    legend.position = "none",
    
    # TITLE
    plot.title = element_text(
      hjust = 0.5,
      face  = "bold",
      size  = 38,
      margin = margin(b = 18)
    ),
    
    # AXIS TITLES
    axis.title.x = element_text(
      face = "bold",
      size = 30,
      margin = margin(t = 16)
    ),
    axis.title.y = element_blank(),
    
    # AXIS TICK LABELS
    axis.text.x = element_text(size = 26),
    axis.text.y = element_text(size = 26)
  )




#horizontal bar chart of wealth/savings   
subset_data %>%
  mutate(
    wealth_cat = factor(
      WEA_SVNGSVL_MCQ,  
      levels = c(1, 2, 3, 4, 8, 9),
      labels = c(
        "<50K",
        "50K–100K",
        "100K to <$1 million",
        "$1 million +",
        "Don't know/No answer",
        "Refused"
      )
    )
  ) %>%
  count(wealth_cat, name = "n") %>%
  arrange(n) %>%
  ggplot(aes(x = n, y = wealth_cat)) +
  geom_col(fill = "#2C7FB8", width = 0.7) +
  labs(
    title = "Total Savings and Investments",
    x = "Number of participants",
    y = NULL
  ) +
  theme_classic(base_size = 26) +
  theme(
    legend.position = "none",
    plot.title = element_text(
      hjust = 0.5, face = "bold", size = 38,
      margin = margin(b = 18)
    ),
    axis.title.x = element_text(
      face = "bold", size = 30,
      margin = margin(t = 16)
    ),
    axis.title.y = element_blank(),
    axis.text.x  = element_text(size = 26),
    axis.text.y  = element_text(size = 26)
  )



```


```{r}

#savings 
table(subset_data$WEA_SVNGSVL_MCQ, useNA = "ifany") #categories are #1 = Less than $50,000; 2 = $50,000–$99,999; 3 = $100,000–$999,999; 4 = $1 million or more; 8 = Don’t know/No answer; 9 = Refused
#there are 103 NAs to start with. 

subset_data$WEA_SVNGSVL_MCQ_RECODED <- dplyr::case_when( #reverse code so that 1=4, 2=3, 3=2, 4=1, 5=8,9 
  subset_data$WEA_SVNGSVL_MCQ == 1 ~ 4,
  subset_data$WEA_SVNGSVL_MCQ == 2 ~ 3,
  subset_data$WEA_SVNGSVL_MCQ == 3 ~ 2,
  subset_data$WEA_SVNGSVL_MCQ == 4 ~ 1,
  subset_data$WEA_SVNGSVL_MCQ %in% c(8, 9) ~ 5,
  TRUE ~ NA_real_   
)
subset_data$WEA_SVNGSVL_MCQ <- subset_data$WEA_SVNGSVL_MCQ_RECODED
table(subset_data$WEA_SVNGSVL_MCQ, useNA = "ifany")#103 NAs 


#education 
# Step 1: Coerce to numeric safely (from character/factor labels)
# Step 2: Set invalid categories to NA (97 = other, 98 = don't know, 99 = refused)
# Step 3: Reverse code only valid numeric values (1–6)
# Step 4: Convert to factor
# Step 5: Optional – drop unused factor levels (e.g., if -90 or others existed)
# Step 6: Check
#check NAs

table(subset_data$ED_HIGH_COM)
subset_data$ED_HIGH_COM <- as.numeric(as.character(subset_data$ED_HIGH_COM))
subset_data$ED_HIGH_COM[subset_data$ED_HIGH_COM %in% c(97, 98, 99)] <- NA
subset_data$ED_HIGH_COM <- ifelse(!is.na(subset_data$ED_HIGH_COM), 7 - subset_data$ED_HIGH_COM, NA)
subset_data$ED_HIGH_COM <- factor(subset_data$ED_HIGH_COM)
subset_data$ED_HIGH_COM <- droplevels(subset_data$ED_HIGH_COM)
table(subset_data$ED_HIGH_COM, useNA = "ifany") #231 NAs 

saveRDS(subset_data, file = "subset_data.rds")
```

## Length of roads within 200m 

```{r}
vars_roads <- list(
  list(col = "Phwy_length_200m_com",   label = "Length of highways within 200m of postal code (m)"),
  list(col = "Locrds_length_200m_com", label = "Length of local roads within 200m of postal code (m)")
)

summary_table_roads <- lapply(vars_roads, function(v) {
  x <- subset_data[[v$col]]
  n_zero_or_missing <- sum(x == 0 | is.na(x), na.rm = TRUE)
  pct_zero_or_missing <- round(n_zero_or_missing / length(x) * 100, 1)
  
  x_filtered <- x[x > 0 & !is.na(x)]  # remove zeros and NAs for stats
  
  tibble(
    Variable                = v$label,
    `Mean (SD)`             = paste0(round(mean(x_filtered), 2), " (", round(sd(x_filtered), 2), ")"),
    `Median [Min, Max]`     = paste0(round(median(x_filtered), 2), " [", round(min(x_filtered), 2), ", ", round(max(x_filtered), 2), "]"),
    `Zero or Missing n (%)` = paste0(n_zero_or_missing, " (", pct_zero_or_missing, "%)")
  )
}) %>% bind_rows()

ft_roads <- flextable(summary_table_roads) %>%
  add_header_lines("Table X. Proximity to Roads") %>%
  bold(part = "header") %>%
  bold(j = "Variable") %>%
  autofit() %>%
  theme_booktabs() %>%
  fontsize(size = 10, part = "all") %>%
  fontsize(i = 1, size = 11, part = "header") %>%
  bold(i = 1, part = "header") %>%
  font(fontname = "Times New Roman", part = "all") %>%
  align(i = 1, align = "center", part = "header") %>%
  align(align = "left", part = "all") %>%
  align(j = c("Mean (SD)", "Median [Min, Max]", "Zero or Missing n (%)"), align = "center", part = "body")

save_as_image(ft_roads, path = "Tables/Table_roads.png", res = 300)
```


```{r}
summary(subset_data$Phwy_length_200m_com) #Length of hwy within 200m of postal code centre
quantile(subset_data$Phwy_length_200m_com,
         probs = c(0, .01, .05, .10, .25, .50, .75, .90, .95, .99),
         na.rm = TRUE)
mean(subset_data$Phwy_length_200m_com == 0, na.rm = TRUE) #96% of the participants (1423 participants) have no hwy nearby. 



summary(subset_data$Locrds_length_200m_com) #Length of local rds within 200m of postal code centre
quantile(subset_data$Locrds_length_200m_com,
         probs = c(0, .01, .05, .10, .25, .50, .75, .90, .95, .99),
         na.rm = TRUE) 

hist(subset_data$Locrds_length_200m_com,
     breaks = 50,
     main = "Length of local rds within 200 m",
     xlab = "Meters",
     col = "grey")


```




