---
title: "Visualizations"
output:
  word_document: default
  pdf_document: default
  html_document: default
date: "2026-04-10"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,       # hides code
  message = FALSE,    # suppresses messages
  warning = FALSE     # suppresses warnings
)
```

# 1. Load and clean data 

```{r}

#load data 

load("data/CLSA_methylation.rds")

library(dplyr)
library(ggplot2)
library(tidyverse)
library(flextable)
library (haven)
library(table1)
library(stargazer)
library(htmltools)

library(table1)
library(Hmisc)
library(htmltools)
library(webshot2)
library(tidyverse)


#not going to include these: 
#"ED_HIGH_COM", "WEA_SVNGSVL_MCQ","temp_1yr_com" LGTNLTYY_01_COM "ENV_HMPRB_NOI_MCQ","pm25_5yrs_com",   "PM25DALYY_02_COM", "no2_5yrs_com", 
 # "Locrds_length_200m_com",   "Phwy_length_200m_com", 

#create a subset 
subset_data <- CLSA_methylation[, c(
  "entity_id.x", 
  
  "AGE_NMBR_COM", "SEX_ASK_COM", "WGHTS_PROV_COM", "Race7",
  
  # Socioeconomic
  "INC_TOT_COM", "WEA_SVNGSVL_MCQ",
  
  # Outcomes
  "GrimAge_EAA", "PhenoAge_EAA", "DunedinPACE", "DNAmIC",
  
  # Environmental exposures
  "pm25_1yr_com", "no2_1yr_com", "so2_1yr_com", "o3_1yr_com", "GRLANYY_26_COM",
    "ENV_HMPRB_CON_MCQ",
  
  #CANUE Linked environmental variables 
  
  "PM25DALYY_01_COM",  "O3CHGYY_01_COM",  "SO2OMIYY_COM", 
  
  
  "grtcc10_01_COM",   "grtcc10_02_COM",   "grtcc10_03_COM",   "grtcc10_04_COM",  "grtcc10_05_COM",  #tree canopy coverage 
  
  "GRUMPYY_01_COM",  "GRUMPYY_02_COM", "GRUMPYY_03_COM",   "GRUMPYY_04_COM",   "GRUMPYY_05_COM",   "GRUMPYY_06_COM",   "GRUMPYY_07_COM",
  "GRUMPYY_08_COM",   "GRUMPYY_09_COM", #%vegetative greenness 
  
  
  
  # Lifestyle / behavioral
  "PA2_DSCR2_MCQ", "ALC_FREQ_COM", "ICQ_SMOKE_COM", #PASE scale: PASE Score - Final (in sharepoint dictionary)
  
  #cell types
  "CD4nv", "CD8nv",   "Mono", "Treg", "CD4mem", "CD8mem", "NK", "Bmem", "Bnv"
  
)]


```




## Recode missing values to N/A

```{r}
# Recode first
subset_data$pm25_1yr_com[subset_data$pm25_1yr_com %in% c(-8888, -9999)] <- NA
subset_data$no2_1yr_com[subset_data$no2_1yr_com %in% c(-8888, -9999)] <- NA
subset_data$so2_1yr_com[subset_data$so2_1yr_com %in% c(-8888, -9999)] <- NA
subset_data$o3_1yr_com[subset_data$o3_1yr_com %in% c(-8888, -9999)] <- NA
subset_data$GRLANYY_26_COM[subset_data$GRLANYY_26_COM %in% c(-8888, -9999)] <- NA
subset_data$ENV_HMPRB_CON_MCQ[subset_data$ENV_HMPRB_CON_MCQ %in% c(-8888, -9999)] <- NA
subset_data$PM25DALYY_01_COM[subset_data$PM25DALYY_01_COM %in% c(-8888, -9999)] <- NA
subset_data$O3CHGYY_01_COM[subset_data$O3CHGYY_01_COM %in% c(-8888, -9999)] <- NA
subset_data$SO2OMIYY_COM[subset_data$SO2OMIYY_COM %in% c(-8888, -9999)] <- NA
subset_data$grtcc10_01_COM[subset_data$grtcc10_01_COM %in% c(-8888, -9999)] <- NA
subset_data$grtcc10_02_COM[subset_data$grtcc10_02_COM %in% c(-8888, -9999)] <- NA
subset_data$grtcc10_03_COM[subset_data$grtcc10_03_COM %in% c(-8888, -9999)] <- NA
subset_data$grtcc10_04_COM[subset_data$grtcc10_04_COM %in% c(-8888, -9999)] <- NA
subset_data$grtcc10_05_COM[subset_data$grtcc10_05_COM %in% c(-8888, -9999)] <- NA
subset_data$GRUMPYY_01_COM[subset_data$GRUMPYY_01_COM %in% c(-8888, -9999)] <- NA
subset_data$GRUMPYY_02_COM[subset_data$GRUMPYY_02_COM %in% c(-8888, -9999)] <- NA
subset_data$GRUMPYY_03_COM[subset_data$GRUMPYY_03_COM %in% c(-8888, -9999)] <- NA
subset_data$GRUMPYY_04_COM[subset_data$GRUMPYY_04_COM %in% c(-8888, -9999)] <- NA
subset_data$GRUMPYY_05_COM[subset_data$GRUMPYY_05_COM %in% c(-8888, -9999)] <- NA
subset_data$GRUMPYY_06_COM[subset_data$GRUMPYY_06_COM %in% c(-8888, -9999)] <- NA
subset_data$GRUMPYY_07_COM[subset_data$GRUMPYY_07_COM %in% c(-8888, -9999)] <- NA
subset_data$GRUMPYY_08_COM[subset_data$GRUMPYY_08_COM %in% c(-8888, -9999)] <- NA
subset_data$GRUMPYY_09_COM[subset_data$GRUMPYY_09_COM %in% c(-8888, -9999)] <- NA

# Then check — should all be 0
sapply(subset_data, function(x) sum(x %in% c(-8888, -9999), na.rm = TRUE))

colSums(is.na(subset_data))#check number of NAs






#show how many NAs are in each variable (in a table)
na_summary <- data.frame(
  Variable = names(subset_data),
  N_Missing = colSums(is.na(subset_data)),
  Pct_Missing = round(colSums(is.na(subset_data)) / nrow(subset_data) * 100, 1)
) %>%
  filter(N_Missing > 0) %>%  # only show variables that have missing data
  arrange(desc(N_Missing))

ft_na <- flextable(na_summary) %>%
  set_header_labels(
    Variable = "Variable",
    N_Missing = "N Missing",
    Pct_Missing = "% Missing"
  ) %>%
  add_header_lines("Table X. Missing Data Summary") %>%
  bold(part = "header") %>%
  bold(j = "Variable") %>%
  autofit() %>%
  theme_booktabs() %>%
  fontsize(size = 10, part = "all") %>%
  fontsize(i = 1, size = 11, part = "header") %>%
  bold(i = 1, part = "header") %>%
  font(fontname = "Times New Roman", part = "all") %>%
  align(i = 1, align = "center", part = "header") %>%
  align(align = "left", part = "all") %>%
  align(j = c("N_Missing", "Pct_Missing"), align = "center", part = "body")

ft_na
save_as_image(ft_na, path = "Tables/Table_missing.png", res = 300)

```





## Reverse code Household Income

```{r}
# BEFORE reversing
cat("=== BEFORE reverse coding ===\n")
table(subset_data$INC_TOT_COM, useNA = "ifany")
# 1 = Less than $20,000; 2 = $20,000–$49,999; 3 = $50,000–$99,999
# 4 = $100,000–$149,999; 5 = $150,000 or more; 8 = Don't know/No answer; 9 = Refused

subset_data$INC_TOT_COM_RECODED <- dplyr::case_when(
  subset_data$INC_TOT_COM == 1 ~ 5,
  subset_data$INC_TOT_COM == 2 ~ 4,
  subset_data$INC_TOT_COM == 3 ~ 3,
  subset_data$INC_TOT_COM == 4 ~ 2,
  subset_data$INC_TOT_COM == 5 ~ 1,
  subset_data$INC_TOT_COM %in% c(8, 9) ~ 6,
  TRUE ~ NA_real_
)

# AFTER reversing
cat("\n=== AFTER reverse coding ===\n")
table(subset_data$INC_TOT_COM_RECODED, useNA = "ifany")
# 1 = $150,000+; 2 = $100,000–$149,999; 3 = $50,000–$99,999
# 4 = $20,000–$49,999; 5 = Less than $20,000; 6 = Don't know/No answer/Refused

subset_data$INC_TOT_COM <- subset_data$INC_TOT_COM_RECODED
cat("\nNAs in INC_TOT_COM after recoding:", sum(is.na(subset_data$INC_TOT_COM)), "\n")
subset_data <- subset_data %>% select(-INC_TOT_COM_RECODED) 

# Create income_cat factor
subset_data <- subset_data %>%
  mutate(
    income_cat = factor(
      INC_TOT_COM,
      levels = c(1, 2, 3, 4, 5, 6),
      labels = c("150K+", "100K–150K", "50K–100K", "20K–50K", "<20K", "Don't know/No answer/Refused")
    )
  )

# Final check
cat("\n=== income_cat factor (final) ===\n")
table(subset_data$income_cat, useNA = "ifany")


```


WEA_SVNGSVL_MCQ

## Reverse code Savings

```{r}

# BEFORE reversing
cat("=== BEFORE reverse coding ===\n")
table(subset_data$WEA_SVNGSVL_MCQ, useNA = "ifany")
# 1 = Less than $50,000; 2 = $50,000–$99,999; 3 = $100,000–<$1 million
# 4 = $1 million or more; 8 = Don't know/No answer; 9 = Refused

subset_data$WEA_SVNGSVL_MCQ_RECODED <- dplyr::case_when(
  subset_data$WEA_SVNGSVL_MCQ == 1 ~ 4,
  subset_data$WEA_SVNGSVL_MCQ == 2 ~ 3,
  subset_data$WEA_SVNGSVL_MCQ == 3 ~ 2,
  subset_data$WEA_SVNGSVL_MCQ == 4 ~ 1,
  subset_data$WEA_SVNGSVL_MCQ %in% c(8, 9) ~ 5,
  TRUE ~ NA_real_
)

# AFTER reversing
cat("\n=== AFTER reverse coding ===\n")
table(subset_data$WEA_SVNGSVL_MCQ_RECODED, useNA = "ifany")
# 1 = $1 million+; 2 = $100,000–<$1 million; 3 = $50,000–$99,999
# 4 = Less than $50,000; 5 = Don't know/No answer/Refused

subset_data$WEA_SVNGSVL_MCQ <- subset_data$WEA_SVNGSVL_MCQ_RECODED

cat("\nNAs in WEA_SVNGSVL_MCQ after recoding:", sum(is.na(subset_data$WEA_SVNGSVL_MCQ)), "\n")
subset_data <- subset_data %>% select(-WEA_SVNGSVL_MCQ_RECODED)

# Create savings_cat factor
subset_data <- subset_data %>%
  mutate(
    savings_cat = factor(
      WEA_SVNGSVL_MCQ,
      levels = c(1, 2, 3, 4, 5),
      labels = c("$1M+", "$100K–<$1M", "$50K–$100K", "<$50K", "Don't know/No answer/Refused")
    )
  )

# Final check
cat("\n=== savings_cat factor (final) ===\n")
table(subset_data$savings_cat, useNA = "ifany")

```















General Code needed for plots 

```{r}
#Custom theme for all figures

report_theme <- theme_classic(base_size = 11) +
  theme(
    legend.position = "none",
    
    plot.title = element_text(
      hjust = 0.5,
      face  = "bold",
      size  = 13,
      margin = margin(b = 8)
    ),
    
    axis.title.x = element_text(
      face = "bold",
      size = 11,
      margin = margin(t = 6)
    ),
    axis.title.y = element_text(
      face = "bold",
      size = 11,
      margin = margin(r = 6)
    ),
    
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10)
  )


```








## Reference info about variable names (new enviro variables) 


Tree canopy:
grtcc10_01_COM: Estimated percent of pixel covered by vegetation greater than 5m in height (2010)
grtcc10_02_COM: Average of estimated percent of pixel covered by vegetation greater than 5m in height within 100m (2010)
grtcc10_03_COM: Average of estimated percent of pixel covered by vegetation greater than 5m in height within 250m (2010)
grtcc10_04_COM: Average of estimated percent of pixel covered by vegetation greater than 5m in height within 500m (2010)
grtcc10_05_COM: Average of estimated percent of pixel covered by vegetation greater than 5m in height within 1 km (2010)




# 2.0 Descriptive overview of sample 
## 2.1 Demographics (age, sex, ethnicity) 


```{r}

subset_data <- readRDS("subset_data.rds")

# age and sex  ------------------------------------------------------

#pie chart for sex
df_sex <- subset_data %>%
  filter(!is.na(SEX_ASK_COM)) %>%
  count(SEX_ASK_COM) %>%
  mutate(
    percent = n / sum(n),
    label = paste0(SEX_ASK_COM, " (", round(percent * 100, 1), "%)")
  )

ggplot(df_sex, aes(x = 1, y = n, fill = SEX_ASK_COM)) +
  geom_col(width = 1, color = "white", linewidth = 1) +
  coord_polar(theta = "y") +
  geom_text(
    aes(label = label),
    position = position_stack(vjust = 0.5),
    size = 10,              # big, slide-friendly
    fontface = "plain"
  ) +
  scale_fill_manual(values = c("F" = "#1f77b4", "M" = "#ff7f0e")) +
  labs(title = "Sex Distribution", x = NULL, y = NULL, fill = NULL) +
  theme_void() +
  theme(
    # remove any remaining axis/radial artifacts
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    
    plot.title = element_text(
      hjust = 0.5,
      face  = "bold",
      size  = 38,
      margin = margin(b = 18)
    ),
    
    legend.position = "right",
    legend.text = element_text(size = 26),
    legend.key.size = grid::unit(1.2, "lines")
  )


#age boxplot 
subset_data %>%
  filter(!is.na(AGE_NMBR_COM)) %>%
  ggplot(aes(x = "", y = AGE_NMBR_COM)) +
  geom_boxplot(
    width = 0.25,
    fill = "white",
    color = "#2C7FB8",
    linewidth = 1.2
  ) +
  scale_y_continuous(
    breaks = seq(
      floor(min(subset_data$AGE_NMBR_COM, na.rm = TRUE) / 5) * 5,
      ceiling(max(subset_data$AGE_NMBR_COM, na.rm = TRUE) / 5) * 5,
      by = 5
    )
  ) +
  labs(
    title = "Age Distribution",
    x = NULL,
    y = "Age (years)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(
      hjust  = 0.5,
      face   = "bold",
      size   = 38,
      margin = margin(b = 18)
    ),
    axis.title.y = element_text(
      face   = "bold",
      size   = 30,
      margin = margin(r = 16)
    ),
    axis.title.x = element_blank(),
    axis.text.y  = element_text(size = 26),
    axis.text.x  = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_line(linetype = "dotted"),
    panel.grid.major.y = element_line(linetype = "dashed"),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background  = element_rect(fill = "white", color = NA)
  )


ggsave("age_distribution.png", width = 12, height = 7, dpi = 600)

# categories of cultural background:"White", "South Asian", "Black", "Asian","Arab", "Latin American", "Other"
subset_data %>%
  count(Race7) %>%
  mutate(
    percent = round(100 * n / sum(n), 1)
  )

  
```

```{r}

# ethnicity/self reported background barplot 
p_ethnicity <- subset_data %>%
  filter(!is.na(Race7)) %>%
  count(Race7, name = "n") %>%
  arrange(n) %>%
  ggplot(aes(x = n, y = Race7)) +
  geom_col(fill = "#2C7FB8", width = 0.7) +
  labs(
    title = "Self-Reported Cultural Background",
    x = "Number of Participants",
    y = NULL
  ) +
  report_theme

ggsave("Figures/ethnicity_barplot.png", plot = p_ethnicity, width = 5, height = 4, dpi = 300)
```


```{r}
#province



province_labels <- c(
  "1"  = "Alberta",
  "2"  = "British Columbia",
  "3"  = "Manitoba",
  "4"  = "New Brunswick",
  "5"  = "Newfoundland and Labrador",
  "6"  = "Nova Scotia",
  "7"  = "Ontario",
  "8"  = "Prince Edward Island",
  "9"  = "Quebec",
  "10" = "Saskatchewan"
)

prov <- subset_data$WGHTS_PROV_COM

prov_counts <- table(factor(prov, levels = names(province_labels)))
prov_pct    <- round(prop.table(prov_counts) * 100, 1)
n_missing   <- sum(is.na(prov))
pct_missing <- round(mean(is.na(prov)) * 100, 1)

summary_table_province <- tibble(
  Province = c(province_labels[names(prov_counts)], "Missing"),
  `n (%)`  = c(
    paste0(prov_counts, " (", prov_pct, "%)"),
    paste0(n_missing, " (", pct_missing, "%)")
  )
)

ft_province <- flextable(summary_table_province) %>%
  add_header_lines("Table X. Distribution of Participants by Province") %>%
  bold(part = "header") %>%
  bold(j = "Province") %>%
  autofit() %>%
  theme_booktabs() %>%
  fontsize(size = 10, part = "all") %>%
  fontsize(i = 1, size = 11, part = "header") %>%
  bold(i = 1, part = "header") %>%
  font(fontname = "Times New Roman", part = "all") %>%
  align(i = 1, align = "center", part = "header") %>%
  align(align = "left", part = "all") %>%
  align(j = "n (%)", align = "center", part = "body")

save_as_image(ft_province, path = "Tables/Table_province.png", res = 300)
```




## 2.2 Socioeconomic variables: Household Income and Wealth 


```{r}

#summary table of household income
#after reversing: 1 = $150,000 or more ; 2 = $100,000–$149,999 ; 3 = $50,000–$99,999; 4 = $20,000–$49,999; 5 = Less than $20,000; 6=  Don't know/No answer/Refused

subset_data %>%
  count(INC_TOT_COM) %>%
  mutate(
    percent = round(100 * n / sum(n), 1)
  )


# household income barplot 
p_income <- subset_data %>%
  count(income_cat, name = "n") %>%
  ggplot(aes(x = n, y = income_cat)) +
  geom_col(fill = "#2C7FB8", width = 0.7) +
  labs(
    title = "Household Income (Past 12 Months)",
    x = "Number of Participants",
    y = NULL
  ) +
  report_theme

ggsave("Figures/household_income_barplot.png", plot = p_income, width = 5, height = 4, dpi = 300)






#savings categories after reverse code are: 1=1 million or more. 2=100k to 1 million. 3=50-100k. 4=<50K. 5=don't know/no answer/refused 
subset_data %>%
  count(WEA_SVNGSVL_MCQ) %>%
  mutate(
    percent = round(100 * n / sum(n), 1)
  )

# savings/wealth barplot 
p_wealth <- subset_data %>%
  mutate(
    wealth_cat = factor(
      WEA_SVNGSVL_MCQ,
      levels = c(1, 2, 3, 4, 8, 9),
      labels = c(
        "<50K",
        "50K–100K",
        "100K to <$1M",
        "$1M+",
        "Don't know/No answer",
        "Refused"
      )
    )
  ) %>%
  count(wealth_cat, name = "n") %>%
  arrange(n) %>%
  ggplot(aes(x = n, y = wealth_cat)) +
  geom_col(fill = "#2C7FB8", width = 0.7) +
  labs(
    title = "Total Savings and Investments",
    x = "Number of Participants",
    y = NULL
  ) +
  report_theme

ggsave("Figures/savings_barplot.png", plot = p_wealth, width = 5, height = 4, dpi = 300)


```



# 3.0 Environmental Variables 

## 3.1 Pollution variables 


### 3.1.1 Compare PM25DALE_AYY_01 and pm25_1yr_com etc. 

PM25DALYY_01_COM
SO2OMIYY_COM
O3CHGYY_01_COM

```{r}
library(ggplot2)
library(tidyr)
library(dplyr)


# ── PM2.5 ──────────────────────────────────────────────────────────────────────
df_long <- subset_data %>%
  select(PM25DALYY_01_COM, pm25_1yr_com) %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") %>%
  filter(!is.na(value))

p_pm25_density <- ggplot(df_long, aes(x = value, fill = variable, color = variable)) +
  geom_density(alpha = 0.4) +
  labs(title = "Distribution Comparison: PM25DALYY_01_COM vs pm25_1yr_com",
       x = "PM2.5 Value", y = "Density") +
  theme_minimal()
ggsave("Figures/pm25_comparison_density.png", plot = p_pm25_density, width = 5, height = 4, dpi = 300)

p_pm25_boxplot <- ggplot(df_long, aes(x = variable, y = value, fill = variable)) +
  geom_boxplot(alpha = 0.6, outlier.shape = 21) +
  labs(title = "Boxplot Comparison: PM25DALYY_01_COM vs pm25_1yr_com",
       x = "", y = "PM2.5 Value") +
  theme_minimal()
ggsave("Figures/pm25_comparison_boxplot.png", plot = p_pm25_boxplot, width = 5, height = 4, dpi = 300)

summary(subset_data[, c("PM25DALYY_01_COM", "pm25_1yr_com")])

# ── SO2 ────────────────────────────────────────────────────────────────────────
df_long <- subset_data %>%
  select(so2_1yr_com, SO2OMIYY_COM) %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") %>%
  filter(!is.na(value))

p_so2_density <- ggplot(df_long, aes(x = value, fill = variable, color = variable)) +
  geom_density(alpha = 0.4) +
  labs(title = "Distribution Comparison: so2_1yr_com vs SO2OMIYY_COM",
       x = "SO2 Value", y = "Density") +
  theme_minimal()
ggsave("Figures/so2_comparison_density.png", plot = p_so2_density, width = 5, height = 4, dpi = 300)

p_so2_boxplot <- ggplot(df_long, aes(x = variable, y = value, fill = variable)) +
  geom_boxplot(alpha = 0.6, outlier.shape = 21) +
  labs(title = "Boxplot Comparison: so2_1yr_com vs SO2OMIYY_COM",
       x = "", y = "SO2 Value") +
  theme_minimal()
ggsave("Figures/so2_comparison_boxplot.png", plot = p_so2_boxplot, width = 5, height = 4, dpi = 300)

summary(subset_data[, c("so2_1yr_com", "SO2OMIYY_COM")])

# ── Ozone ──────────────────────────────────────────────────────────────────────
df_long <- subset_data %>%
  select(O3CHGYY_01_COM, o3_1yr_com) %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") %>%
  filter(!is.na(value))

p_o3_density <- ggplot(df_long, aes(x = value, fill = variable, color = variable)) +
  geom_density(alpha = 0.4) +
  labs(title = "Distribution Comparison: O3CHGYY_01_COM vs o3_1yr_com",
       x = "O3 Value", y = "Density") +
  theme_minimal()
ggsave("Figures/o3_comparison_density.png", plot = p_o3_density, width = 5, height = 4, dpi = 300)

p_o3_boxplot <- ggplot(df_long, aes(x = variable, y = value, fill = variable)) +
  geom_boxplot(alpha = 0.6, outlier.shape = 21) +
  labs(title = "Boxplot Comparison: O3CHGYY_01_COM vs o3_1yr_com",
       x = "", y = "O3 Value") +
  theme_minimal()
ggsave("Figures/o3_comparison_boxplot.png", plot = p_o3_boxplot, width = 5, height = 4, dpi = 300)

summary(subset_data[, c("O3CHGYY_01_COM", "o3_1yr_com")])

```



```{r}
ft <- flextable(summary_table) %>%
  add_header_lines("Table X. Air Quality Exposures") %>%
  bold(part = "header") %>%
  bold(j = "Variable") %>%
  autofit() %>%
  theme_booktabs() %>%
  fontsize(size = 10, part = "all") %>%
  fontsize(i = 1, size = 11, part = "header") %>%  # title slightly bigger
  bold(i = 1, part = "header") %>%                 # title bold
  font(fontname = "Times New Roman", part = "all") %>%
  align(i = 1, align = "center", part = "header") %>%  # title centered
  align(align = "left", part = "all") %>%
  align(j = c("Mean (SD)", "Median [Min, Max]", "Missing"), align = "center", part = "body")

save_as_image(ft, path = "Tables/Table_airquality.png", res = 300)
```

```{r}
#air pollution dashboard 
air_vars <- list(
  list(col = "pm25_1yr_com",  title = "PM2.5 (µg/m³)", file = "pm25",  binwidth = 0.5, limits = c(3, 11),  breaks = seq(3, 11, by = 1)),
  list(col = "no2_1yr_com",   title = "NO2 (ppb)",      file = "no2",   binwidth = 1,   limits = c(3, 16),  breaks = seq(3, 16, by = 2)),
  list(col = "so2_1yr_com",   title = "SO2 (ppb)",      file = "so2",   binwidth = 0.1, limits = c(0, 3.2), breaks = seq(0, 3.2, by = 0.5)),
  list(col = "o3_1yr_com",    title = "Ozone (ppb)",    file = "ozone", binwidth = 1,   limits = c(15, 33), breaks = seq(15, 33, by = 2))
)

air_plots <- list()

for (v in air_vars) {
  p <- ggplot(subset_data, aes(x = .data[[v$col]])) +
    geom_histogram(binwidth = v$binwidth, fill = "steelblue", color = "white") +
    scale_x_continuous(breaks = v$breaks, limits = v$limits) +
    labs(title = v$title, x = v$title, y = "Count") +
    report_theme +
    theme(plot.title = element_text(size = 9))
  
  air_plots[[v$col]] <- p
  ggsave(paste0("Figures/", v$file, "_histogram.png"), plot = p, width = 5, height = 4, dpi = 300)
}

dashboard_air <- wrap_plots(air_plots, ncol = 2) +
  plot_annotation(
    title = "Air Pollution Distributions",
    theme = theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 14))
  )

ggsave("Figures/air_pollution_dashboard.png", plot = dashboard_air, width = 10, height = 8, dpi = 300)

```













## 3.2 Condensation/noise

```{r}
## 3.3 Condensation and noise 
vars_housing <- list(
  list(col = "ENV_HMPRB_CON_MCQ", label = "Condensation"),
  list(col = "ENV_HMPRB_NOI_MCQ", label = "Noise")
)

summary_table_housing <- lapply(vars_housing, function(v) {
  x <- subset_data[[v$col]]
  tibble(
    Variable  = v$label,
    `Yes n (%)` = paste0(sum(x == 1, na.rm=TRUE), " (", round(mean(x == 1, na.rm=TRUE)*100, 1), "%)"),
    `No n (%)`  = paste0(sum(x == 0, na.rm=TRUE), " (", round(mean(x == 0, na.rm=TRUE)*100, 1), "%)"),
    `Missing`   = paste0(sum(is.na(x)), " (", round(mean(is.na(x))*100, 1), "%)")
  )
}) %>% bind_rows()

ft_housing <- flextable(summary_table_housing) %>%
  add_header_lines("Table X. Problems with Housing") %>%
  bold(part = "header") %>%
  bold(j = "Variable") %>%
  autofit() %>%
  theme_booktabs() %>%
  fontsize(size = 10, part = "all") %>%
  fontsize(i = 1, size = 11, part = "header") %>%
  bold(i = 1, part = "header") %>%
  font(fontname = "Times New Roman", part = "all") %>%
  align(i = 1, align = "center", part = "header") %>%
  align(align = "left", part = "all") %>%
  align(j = c("Yes n (%)", "No n (%)", "Missing"), align = "center", part = "body")

save_as_image(ft_housing, path = "Tables/Table_housing.png", res = 300)
```


```{r}

#Problems with condensation
condensation_bar <- ggplot(subset_data, aes(x = factor(ENV_HMPRB_CON_MCQ))) +
  geom_bar(fill = "steelblue", color = "white") +
  geom_text(
    aes(label = paste0(round(after_stat(count/sum(count)*100), 1), "%")),
    stat = "count",
    vjust = -0.5,
    size = 4
  ) +
  scale_x_discrete(labels = c("0" = "No", "1" = "Yes")) +
  labs(
    title = "Problems with Condensation in Home",
    x = "Condensation",
    y = "Count"
  ) +
  report_theme +
  theme(plot.margin = margin(t = 20, r = 10, b = 10, l = 10)) +
  coord_cartesian(clip = "off")

ggsave("Figures/condensation_bar.png", plot = condensation_bar, width = 5, height = 4, dpi = 300)




#Problems with noise

noise_bar <- ggplot(subset_data, aes(x = factor(ENV_HMPRB_NOI_MCQ))) +
  geom_bar(fill = "steelblue", color = "white") +
  geom_text(
    aes(label = paste0(round(after_stat(count/sum(count)*100), 1), "%")),
    stat = "count",
    vjust = -0.5,
    size = 4
  ) +
  scale_x_discrete(labels = c("0" = "No", "1" = "Yes")) +
  labs(
    title = "Problems with Noise",
    x = "Noise",
    y = "Count"
  ) +
  report_theme +
  theme(plot.margin = margin(t = 20, r = 10, b = 10, l = 10)) +
  coord_cartesian(clip = "off")

ggsave("Figures/noise_bar.png", plot = noise_bar, width = 5, height = 4, dpi = 300)



```












## 3.3 Tree Canopy Coverage (TCC) and % vegetation 
```{r}
# Table 1 - Tree Canopy Coverage
vars_canopy <- list(
  list(col = "grtcc10_01_COM", label = "Mean % of pixel covered by vegetation >5m"),
  list(col = "grtcc10_02_COM", label = "Mean % of pixel covered by vegetation >5m, within 100m"),
  list(col = "grtcc10_03_COM", label = "Mean % of pixel covered by vegetation >5m, within 250m"),
  list(col = "grtcc10_04_COM", label = "Mean % of pixel covered by vegetation >5m, within 500m"),
  list(col = "grtcc10_05_COM", label = "Mean % of pixel covered by vegetation >5m, within 1km")
)

summary_table_canopy <- lapply(vars_canopy, function(v) {
  x <- subset_data[[v$col]]
  tibble(
    Variable            = v$label,
    `Mean (SD)`         = paste0(round(mean(x, na.rm=TRUE), 2), " (", round(sd(x, na.rm=TRUE), 2), ")"),
    `Median [Min, Max]` = paste0(round(median(x, na.rm=TRUE), 2), " [", round(min(x, na.rm=TRUE), 2), ", ", round(max(x, na.rm=TRUE), 2), "]"),
    `Missing`           = paste0(sum(is.na(x)), " (", round(mean(is.na(x))*100, 1), "%)")
  )
}) %>% bind_rows()

ft_canopy <- flextable(summary_table_canopy) %>%
  add_header_lines("Table X. Tree Canopy Coverage") %>%
  bold(part = "header") %>%
  bold(j = "Variable") %>%
  autofit() %>%
  theme_booktabs() %>%
  fontsize(size = 10, part = "all") %>%
  fontsize(i = 1, size = 11, part = "header") %>%
  flextable::bold(i = 1, part = "header") %>%
  font(fontname = "Times New Roman", part = "all") %>%
  align(i = 1, align = "center", part = "header") %>%
  align(align = "left", part = "all") %>%
  align(j = c("Mean (SD)", "Median [Min, Max]", "Missing"), align = "center", part = "body")

save_as_image(ft_canopy, path = "Tables/Table_treecanopy.png", res = 300)

```



```{r}
# Tree canopy histogram dashboard 
canopy_vars <- list(
  list(col = "grtcc10_01_COM", title = "Estimated % Pixel Covered by Vegetation >5m",           file = "canopy_01"),
  list(col = "grtcc10_02_COM", title = "Mean % Pixel Covered by Vegetation >5m within 100m",    file = "canopy_02"),
  list(col = "grtcc10_03_COM", title = "Mean % Pixel Covered by Vegetation >5m within 250m",    file = "canopy_03"),
  list(col = "grtcc10_04_COM", title = "Mean % Pixel Covered by Vegetation >5m within 500m",    file = "canopy_04"),
  list(col = "grtcc10_05_COM", title = "Mean % Pixel Covered by Vegetation >5m within 1km",     file = "canopy_05")
)

for (v in canopy_vars) {# Save individual histograms
  p <- ggplot(subset_data, aes(x = .data[[v$col]])) +
    geom_histogram(binwidth = 5, fill = "forestgreen", color = "white") +
    scale_x_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60, 70, 80)) +
    labs(title = v$title, x = "Tree Canopy Cover (%)", y = "Count") +
    report_theme
  ggsave(paste0("Figures/", v$file, "_histogram.png"), plot = p, width = 5, height = 4, dpi = 300)
}

canopy_plots <- list()# Dashboard

for (v in canopy_vars) {
  p <- ggplot(subset_data, aes(x = .data[[v$col]])) +
    geom_histogram(binwidth = 5, fill = "forestgreen", color = "white") +
    scale_x_continuous(breaks = c(0, 20, 40, 60, 80)) +
    labs(title = v$title, x = "Tree Canopy Cover (%)", y = "Count") +
    report_theme +
    theme(plot.title = element_text(size = 8))
  
  canopy_plots[[v$col]] <- p
}

dashboard_canopy <- wrap_plots(canopy_plots, ncol = 3) +# 5 plots — arrange in 2 rows (3 top, 2 bottom)
  plot_annotation(
    title = "Tree Canopy Coverage Distributions",
    theme = theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 14))
  )

ggsave("Figures/tree_canopy_dashboard.png", plot = dashboard_canopy, width = 15, height = 9, dpi = 300)


```




## 3.4  % vegetation 

```{r}
# Table 2 - Vegetative Greenness
vars_greenness <- list(
  list(col = "GRUMPYY_01_COM", label = "Mean of annual % vegetative greenness, At postal code"),
  list(col = "GRUMPYY_02_COM", label = "Mean of annual % vegetative greenness, within 100m"),
  list(col = "GRUMPYY_03_COM", label = "Mean of annual % vegetative greenness, within 250m"),
  list(col = "GRUMPYY_04_COM", label = "Mean of annual % vegetative greenness, within 500m"),
  list(col = "GRUMPYY_05_COM", label = "Mean of annual % vegetative greenness, within 1km"),
  list(col = "GRUMPYY_06_COM", label = "Max of annual % vegetative greenness, within 100m"),
  list(col = "GRUMPYY_07_COM", label = "Max of annual % vegetative greenness, within 250m"),
  list(col = "GRUMPYY_08_COM", label = "Max of annual % vegetative greenness, within 500m"),
  list(col = "GRUMPYY_09_COM", label = "Max of annual % vegetative greenness, within 1km")
)

summary_table_greenness <- lapply(vars_greenness, function(v) {
  x <- subset_data[[v$col]]
  tibble(
    Variable            = v$label,
    `Mean (SD)`         = paste0(round(mean(x, na.rm=TRUE), 2), " (", round(sd(x, na.rm=TRUE), 2), ")"),
    `Median [Min, Max]` = paste0(round(median(x, na.rm=TRUE), 2), " [", round(min(x, na.rm=TRUE), 2), ", ", round(max(x, na.rm=TRUE), 2), "]"),
    `Missing`           = paste0(sum(is.na(x)), " (", round(mean(is.na(x))*100, 1), "%)")
  )
}) %>% bind_rows()

ft_greenness <- flextable(summary_table_greenness) %>%
  add_header_lines("Table X. Percent Vegetative Greenness") %>%
  bold(part = "header") %>%
  bold(j = "Variable") %>%
  autofit() %>%
  theme_booktabs() %>%
  fontsize(size = 10, part = "all") %>%
  fontsize(i = 1, size = 11, part = "header") %>%
  bold(i = 1, part = "header") %>%
  font(fontname = "Times New Roman", part = "all") %>%
  align(i = 1, align = "center", part = "header") %>%
  align(align = "left", part = "all") %>%
  align(j = c("Mean (SD)", "Median [Min, Max]", "Missing"), align = "center", part = "body")

save_as_image(ft_greenness, path = "Tables/Table_percentveg.png", res = 300)
```

```{r}

#Percent veg greenness histograms (9 of them)

greenness_vars <- list(
  list(col = "GRUMPYY_01_COM", title = "Annual % Vegetative Greenness at Postal Code",        file = "greenness_01"),
  list(col = "GRUMPYY_02_COM", title = "Mean Annual % Vegetative Greenness within 100m",       file = "greenness_02"),
  list(col = "GRUMPYY_03_COM", title = "Mean Annual % Vegetative Greenness within 250m",       file = "greenness_03"),
  list(col = "GRUMPYY_04_COM", title = "Mean Annual % Vegetative Greenness within 500m",       file = "greenness_04"),
  list(col = "GRUMPYY_05_COM", title = "Mean Annual % Vegetative Greenness within 1000m",      file = "greenness_05"),
  list(col = "GRUMPYY_06_COM", title = "Max Annual % Vegetative Greenness within 100m",        file = "greenness_06"),
  list(col = "GRUMPYY_07_COM", title = "Max Annual % Vegetative Greenness within 250m",        file = "greenness_07"),
  list(col = "GRUMPYY_08_COM", title = "Max Annual % Vegetative Greenness within 500m",        file = "greenness_08"),
  list(col = "GRUMPYY_09_COM", title = "Max Annual % Vegetative Greenness within 1000m",       file = "greenness_09")
)

for (v in greenness_vars) {
  p <- ggplot(subset_data, aes(x = .data[[v$col]])) +
    geom_histogram(binwidth = 5, fill = "forestgreen", color = "white") +
    scale_x_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90)) +
    labs(title = v$title, x = "Vegetative Greenness (%)", y = "Count") +
    report_theme
  ggsave(paste0("Figures/", v$file, "_histogram.png"), plot = p, width = 5, height = 4, dpi = 300)
}

library(patchwork)

# Use your loop to store plots in a list instead of saving individually
greenness_plots <- list()

for (v in greenness_vars) {
  p <- ggplot(subset_data, aes(x = .data[[v$col]])) +
    geom_histogram(binwidth = 5, fill = "forestgreen", color = "white") +
    scale_x_continuous(breaks = c(0, 20, 40, 60, 80, 100)) +
    labs(title = v$title, x = "Greenness (%)", y = "Count") +
    report_theme +
    theme(plot.title = element_text(size = 8))  # smaller title to fit
  
  greenness_plots[[v$col]] <- p
}

# Arrange in a 3x3 grid
dashboard <- wrap_plots(greenness_plots, ncol = 3) +
  plot_annotation(
    title = "Vegetative Greenness Distributions",
    theme = theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 14))
  )

ggsave("Figures/greenness_dashboard.png", plot = dashboard, width = 15, height = 12, dpi = 300)




ndvi_500m <- ggplot(subset_data, aes(x = GRLANYY_26_COM)) +
  geom_histogram(binwidth = 5, fill = "forestgreen", color = "white") +
  scale_x_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100)) +
  labs(
    title = "Maximum Annual NDVI within 500m",
    x = "NDVI (%)",
    y = "Count"
  ) +
  report_theme

ggsave("Figures/ndvi_500m_histogram.png", plot = ndvi_500m, width = 5, height = 4, dpi = 300)
GRLANYY_26_COM
```



## 3.5  NDVI 


```{r}
vars_ndvi <- list(
  list(col = "GRLANYY_26_COM", label = "Maximum annual mean NVDI within 250m")
)

summary_table_ndvi <- lapply(vars_ndvi, function(v) {
  x <- subset_data[[v$col]]
  tibble(
    Variable            = v$label,
    `Mean (SD)`         = paste0(round(mean(x, na.rm=TRUE), 2), " (", round(sd(x, na.rm=TRUE), 2), ")"),
    `Median [Min, Max]` = paste0(round(median(x, na.rm=TRUE), 2), " [", round(min(x, na.rm=TRUE), 2), ", ", round(max(x, na.rm=TRUE), 2), "]"),
    `Missing`           = paste0(sum(is.na(x)), " (", round(mean(is.na(x))*100, 1), "%)")
  )
}) %>% bind_rows()

ft_ndvi <- flextable(summary_table_ndvi) %>%
  add_header_lines("Table X. Normalized Difference Vegetation Index (NDVI)") %>%
  bold(part = "header") %>%
  bold(j = "Variable") %>%
  autofit() %>%
  theme_booktabs() %>%
  fontsize(size = 10, part = "all") %>%
  fontsize(i = 1, size = 11, part = "header") %>%
  bold(i = 1, part = "header") %>%
  font(fontname = "Times New Roman", part = "all") %>%
  align(i = 1, align = "center", part = "header") %>%
  align(align = "left", part = "all") %>%
  align(j = c("Mean (SD)", "Median [Min, Max]", "Missing"), align = "center", part = "body")

save_as_image(ft_ndvi, path = "Tables/Table_ndvi.png", res = 300)
```


```{r}
ndvi_500m <- ggplot(subset_data, aes(x = GRLANYY_26_COM)) +
  geom_histogram(binwidth = 0.02, fill = "forestgreen", color = "white") +
  scale_x_continuous(
    breaks = seq(0.60, 1.00, by = 0.05),
    limits = c(0.60, 1.00)
  ) +
  labs(
    title = "Maximum Annual NDVI within 500m",
    x = "NDVI",
    y = "Count"
  ) +
  report_theme

ggsave("Figures/ndvi_500m_histogram.png", plot = ndvi_500m, width = 5, height = 4, dpi = 300)

```




## 3.6 Length of roads within 200m 

```{r}
vars_roads <- list(
  list(col = "Phwy_length_200m_com",   label = "Length of highways within 200m of postal code (m)"),
  list(col = "Locrds_length_200m_com", label = "Length of local roads within 200m of postal code (m)")
)

summary_table_roads <- lapply(vars_roads, function(v) {
  x <- subset_data[[v$col]]
  n_zero_or_missing <- sum(x == 0 | is.na(x), na.rm = TRUE)
  pct_zero_or_missing <- round(n_zero_or_missing / length(x) * 100, 1)
  
  x_filtered <- x[x > 0 & !is.na(x)]  # remove zeros and NAs for stats
  
  tibble(
    Variable                = v$label,
    `Mean (SD)`             = paste0(round(mean(x_filtered), 2), " (", round(sd(x_filtered), 2), ")"),
    `Median [Min, Max]`     = paste0(round(median(x_filtered), 2), " [", round(min(x_filtered), 2), ", ", round(max(x_filtered), 2), "]"),
    `Zero or Missing n (%)` = paste0(n_zero_or_missing, " (", pct_zero_or_missing, "%)")
  )
}) %>% bind_rows()

ft_roads <- flextable(summary_table_roads) %>%
  add_header_lines("Table X. Proximity to Roads") %>%
  bold(part = "header") %>%
  bold(j = "Variable") %>%
  autofit() %>%
  theme_booktabs() %>%
  fontsize(size = 10, part = "all") %>%
  fontsize(i = 1, size = 11, part = "header") %>%
  bold(i = 1, part = "header") %>%
  font(fontname = "Times New Roman", part = "all") %>%
  align(i = 1, align = "center", part = "header") %>%
  align(align = "left", part = "all") %>%
  align(j = c("Mean (SD)", "Median [Min, Max]", "Zero or Missing n (%)"), align = "center", part = "body")

save_as_image(ft_roads, path = "Tables/Table_roads.png", res = 300)
```


```{r}
summary(subset_data$Phwy_length_200m_com) #Length of hwy within 200m of postal code centre
quantile(subset_data$Phwy_length_200m_com,
         probs = c(0, .01, .05, .10, .25, .50, .75, .90, .95, .99),
         na.rm = TRUE)
mean(subset_data$Phwy_length_200m_com == 0, na.rm = TRUE) #96% of the participants (1423 participants) have no hwy nearby. 



summary(subset_data$Locrds_length_200m_com) #Length of local rds within 200m of postal code centre
quantile(subset_data$Locrds_length_200m_com,
         probs = c(0, .01, .05, .10, .25, .50, .75, .90, .95, .99),
         na.rm = TRUE) 

hist(subset_data$Locrds_length_200m_com,
     breaks = 50,
     main = "Length of local rds within 200 m",
     xlab = "Meters",
     col = "grey")


```




# 4.0 Aging outcome variables: summary stats 

```{r}
#all numeric, no NAs 
str(subset_data$GrimAge_EAA) 
str(subset_data$PhenoAge_EAA) 
str(subset_data$DunedinPACE) 
str(subset_data$DNAmIC) 
```




# 5.0 Environmental variables by income (Pollution, condensation) 

## Pollution by income 
```{r}

#pollution boxplots by income group 


plot_data <- subset_data
ggplot(data = plot_data, aes(x = income_cat, y = pm25_1yr_com)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(
    title = "PM2.5 1 y average, by Income Group",
    x = "Household Income",
    y = "PM2.5 (µg/m³)"
  ) +
  report_theme

ggsave("Figures/boxplot_pm25.png", width = 5, height = 4, dpi = 300)



df %>% #check that there is significance for PM2.5 
  group_by(income_cat) %>%
  summarise(
    n = n(),
    median_pm25 = median(pm25_1yr_com, na.rm = TRUE),
    mean_pm25 = round(mean(pm25_1yr_com, na.rm = TRUE), 2)
  )
kruskal.test(pm25_1yr_com ~ income_cat, data = df)



ggplot(df_no2, aes(x = income_cat, y = no2_1yr_com)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(
    title = "NO2 1 y average, by Income Group",
    x = "Household Income",
    y = "NO2 (ppb)"
  ) +
  report_theme

ggsave("Figures/boxplot_no2.png", width = 5, height = 4, dpi = 300)



ggplot(df_so2, aes(x = income_cat, y = so2_1yr_com)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(
    title = "SO2 1 y average, by Income Group",
    x = "Household Income",
    y = "SO2 (ppb)"
  ) +
  report_theme

ggsave("Figures/boxplot_so2.png", width = 5, height = 4, dpi = 300)



ggplot(df_o3, aes(x = income_cat, y = o3_1yr_com)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(
    title = "Ozone 1 y average, by Income Group",
    x = "Household Income",
    y = "Ozone (ppb)"
  ) +
  report_theme

ggsave("Figures/boxplot_o3.png", width = 5, height = 4, dpi = 300)


ggplot(df_greenness, aes(x = income_cat, y = GRLANYY_26_COM)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(
    title = "Max NDVI within 500m, by Income Group",
    x = "Household Income",
    y = "Greenness (NDVI)"
  ) +
  report_theme

ggsave("Figures/boxplot_ndvi.png", width = 5, height = 4, dpi = 300)


```



## Pollution by income using the potentially better variables 


PM25DALYY_01_COM
SO2OMIYY_COM
O3CHGYY_01_COM


```{r}

#PM25DALYY_01_COM by Income Group 

plot_data <- subset_data

ggplot(data = plot_data, aes(x = income_cat, y = PM25DALYY_01_COM)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(
    title = "PM2.5 1 year average, by Income Group",
    x = "Household Income",
    y = "PM2.5 (satellite, model-based) (µg/m³)"
  ) +
  report_theme
ggsave("Figures/boxplot_PM25DALYY_01_COM.png", width = 5, height = 4, dpi = 300)



plot_data %>% #check that there is significance for PM2.5 
  group_by(income_cat) %>%
  summarise(
    n = n(),
    median_pm25 = median(PM25DALYY_01_COM, na.rm = TRUE),
    mean_pm25 = round(mean(PM25DALYY_01_COM, na.rm = TRUE), 2)
  )
kruskal.test(PM25DALYY_01_COM ~ income_cat, data = df)







ggplot(plot_data, aes(x = income_cat, y = SO2OMIYY_COM)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(
    title = "SO2 3 year annual average at postal code, by Income Group",
    x = "Household Income",
    y = "3 y average SO2 (satellite derived)(ppb)"
  ) +
  report_theme

ggsave("Figures/boxplot_SO2OMIYY_COM.png", width = 5, height = 4, dpi = 300)



ggplot(plot_data, aes(x = income_cat, y = O3CHGYY_01_COM)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(
    title = "O3 annual average ground-level at postal code, by Income Group",
    x = "Household Income",
    y = "Annual average O3 (model-based) (ppb)"
  ) +
  report_theme

ggsave("Figures/boxplot_O3CHGYY_01_COM.png", width = 5, height = 4, dpi = 300)




```


##Pollution by wealth (savings)




```{r}
# PM25DALYY_01_COM by Savings/Wealth Group
plot_data <- subset_data

ggplot(data = plot_data, aes(x = savings_cat, y = PM25DALYY_01_COM)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(
    title = "PM2.5 1 year average, by Total Savings/Investments",
    x = "Total Savings/Investments",
    y = "PM2.5 (satellite, model-based) (µg/m³)"
  ) +
  report_theme
ggsave("Figures/boxplot_PM25DALYY_01_COM_savings.png", width = 5, height = 4, dpi = 300)


ggplot(plot_data, aes(x = savings_cat, y = SO2OMIYY_COM)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(
    title = "SO2 3 year annual average, by Total Savings/Investments",
    x = "Total Savings/Investments",
    y = "3 y average SO2 (satellite derived) (ppb)"
  ) +
  report_theme
ggsave("Figures/boxplot_SO2OMIYY_COM_savings.png", width = 5, height = 4, dpi = 300)

ggplot(plot_data, aes(x = savings_cat, y = O3CHGYY_01_COM)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(
    title = "O3 annual average, by Total Savings/Investments",
    x = "Total Savings/Investments",
    y = "Annual average O3 (model-based) (ppb)"
  ) +
  report_theme
ggsave("Figures/boxplot_O3CHGYY_01_COM_savings.png", width = 5, height = 4, dpi = 300)


```



















### 5.1.2 PM2.5 median dot plot, by income group 


```{r}
df_pm25_summary <- df %>%
  group_by(income_cat) %>%
  summarise(
    n = n(),
    median_pm25 = median(pm25_1yr_com, na.rm = TRUE)
  )

p_pm25_dot <- ggplot(df_pm25_summary, aes(x = income_cat, y = median_pm25)) +
  geom_point(size = 4, color = "#2C7FB8", aes(shape = "Median")) +
  geom_text(aes(label = paste0("n=", n)), vjust = 2, size = 3.5) +
  scale_shape_manual(name = "", values = c("Median" = 16)) +
  scale_y_continuous(limits = c(6.0, 6.8)) +
  labs(title = "Median PM2.5 Exposure by Income Group",
       subtitle = "Kruskal-Wallis p = 0.032",
       x = "Household Income",
       y = "Median PM2.5 (µg/m³)") +
  report_theme

ggsave("Figures/pm25_median_dot.png", plot = p_pm25_dot,
       width = 6, height = 4, dpi = 300)
```



```{r}
air_income_plots <- list()

for (v in air_income_vars) {
  p <- subset_data %>%
    filter(income_cat %in% c("<20K", "20K–50K", "50K–100K", "100K–150K", "150K+")) %>%
    ggplot(aes(x = income_cat, y = .data[[v$col]], fill = income_cat)) +
    geom_boxplot(outlier.alpha = 0.3, width = 0.6) +
    scale_fill_brewer(palette = "Blues") +
    labs(
      title = v$title,
      x = "Household Income",
      y = v$y_lab
    ) +
    report_theme +
    theme(
      plot.title = element_text(size = 13, face = "bold"),
      axis.title = element_text(size = 11),
      axis.text = element_text(size = 9),
      legend.position = "none"
    )
  
  air_income_plots[[v$col]] <- p
}

dashboard_air_income <- wrap_plots(air_income_plots, ncol = 2) &
  theme(
    plot.margin = margin(t = 15, r = 15, b = 15, l = 15),
    plot.background = element_rect(color = "grey60", linewidth = 1, fill = "white"),
    panel.background = element_rect(fill = "white")
  )

dashboard_air_income <- dashboard_air_income +
  plot_annotation(
    title = "Air Pollution by Household Income",
    theme = theme(
      plot.title    = element_text(hjust = 0.5, face = "bold", size = 22),
      plot.subtitle = element_text(hjust = 0.5, size = 11, color = "grey40"),
      plot.caption  = element_text(size = 9, color = "grey40"),
      plot.margin   = margin(t = 20, r = 10, b = 10, l = 10),
      plot.background = element_rect(fill = "grey90", color = NA)
    )
  )

ggsave("Figures/air_pollution_income_dashboard.png", plot = dashboard_air_income, width = 16, height = 13, dpi = 300)
```





























## 5.2 Condensation by income 

```{r}
#condensation by income

cond_only <- subset_data %>% #keep only those with condensation problems (54 people)
  filter(ENV_HMPRB_CON_MCQ == 1)

cond_counts <- cond_only %>%
  filter(INC_TOT_COM %in% 1:5) %>%
  mutate(income_cat = factor(INC_TOT_COM, levels = 1:5, labels = income_labels)) %>%
  count(income_cat, name = "n") %>%
  mutate(pct = n / sum(n))

cond_counts

ggplot(cond_counts, aes(x = income_cat, y = n)) +
  geom_col(fill = "#2C7FB8", width = 0.6) +
  geom_text(
    aes(label = paste0(n, " (", percent(pct, accuracy = 0.1), ")")),
    vjust = -0.35,
    size = 3.2  # scaled to base_size 11 theme
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Homes with Condensation, by Income Group",
    x = "Household Income",
    y = "Homes with Condensation"
  ) +
  report_theme

ggsave("Figures/barplot_condensation.png", width = 5, height = 4, dpi = 300)





ggplot(df, aes(x = income_cat, y = pm25_1yr_com)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(
    title = "PM2.5 1 y average, by Income Group",
    x = "Household Income",
    y = "PM2.5 (µg/m³)"
  ) +
  report_theme

ggsave("Figures/boxplot_pm25.png", width = 5, height = 4, dpi = 300)

```








## 5.3 Tree canopy coverage by income 

```{r}

#tree canopy by income, boxplots 
p_veg1 <- ggplot(df, aes(x = income_cat, y = grtcc10_01_COM)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(title = "Vegetation Cover (>5m) by Income Group",
       x = "Household Income", y = "% Pixel Covered by Vegetation >5m") +
  report_theme
ggsave("Figures/boxplot_vegetation_01.png", plot = p_veg1, width = 5, height = 4, dpi = 300)


p_veg2 <- ggplot(df, aes(x = income_cat, y = grtcc10_02_COM)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(title = "Vegetation Cover (>5m) by Income, 100m",
       x = "Household Income", y = "% Pixel Covered by Vegetation >5m") +
  report_theme
ggsave("Figures/boxplot_vegetation_02.png", plot = p_veg2, width = 5, height = 4, dpi = 300)


p_veg3 <- ggplot(df, aes(x = income_cat, y = grtcc10_03_COM)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(title = "Vegetation Cover (>5m) by Income, 250m",
       x = "Household Income", y = "% Pixel Covered by Vegetation >5m") +
  report_theme
ggsave("Figures/boxplot_vegetation_03.png", plot = p_veg3, width = 5, height = 4, dpi = 300)


p_veg4 <- ggplot(df, aes(x = income_cat, y = grtcc10_04_COM)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(title = "Vegetation Cover (>5m) by Income, 500m",
       x = "Household Income", y = "% Pixel Covered by Vegetation >5m") +
  report_theme
ggsave("Figures/boxplot_vegetation_04.png", plot = p_veg4, width = 5, height = 4, dpi = 300)


p_veg5 <- ggplot(df, aes(x = income_cat, y = grtcc10_05_COM)) +
  geom_boxplot(fill = "#2C7FB8", outlier.alpha = 0.5, width = 0.6) +
  labs(title = "Vegetation Cover (>5m) by Income, 1km",
       x = "Household Income", y = "% Pixel Covered by Vegetation >5m") +
  report_theme
ggsave("Figures/boxplot_vegetation_05.png", plot = p_veg5, width = 5, height = 4, dpi = 300)





df <- subset_data %>%
  filter(INC_TOT_COM %in% 1:5, !is.na(grtcc10_01_COM)) %>%
  mutate(
    income_cat = factor(INC_TOT_COM, levels = 1:5, labels = income_labels)
  )

ggplot(df, aes(x = income_cat, y = grtcc10_01_COM)) +
  geom_boxplot(
    fill = "#2C7FB8",
    outlier.alpha = 0.5,
    width = 0.6
  ) +
  labs(
    title = "Vegetation Cover (>5m) by Income Group",
    x = "Household Income",
    y = "Estimated % Pixel Covered by Vegetation >5m"
  ) +
  theme_classic(base_size = 26) +
  theme(
    legend.position = "none",
    
    plot.title = element_text(
      hjust = 0.5,
      face  = "bold",
      size  = 38,
      margin = margin(b = 18)
    ),
    
    axis.title.x = element_text(
      face = "bold",
      size = 30,
      margin = margin(t = 16)
    ),
    axis.title.y = element_text(
      face = "bold",
      size = 30,
      margin = margin(r = 16)
    ),
    
    axis.text.x = element_text(size = 26),
    axis.text.y = element_text(size = 26)
  )
```





## 5.4 Vegetative greenness fractions by income 
There is no pattern by income group. The values are all clustered between 26–31%.
The GRUMPY variable measures all green vegetation including grass, shrubs, parks — not just tree canopy in residential areas.


```{r}

library(flextable)

df_green_summary <- df %>%
  group_by(income_cat) %>%
  summarise(
    n = n(),
    median_green = median(GRUMPYY_01_COM, na.rm = TRUE),
    mean_green = round(mean(GRUMPYY_01_COM, na.rm = TRUE), 1)
  ) %>%
  rename(
    `Income Group` = income_cat,
    `N` = n,
    `Median Greenness (%)` = median_green,
    `Mean Greenness (%)` = mean_green
  )

ft_green <- flextable(df_green_summary) %>%
  flextable::add_header_lines("Table X. Vegetation Greenness (GRUMPY) by Income Group") %>%
  flextable::bold(part = "header") %>%
  flextable::bold(j = "Income Group") %>%
  flextable::autofit() %>%
  flextable::theme_booktabs() %>%
  flextable::fontsize(size = 10, part = "all") %>%
  flextable::fontsize(i = 1, size = 11, part = "header") %>%
  flextable::bold(i = 1, part = "header") %>%
  flextable::font(fontname = "Times New Roman", part = "all") %>%
  flextable::align(i = 1, align = "center", part = "header") %>%
  flextable::align(align = "left", part = "all") %>%
  flextable::align(j = c("N", "Median Greenness (%)", "Mean Greenness (%)"), align = "center", part = "body")

save_as_image(ft_green, path = "Tables/Table_veg_green_fraction_byincome.png", res = 300)

```




## 5.5 Road length (nearby) by income 
This was not a good variable to look at. 

```{r}
#median hwy length within 200 m, by income group, just the participants living near a hwy 

p_hwy_dot <- ggplot(df_hwy_summary, aes(x = income_cat, y = median_length)) +
  geom_point(size = 4, color = "#2C7FB8", aes(shape = "Median")) +
  geom_text(aes(label = paste0("n=", n)), vjust = 2, size = 3.5) +
  scale_shape_manual(name = "", values = c("Median" = 16)) +
  scale_y_continuous(limits = c(40, 105)) +
  labs(title = "Median Highway Length within 200m by Income Group",
       subtitle = "Among participants living near a highway only",
       x = "Household Income",
       y = "Median Highway Length (m)") +
  report_theme
ggsave("Figures/highway_median_dot.png", plot = p_hwy_dot, 
       width = 6, height = 4, dpi = 300)


#median road length within 200 m, by income group, just the participants living near a road 

df_local <- df %>% filter(Locrds_length_200m_com > 0)

df_local_summary <- df_local %>%
  group_by(income_cat) %>%
  summarise(
    n = n(),
    median_length = median(Locrds_length_200m_com, na.rm = TRUE)
  )

p_local_dot <- ggplot(df_local_summary, aes(x = income_cat, y = median_length)) +
  geom_point(size = 4, color = "#2C7FB8", aes(shape = "Median")) +
  geom_text(aes(label = paste0("n=", n)), vjust = 2, size = 3.5) +
  scale_shape_manual(name = "", values = c("Median" = 16)) +
  labs(title = "Median Local Road Length within 200m by Income Group",
       subtitle = "Among participants living near a local road only",
       x = "Household Income",
       y = "Median Local Road Length (m)") +
  report_theme

ggsave("Figures/localroad_median_dot.png", plot = p_local_dot,
       width = 6, height = 4, dpi = 300)

```































# 6.0 Multiple linear regressions
## 6.1 Income predicting aging (stratified by age)

```{r}

#make variables into factors first 
subset_data$SEX_ASK_COM  <- as.factor(subset_data$SEX_ASK_COM)
subset_data$Race7        <- as.factor(subset_data$Race7)
subset_data$INC_TOT_COM  <- as.factor(subset_data$INC_TOT_COM)
subset_data$WGHTS_PROV_COM <- as.factor(subset_data$WGHTS_PROV_COM)
str(subset_data[, c("SEX_ASK_COM", "Race7", "INC_TOT_COM", "WGHTS_PROV_COM")])#now these are factors 


class(subset_data$AGE_NMBR_COM)#these are integers, ok to leave alone
class(subset_data$CD4nv) 


# Regression: Income predicting biological aging
#stratified by age 
subset_data$age_group <- ifelse(subset_data$AGE_NMBR_COM < 65, "45-64", "65-85")
subset_data_firstgrp  <- subset_data[subset_data$age_group == "45-64", ]
subset_data_secondgrp <- subset_data[subset_data$age_group == "65-85", ]

table(subset_data$INC_TOT_COM) #make sure it is reverse coded 


#GrimAGe 
library(broom)
model <- lm(GrimAge_EAA ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data_firstgrp)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "GrimAge_model_results_firstgrp.csv", row.names = FALSE)


model <- lm(GrimAge_EAA ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data_secondgrp)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "GrimAge_model_results_firstgrp.csv", row.names = FALSE)




#PhenoAge 
library(broom)
model <- lm(PhenoAge_EAA ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data_firstgrp)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "GrimAge_model_results_firstgrp.csv", row.names = FALSE)


model <- lm(PhenoAge_EAA ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data_secondgrp)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "GrimAge_model_results_firstgrp.csv", row.names = FALSE)


#DunedinPACE 

model <- lm(DunedinPACE ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data_firstgrp)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "GrimAge_model_results_firstgrp.csv", row.names = FALSE)


model <- lm(DunedinPACE ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data_secondgrp)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "GrimAge_model_results_firstgrp.csv", row.names = FALSE)



#DNAmIC 
model <- lm(DNAmIC ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data_firstgrp)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "GrimAge_model_results_firstgrp.csv", row.names = FALSE)


model <- lm(DNAmIC ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data_secondgrp)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "GrimAge_model_results_firstgrp.csv", row.names = FALSE)


```

## 6.2 Income predicting environmental 
### 6.2.1 Income predicting air pollution 

```{r}
#Income predicting PM2.5 
library(broom)
model <- lm(pm25_1yr_com ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "Income_PM2.5.csv", row.names = FALSE)


library(broom)
model <- lm(PM25DALYY_01_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "Income_PM2.5.csv", row.names = FALSE)




#Income predicting SO2 
model <- lm(so2_1yr_com ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "Income_SO2.csv", row.names = FALSE)



model <- lm(SO2OMIYY_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + #other SO2 variable (3 y average)
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "Income_SO2.csv", row.names = FALSE)






#Income predicting NO2 
model <- lm(no2_1yr_com ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "Income_NO2.csv", row.names = FALSE)



#Income predicting ozone
model <- lm(o3_1yr_com ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "Income_ozone.csv", row.names = FALSE)


model <- lm(O3CHGYY_01_COM ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "Income_ozone.csv", row.names = FALSE)


```

### 6.2.1 Income predicting condensation  
```{r}


#Income predicting condensation
model <- lm(ENV_HMPRB_CON_MCQ ~ INC_TOT_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "Income_condensation.csv", row.names = FALSE)

```






## 6.3 Environmental exposure predicting aging 
```{r}

#pm2.5
model <- lm(GrimAge_EAA ~ pm25_1yr_com + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "pm25_grimage.csv", row.names = FALSE)

model <- lm(PhenoAge_EAA ~ pm25_1yr_com + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "pm25_phenoage.csv", row.names = FALSE)

model <- lm(DunedinPACE ~ pm25_1yr_com + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "pm25_Duned.csv", row.names = FALSE)

model <- lm(DNAmIC ~ pm25_1yr_com + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "pm25_IC.csv", row.names = FALSE)




#no2_1yr_com
model <- lm(GrimAge_EAA ~ no2_1yr_com + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "pm25_grimage.csv", row.names = FALSE)

model <- lm(PhenoAge_EAA ~ no2_1yr_com + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "pm25_phenoage.csv", row.names = FALSE)

model <- lm(DunedinPACE ~ no2_1yr_com + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "pm25_Duned.csv", row.names = FALSE)

model <- lm(DNAmIC ~ no2_1yr_com + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "pm25_IC.csv", row.names = FALSE)




#so2_1yr_com
model <- lm(GrimAge_EAA ~ so2_1yr_com + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "pm25_grimage.csv", row.names = FALSE)


model <- lm(PhenoAge_EAA ~ so2_1yr_com + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "pm25_phenoage.csv", row.names = FALSE)


model <- lm(DunedinPACE ~ so2_1yr_com + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "pm25_Duned.csv", row.names = FALSE)


model <- lm(DNAmIC ~ so2_1yr_com + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "pm25_IC.csv", row.names = FALSE)




#ozone
model <- lm(GrimAge_EAA ~ o3_1yr_com + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "pm25_grimage.csv", row.names = FALSE)


model <- lm(PhenoAge_EAA ~ o3_1yr_com + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "pm25_phenoage.csv", row.names = FALSE)


model <- lm(DunedinPACE ~ o3_1yr_com + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "pm25_Duned.csv", row.names = FALSE)


model <- lm(DNAmIC ~ o3_1yr_com + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "pm25_IC.csv", row.names = FALSE)





#NDVI
model <- lm(GrimAge_EAA ~ GRLANYY_26_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "GrimAge_model_results_firstgrp.csv", row.names = FALSE)

model <- lm(PhenoAge_EAA ~ GRLANYY_26_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "GrimAge_model_results_firstgrp.csv", row.names = FALSE)

model <- lm(DunedinPACE ~ GRLANYY_26_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "GrimAge_model_results_firstgrp.csv", row.names = FALSE)

model <- lm(DNAmIC ~ GRLANYY_26_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "GrimAge_model_results_firstgrp.csv", row.names = FALSE)







#Average of estimated percent of pixel covered by vegetation greater than 5m in height within 100m (2010)
#grtcc10_02_COM

model <- lm(GrimAge_EAA ~ grtcc10_05_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "GrimAge_treecanopy.csv", row.names = FALSE)

model <- lm(PhenoAge_EAA ~ grtcc10_05_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "PhenoAge_treecanopy.csv", row.names = FALSE)

model <- lm(DunedinPACE ~ grtcc10_05_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "Dune_treecanopy.csv", row.names = FALSE)

model <- lm(DNAmIC ~ grtcc10_05_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "IC_treecanopy.csv", row.names = FALSE)





#Max of annual percentages of vegetative greeness within 100 m 
#GRUMPYY_06_COM

model <- lm(GrimAge_EAA ~ GRUMPYY_09_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "GrimAge_%veg.csv", row.names = FALSE)

model <- lm(PhenoAge_EAA ~ GRUMPYY_09_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "PhenoAge_%veg.csv", row.names = FALSE)

model <- lm(DunedinPACE ~ GRUMPYY_09_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "Dune_%veg.csv", row.names = FALSE)

model <- lm(DNAmIC ~ GRUMPYY_09_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "IC_%vegy.csv", row.names = FALSE)










#Condensation
model <- lm(GrimAge_EAA ~ ENV_HMPRB_CON_MCQ + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "GrimAge_model_results_firstgrp.csv", row.names = FALSE)



model <- lm(PhenoAge_EAA ~ ENV_HMPRB_CON_MCQ + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "GrimAge_model_results_firstgrp.csv", row.names = FALSE)



model <- lm(DunedinPACE ~ ENV_HMPRB_CON_MCQ + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "GrimAge_model_results_firstgrp.csv", row.names = FALSE)



model <- lm(DNAmIC ~ ENV_HMPRB_CON_MCQ + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "GrimAge_model_results_firstgrp.csv", row.names = FALSE)




```


### 6.3.1 Enviro/aging with other air pollution variables 

```{r}


PM25DALYY_01_COM
SO2OMIYY_COM
O3CHGYY_01_COM


#pm2.5
model <- lm(GrimAge_EAA ~ PM25DALYY_01_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "pm25_grimage.csv", row.names = FALSE)

model <- lm(PhenoAge_EAA ~ PM25DALYY_01_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "pm25_phenoage.csv", row.names = FALSE)

model <- lm(DunedinPACE ~ PM25DALYY_01_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "pm25_Duned.csv", row.names = FALSE)

model <- lm(DNAmIC ~ PM25DALYY_01_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)
write.csv(results, "pm25_IC.csv", row.names = FALSE)





#so2_1yr_com
model <- lm(GrimAge_EAA ~ SO2OMIYY_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "pm25_grimage.csv", row.names = FALSE)


model <- lm(PhenoAge_EAA ~ SO2OMIYY_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "pm25_phenoage.csv", row.names = FALSE)


model <- lm(DunedinPACE ~ SO2OMIYY_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "pm25_Duned.csv", row.names = FALSE)


model <- lm(DNAmIC ~ SO2OMIYY_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "pm25_IC.csv", row.names = FALSE)




#ozone
model <- lm(GrimAge_EAA ~ O3CHGYY_01_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "pm25_grimage.csv", row.names = FALSE)


model <- lm(PhenoAge_EAA ~ O3CHGYY_01_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "pm25_phenoage.csv", row.names = FALSE)


model <- lm(DunedinPACE ~ O3CHGYY_01_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "pm25_Duned.csv", row.names = FALSE)


model <- lm(DNAmIC ~ O3CHGYY_01_COM + SEX_ASK_COM + AGE_NMBR_COM + 
              I(AGE_NMBR_COM^2) + WGHTS_PROV_COM + Race7
            + CD4nv + CD8nv + Mono + Treg + CD4mem + CD8mem + NK + Bmem + Bnv,
            data = subset_data)
results <- tidy(model, conf.int = TRUE, conf.level = 0.95)# Tidy table with estimates, SE, t, p, and 95% CI
write.csv(results, "pm25_IC.csv", row.names = FALSE)


```






























# 7 To do 
-median dot plots for NO2, SO2, Ozone 
-visualizations of tree canopy and veg. 


# 8 plots for presentation 


```{r}
air_income_plots_pres <- list()

for (v in air_income_vars) {
  p <- subset_data %>%
    filter(income_cat %in% c("<20K", "20K–50K", "50K–100K", "100K–150K", "150K+")) %>%
    ggplot(aes(x = income_cat, y = .data[[v$col]], fill = income_cat)) +
    geom_boxplot(outlier.alpha = 0.3, width = 0.6) +
    scale_fill_brewer(palette = "Blues") +
    labs(
      title = v$title,
      x = "Household Income",
      y = v$y_lab
    ) +
    report_theme +
    report_theme +
theme(
  plot.title      = element_text(size = 22, face = "bold"),
  axis.title.x    = element_text(size = 20, face = "bold", margin = margin(t = 6)),
  axis.title.y    = element_text(size = 20, face = "bold", margin = margin(r = 6)),
  axis.text.x     = element_text(size = 18),
  axis.text.y     = element_text(size = 18),
  legend.position = "none",
  plot.margin      = margin(t = 15, r = 15, b = 15, l = 15),
  plot.background  = element_rect(color = "grey60", linewidth = 1, fill = "white"),
  panel.background = element_rect(fill = "white")
)
  
  
  
  
  air_income_plots_pres[[v$col]] <- p
}

dashboard_air_income_pres <- wrap_plots(air_income_plots_pres, ncol = 2) +
  plot_annotation(
    title = "Air Pollution by Household Income",
    theme = theme(
      plot.title      = element_text(hjust = 0.5, face = "bold", size = 28),
      plot.margin     = margin(t = 20, r = 10, b = 10, l = 10),
      plot.background = element_rect(fill = "grey90", color = NA)
    )
  )


ggsave("Figures/air_pollution_income_dashboard_PRESENTATION.png",
       plot = dashboard_air_income_pres, width = 18, height = 14, dpi = 300)





```






#9 Imputation 











# Not using yet 



## Table1 of All variables (Summary Stats)
```{r}

# Table1 all the variables (label and make into factor first)  -----------------------------------------------

#demographic variables age, sex, race 
subset_data$AGE_NMBR_COM <- as.numeric(subset_data$AGE_NMBR_COM) 
subset_data$SEX_ASK_COM <- factor(subset_data$SEX_ASK_COM) #F is first 
subset_data$WGHTS_PROV_COM <- factor(subset_data$WGHTS_PROV_COM) #AB is first 
subset_data$Race7 <- factor(subset_data$Race7,
                            levels = c("White", "South Asian", "Black", "Asian",
                                       "Arab", "Other")) #White is reference

label(subset_data$SEX_ASK_COM) <- "Sex (M/F)"#label categorical variables 
label(subset_data$AGE_NMBR_COM) <- "Age (years)"
label(subset_data$Race7) <- "Cultural background"

table1(~ SEX_ASK_COM + AGE_NMBR_COM + Race7, data = subset_data)

#SEC variables 
table(subset_data$INC_TOT_COM)
subset_data$INC_TOT_COM_factored <- factor(subset_data$INC_TOT_COM,#household income
                                                levels = c(1, 2, 3, 4, 5),
                                                labels = c("150K+",
                                                      "100K to 150K",
                                                      "50K–100K",
                                                      "20K-50K",
                                                      "<20K"))
                           

table(subset_data$ED_HIGH_COM)
subset_data$ED_HIGH_COM_factored <- factor(subset_data$ED_HIGH_COM,#education
                                                levels = c(1, 2, 3, 4, 5, 6),
                                                labels = c("01 = Above bachelor’s degg",
                                                           "02 = Bachelor’s deg",
                                                           "03 = Uni cert below bachelor’s",
                                                           "04 = Non-univ college diplomas",
                                                           "05 = Trade cert/diploma",
                                                           "06 = Above bachelor’s deg"))


table(subset_data$WEA_SVNGSVL_MCQ)
subset_data$WEA_SVNGSVL_MCQ_factored <- factor(subset_data$WEA_SVNGSVL_MCQ,#savings
                                                    levels = c(1, 2, 3, 4, 5),
                                                    labels = c("$1 million +", 
                                                               "100K to <$1 million",
                                                               "50K–100K",
                                                               "<50K",
                                                               "Don't know/No answer/Refused"
                                                               ))



label(subset_data$INC_TOT_COM_factored)          <- "Total Household Income (past 12 months)"
label(subset_data$ED_HIGH_COM_factored)           <- "Highest Education Obtained"
label(subset_data$WEA_SVNGSVL_MCQ_factored)       <- "Total Savings and Investments"

table1(~ INC_TOT_COM_factored + ED_HIGH_COM_factored + WEA_SVNGSVL_MCQ_factored, data = subset_data)



#environmental variables 

table(subset_data$ENV_HMPRB_CON_MCQ)
subset_data$ENV_HMPRB_CON_MCQ_factored <- factor(subset_data$ENV_HMPRB_CON_MCQ,#condensation
                                                      levels = c(0, 1),
                                                      labels = c("No", "Yes"))

table(subset_data$ENV_HMPRB_NOI_MCQ)
subset_data$ENV_HMPRB_NOI_MCQ_factored <- factor(subset_data$ENV_HMPRB_NOI_MCQ,#noise
                                                      levels = c(0, 1),
                                                      labels = c("No", "Yes"))



label(subset_data$ENV_HMPRB_CON_MCQ_factored)     <- "Home Problems: Condensation"
label(subset_data$ENV_HMPRB_NOI_MCQ_factored)     <- "Home Problems: Noise"
label(subset_data$pm25_1yr_com)          <- "PM2.5 (1-year average)"
label(subset_data$no2_1yr_com)          <- "NO2 (1-year average)"
label(subset_data$so2_1yr_com)          <- "SO2 (1-year average)"
label(subset_data$o3_1yr_com)           <- "Ozone (1-year average)"
label(subset_data$GRLANYY_26_COM)       <- "Greenness (Maximum of annual mean NDVI within 250m)"
label(subset_data$temp_1yr_com)         <- "Temperature (1-year average)"
label(subset_data$LGTNLTYY_01_COM)      <- "Nighttime Light Brightness"
label(subset_data$Locrds_length_200m_com)  <- "Length of local rds within 200m of postal code centre"
label(subset_data$Phwy_length_200m_com) <- "Length of hwy within 200m of postal code centre" 

table1(~ ENV_HMPRB_CON_MCQ_factored + ENV_HMPRB_NOI_MCQ_factored + pm25_1yr_com + no2_1yr_com + so2_1yr_com 
       + o3_1yr_com +  GRLANYY_26_COM + temp_1yr_com + LGTNLTYY_01_COM + 
         Locrds_length_200m_com + Phwy_length_200m_com, data = subset_data)



label(subset_data$grtcc10_01_COM)  <- "Mean % of pixel covered by vegetation >5m"
label(subset_data$grtcc10_02_COM)  <- "Mean % of pixel covered by vegetation >5m within 100m"
label(subset_data$grtcc10_03_COM)  <- "Mean % of pixel covered by vegetation >5m within 250m"
label(subset_data$grtcc10_04_COM)  <- "Mean % of pixel covered by vegetation >5m within 500m"
label(subset_data$grtcc10_05_COM)  <- "Mean % of pixel covered by vegetation >5m within 1km"

label(subset_data$GRUMPYY_01_COM)  <- "Annual % vegetative greenness at postal code"
label(subset_data$GRUMPYY_02_COM)  <- "Mean of annual % of vegetative greenness within 100m"
label(subset_data$GRUMPYY_03_COM)  <- "Mean of annual % of vegetative greenness within 250m" 
label(subset_data$GRUMPYY_04_COM)  <- "Mean of annual % of vegetative greenness within 500m"
label(subset_data$GRUMPYY_05_COM)  <- "Mean of annual % of vegetative greenness within 1km"
label(subset_data$GRUMPYY_06_COM)  <- "Max of annual % of vegetative greenness within 100m" 
label(subset_data$GRUMPYY_07_COM)  <- "Max of annual % of vegetative greenness within 250m" 
label(subset_data$GRUMPYY_08_COM)  <- "Max of annual % of vegetative greenness within 500m" 
label(subset_data$GRUMPYY_09_COM)  <- "Max of annual % of vegetative greenness within 1km" 



table1(~ grtcc10_01_COM + grtcc10_02_COM + grtcc10_03_COM + grtcc10_04_COM + grtcc10_05_COM 
       + GRUMPYY_01_COM +  GRUMPYY_02_COM + GRUMPYY_03_COM + GRUMPYY_04_COM + 
         GRUMPYY_05_COM + GRUMPYY_06_COM + GRUMPYY_07_COM + GRUMPYY_08_COM + GRUMPYY_09_COM, data = subset_data)



#Aging measures (outcomes)

table1(~ GrimAge_EAA + PhenoAge_EAA + DunedinPACE + DNAmIC, data = subset_data)


```




## Covariates Smoking, Alcohol, PASE

```{r}

#PASE
str(subset_data$PA2_DSCR2_MCQ) #PASE is numeric variable 

#smoke
str(subset_data$ICQ_SMOKE_COM) #smoke is integer (though it is categorical...)
table(subset_data$ICQ_SMOKE_COM) #1=Yes (I currently smoke), 2=No (I don't smoke and I never have), 3=Former (I don't smoke now but I have in the past)
subset_data$ICQ_SMOKE_COM_RECODED <- dplyr::case_when( #recode to 1=No I don't smoke and never have. 2=former. 3= yes i currently smoke 
  subset_data$ICQ_SMOKE_COM == 2 ~ 1,
  subset_data$ICQ_SMOKE_COM == 3 ~ 2,
  subset_data$ICQ_SMOKE_COM == 1 ~ 3,
  TRUE ~ NA_real_   
)
table(subset_data$ICQ_SMOKE_COM_RECODED)
subset_data$ICQ_SMOKE_COM <- subset_data$ICQ_SMOKE_COM_RECODED
sum(is.na(subset_data$ICQ_SMOKE_COM)) #number of NAs is 229


#alcohol
str(subset_data$ALC_FREQ_COM) #alcohol is integer (though it is categorical...)
table(subset_data$ALC_FREQ_COM)  

01			Almost every day (incl. 6 times a week)
02			4-5 times a week
03			2-3 times a week
04			Once a week
05			2-3 times a month
06			About once a month
07			Less than once a month
96=never. 98=don't know/no answer 

subset_data$ALC_FREQ_COM_RECODED <- dplyr::case_when( #recode to 1=Least drinking. 
  subset_data$ALC_FREQ_COM == 7 ~ 1,
  subset_data$ALC_FREQ_COM == 6 ~ 2,
  subset_data$ALC_FREQ_COM == 5 ~ 3,
  subset_data$ALC_FREQ_COM == 4 ~ 4,
  subset_data$ALC_FREQ_COM == 3 ~ 5,
  subset_data$ALC_FREQ_COM == 2 ~ 6,
  subset_data$ALC_FREQ_COM == 1 ~ 7,
  subset_data$ALC_FREQ_COM == 96 ~ 8,
  subset_data$ALC_FREQ_COM == 98 ~ NA_real_,  
  TRUE ~ NA_real_
)
table(subset_data$ALC_FREQ_COM_RECODED)
subset_data$ALC_FREQ_COM <- subset_data$ALC_FREQ_COM_RECODED
sum(is.na(subset_data$ALC_FREQ_COM)) #number of NAs is 257


#make sure the variables are appropriate types 

vars_to_check <- c(
  "AGE_NMBR_COM", "SEX_ASK_COM", "WGHTS_PROV_COM", "Race7",
  
  # Socioeconomic
  "INC_TOT_COM", "ED_HIGH_COM", "WEA_SVNGSVL_MCQ",
  
  # Outcomes
  "GrimAge_EAA", "PhenoAge_EAA", "DunedinPACE", "DNAmIC",
  
  # Environmental exposures
  "pm25_1yr_com", "no2_1yr_com", "o3_1yr_com", "GRLANYY_26_COM", "temp_1yr_com", 
  "LGTNLTYY_01_COM", "ENV_HMPRB_CON_MCQ", "ENV_HMPRB_NOI_MCQ",   
  
  # Lifestyle / behavioral
  "PA2_DSCR2_MCQ", "ALC_FREQ_COM", "ICQ_SMOKE_COM"
)
sapply(subset_data[vars_to_check], class)

```


# Delete 

## Charts: age, sex, ethnicity, income, wealth  
```{r}

#sex and age 
barplot(table(subset_data$SEX_ASK_COM),
        main = "Sex Distribution",
        col = "lightblue",
        ylab = "Count",
        las = 2,
        cex.names = 0.8)


# age and sex  ------------------------------------------------------

#pie chart for sex
df_sex <- subset_data %>%
  filter(!is.na(SEX_ASK_COM)) %>%
  count(SEX_ASK_COM) %>%
  mutate(
    percent = n / sum(n),
    label = paste0(SEX_ASK_COM, " (", round(percent * 100, 1), "%)")
  )

ggplot(df_sex, aes(x = 1, y = n, fill = SEX_ASK_COM)) +
  geom_col(width = 1, color = "white", linewidth = 1) +
  coord_polar(theta = "y") +
  geom_text(
    aes(label = label),
    position = position_stack(vjust = 0.5),
    size = 10,              # big, slide-friendly
    fontface = "plain"
  ) +
  scale_fill_manual(values = c("F" = "#1f77b4", "M" = "#ff7f0e")) +
  labs(title = "Sex Distribution", x = NULL, y = NULL, fill = NULL) +
  theme_void() +
  theme(
    # remove any remaining axis/radial artifacts
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    
    plot.title = element_text(
      hjust = 0.5,
      face  = "bold",
      size  = 38,
      margin = margin(b = 18)
    ),
    
    legend.position = "right",
    legend.text = element_text(size = 26),
    legend.key.size = grid::unit(1.2, "lines")
  )



#age boxplot 
subset_data %>%
  filter(!is.na(AGE_NMBR_COM)) %>%
  ggplot(aes(x = "", y = AGE_NMBR_COM)) +
  geom_boxplot(
    width = 0.25,
    fill = "white",
    color = "#2C7FB8",
    linewidth = 1.2
  ) +
  scale_y_continuous(
    breaks = seq(
      floor(min(subset_data$AGE_NMBR_COM, na.rm = TRUE) / 5) * 5,
      ceiling(max(subset_data$AGE_NMBR_COM, na.rm = TRUE) / 5) * 5,
      by = 5
    )
  ) +
  labs(
    title = "Age Distribution",
    x = NULL,
    y = "Age (years)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(
      hjust  = 0.5,
      face   = "bold",
      size   = 38,
      margin = margin(b = 18)
    ),
    axis.title.y = element_text(
      face   = "bold",
      size   = 30,
      margin = margin(r = 16)
    ),
    axis.title.x = element_blank(),
    axis.text.y  = element_text(size = 26),
    axis.text.x  = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_line(linetype = "dotted"),
    panel.grid.major.y = element_line(linetype = "dashed"),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background  = element_rect(fill = "white", color = NA)
  )



# ethnicity, income, wealth ---------------------------------------------------
#cultural background
#horizontal bar plots

subset_data %>%
  filter(!is.na(Race7)) %>%
  count(Race7, name = "n") %>%
  arrange(n) %>%
  ggplot(aes(x = n, y = Race7)) +
  geom_col(fill = "#2C7FB8", width = 0.7) +
  labs(
    title = "Self-Reported Cultural Background",
    x = "Number of participants",
    y = NULL
  ) +
  theme_classic(base_size = 26) +   # ⬅️ bigger global base
  theme(
    legend.position = "none",
    
    # TITLE — BIG
    plot.title = element_text(
      hjust = 0.5,
      face  = "bold",
      size  = 38,          # ⬅️ increased
      margin = margin(b = 18)
    ),
    
    # AXIS TITLES — BIGGER
    axis.title.x = element_text(
      face = "bold",
      size = 30,          # ⬅️ increased
      margin = margin(t = 16)
    ),
    axis.title.y = element_blank(),
    
    # AXIS TICK LABELS — BIGGER
    axis.text.x = element_text(size = 26),   # ⬅️ increased
    axis.text.y = element_text(size = 26)
  )




#horizontal bar chart of household income categories 
subset_data %>%
  mutate(
    income_cat = factor(
      INC_TOT_COM,  # personal/household income
      levels = c(1, 2, 3, 4, 5, 7, 8, 9),
      labels = c(
        "<20K",
        "20K–50K",
        "50K–100K",
        "100K–150K",
        "150K+",
        "Missing",
        "Don't know/No answer",
        "Refused"
      )
    )
  ) %>%
  count(income_cat, name = "n") %>%
  arrange(n) %>%
  ggplot(aes(x = n, y = income_cat)) +
  geom_col(fill = "#2C7FB8", width = 0.7) +
  labs(
    title = "Household Income (Past 12 Months)",
    x = "Number of participants",
    y = NULL
  ) +
  theme_classic(base_size = 26) +
  theme(
    legend.position = "none",
    
    # TITLE
    plot.title = element_text(
      hjust = 0.5,
      face  = "bold",
      size  = 38,
      margin = margin(b = 18)
    ),
    
    # AXIS TITLES
    axis.title.x = element_text(
      face = "bold",
      size = 30,
      margin = margin(t = 16)
    ),
    axis.title.y = element_blank(),
    
    # AXIS TICK LABELS
    axis.text.x = element_text(size = 26),
    axis.text.y = element_text(size = 26)
  )




#horizontal bar chart of wealth/savings   
subset_data %>%
  mutate(
    wealth_cat = factor(
      WEA_SVNGSVL_MCQ,  
      levels = c(1, 2, 3, 4, 8, 9),
      labels = c(
        "<50K",
        "50K–100K",
        "100K to <$1 million",
        "$1 million +",
        "Don't know/No answer",
        "Refused"
      )
    )
  ) %>%
  count(wealth_cat, name = "n") %>%
  arrange(n) %>%
  ggplot(aes(x = n, y = wealth_cat)) +
  geom_col(fill = "#2C7FB8", width = 0.7) +
  labs(
    title = "Total Savings and Investments",
    x = "Number of participants",
    y = NULL
  ) +
  theme_classic(base_size = 26) +
  theme(
    legend.position = "none",
    plot.title = element_text(
      hjust = 0.5, face = "bold", size = 38,
      margin = margin(b = 18)
    ),
    axis.title.x = element_text(
      face = "bold", size = 30,
      margin = margin(t = 16)
    ),
    axis.title.y = element_blank(),
    axis.text.x  = element_text(size = 26),
    axis.text.y  = element_text(size = 26)
  )



```


```{r}

#savings 
table(subset_data$WEA_SVNGSVL_MCQ, useNA = "ifany") #categories are #1 = Less than $50,000; 2 = $50,000–$99,999; 3 = $100,000–$999,999; 4 = $1 million or more; 8 = Don’t know/No answer; 9 = Refused
#there are 103 NAs to start with. 

subset_data$WEA_SVNGSVL_MCQ_RECODED <- dplyr::case_when( #reverse code so that 1=4, 2=3, 3=2, 4=1, 5=8,9 
  subset_data$WEA_SVNGSVL_MCQ == 1 ~ 4,
  subset_data$WEA_SVNGSVL_MCQ == 2 ~ 3,
  subset_data$WEA_SVNGSVL_MCQ == 3 ~ 2,
  subset_data$WEA_SVNGSVL_MCQ == 4 ~ 1,
  subset_data$WEA_SVNGSVL_MCQ %in% c(8, 9) ~ 5,
  TRUE ~ NA_real_   
)
subset_data$WEA_SVNGSVL_MCQ <- subset_data$WEA_SVNGSVL_MCQ_RECODED
table(subset_data$WEA_SVNGSVL_MCQ, useNA = "ifany")#103 NAs 


#education 
# Step 1: Coerce to numeric safely (from character/factor labels)
# Step 2: Set invalid categories to NA (97 = other, 98 = don't know, 99 = refused)
# Step 3: Reverse code only valid numeric values (1–6)
# Step 4: Convert to factor
# Step 5: Optional – drop unused factor levels (e.g., if -90 or others existed)
# Step 6: Check
#check NAs

table(subset_data$ED_HIGH_COM)
subset_data$ED_HIGH_COM <- as.numeric(as.character(subset_data$ED_HIGH_COM))
subset_data$ED_HIGH_COM[subset_data$ED_HIGH_COM %in% c(97, 98, 99)] <- NA
subset_data$ED_HIGH_COM <- ifelse(!is.na(subset_data$ED_HIGH_COM), 7 - subset_data$ED_HIGH_COM, NA)
subset_data$ED_HIGH_COM <- factor(subset_data$ED_HIGH_COM)
subset_data$ED_HIGH_COM <- droplevels(subset_data$ED_HIGH_COM)
table(subset_data$ED_HIGH_COM, useNA = "ifany") #231 NAs 

saveRDS(subset_data, file = "subset_data.rds")
```




