---
title: "Imputation"
output:
  word_document: default
  pdf_document: default
  html_document: default
date: "2026-05-20"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,       # hides code
  message = FALSE,    # suppresses messages
  warning = FALSE     # suppresses warnings
)
```


set pred
make sure certain variables don't predict each other 
think about which variables should be used to predict the others... 
multi imp=true. make sure it's looking at all of the datasets. 
do diagnostics after to make sure that the values look stable. bwplot (in keenan's code)


#1 create subset of variables 
Use other script for this 


#2 check missingness in each variable within subset 

Notes on province and PM2.5:
Province 3: 101 missing out of 156 total (~65% missing)
Province 7: 109 missing out of 309 total (~35% missing)


```{r}
library(mice)
install.packages("DataExplorer")
library(DataExplorer)
plot_missing(subset_data)
md.pattern(subset_data)


colSums(is.na(subset_data)) #how many NAs for each variable 

```


#3 what are the types of variables? 

```{r}
#what class are the variables 
data.frame(
  variable = names(subset_data),
  class = sapply(subset_data, class),
  n_unique = sapply(subset_data, function(x) length(unique(x))),
  n_missing = sapply(subset_data, function(x) sum(is.na(x)))
)

#change categorical variables into factors, not integers 
subset_data$SEX_ASK_COM <- as.factor(subset_data$SEX_ASK_COM)
subset_data$WGHTS_PROV_COM <- as.factor(subset_data$WGHTS_PROV_COM)
subset_data$ENV_HMPRB_CON_MCQ <- as.factor(subset_data$ENV_HMPRB_CON_MCQ)
subset_data$ALC_FREQ_COM <- as.ordered(subset_data$ALC_FREQ_COM)
subset_data$ICQ_SMOKE_COM <- as.factor(subset_data$ICQ_SMOKE_COM)
subset_data$INC_TOT_COM <- as.ordered(subset_data$INC_TOT_COM)

#check that it worked 
data.frame(
  variable = names(subset_data),
  class = sapply(subset_data, function(x) paste(class(x), collapse = ", "))
)


```


#4 Initialize MICE and note which variables to not impute. 

```{r}
# Initialize MICE
init = mice(subset_data, maxit = 0)
my_methods = init$method
pred = make.predictorMatrix(subset_data)


# outcomes: don't impute them, but keep as predictors
pred["GrimAge_EAA", ]  = 0  
pred["PhenoAge_EAA", ] = 0
pred["DunedinPACE", ]  = 0
pred["DNAmIC", ]       = 0

# remove entity_id entirely
pred[, "entity_id.x"]  = 0
pred["entity_id.x", ]  = 0


# Cell types have no missingness so don't need to be imputed
# but keep them as predictors (leave their columns as-is)
pred["CD4nv", ]  = 0
pred["CD8nv", ]  = 0
pred["Mono", ]   = 0
pred["Treg", ]   = 0
pred["CD4mem", ] = 0
pred["CD8mem", ] = 0
pred["NK", ]     = 0
pred["Bmem", ]   = 0
pred["Bnv", ]    = 0


```


# 5 Assign imputation methods for each variable

```{r}
# Specify imputation methods for each variable

# --- Do NOT impute these ---
my_methods["entity_id.x"] = "" # dont impute entity_id
my_methods["GrimAge_EAA"]      = ""   # outcome - do not impute
my_methods["PhenoAge_EAA"]     = ""   # outcome - do not impute
my_methods["DunedinPACE"]      = ""   # outcome - do not impute
my_methods["DNAmIC"]           = ""   # outcome - do not impute

# cell types - no missingness, do not impute
my_methods["CD4nv"]   = ""
my_methods["CD8nv"]   = ""
my_methods["Mono"]    = ""
my_methods["Treg"]    = ""
my_methods["CD4mem"]  = ""
my_methods["CD8mem"]  = ""
my_methods["NK"]      = ""
my_methods["Bmem"]    = ""
my_methods["Bnv"]     = ""

# --- Impute these ---
my_methods["AGE_NMBR_COM"]      = "pmm"     # continuous
my_methods["SEX_ASK_COM"]       = "logreg"  # binary
my_methods["WGHTS_PROV_COM"]    = "pmm"     # categorical >2 levels -> pmm for stability
my_methods["Race7"]             = "pmm"     # categorical >2 levels -> pmm for stability
my_methods["INC_TOT_COM"]       = "pmm"     # ordinal categorical -> pmm for stability
my_methods["ALC_FREQ_COM"]      = "pmm"     # ordinal categorical -> pmm for stability
my_methods["ICQ_SMOKE_COM"]     = "pmm"     # categorical >2 levels -> pmm for stability
my_methods["ENV_HMPRB_CON_MCQ"] = "logreg"  # binary
my_methods["PA2_DSCR2_MCQ"]     = "pmm"     # continuous

# environmental exposures
my_methods["pm25_1yr_com"]      = "pmm"     # continuous
my_methods["no2_1yr_com"]       = "pmm"     # continuous
my_methods["so2_1yr_com"]       = "pmm"     # continuous
my_methods["o3_1yr_com"]        = "pmm"     # continuous
my_methods["GRLANYY_26_COM"]    = "pmm"     # continuous (NDVI)

# tree canopy coverage
my_methods["grtcc10_01_COM"]    = "pmm"     # continuous
my_methods["grtcc10_02_COM"]    = "pmm"     # continuous
my_methods["grtcc10_03_COM"]    = "pmm"     # continuous
my_methods["grtcc10_04_COM"]    = "pmm"     # continuous
my_methods["grtcc10_05_COM"]    = "pmm"     # continuous

# vegetation coverage
my_methods["GRUMPYY_01_COM"]    = "pmm"     # continuous
my_methods["GRUMPYY_02_COM"]    = "pmm"     # continuous
my_methods["GRUMPYY_03_COM"]    = "pmm"     # continuous
my_methods["GRUMPYY_04_COM"]    = "pmm"     # continuous
my_methods["GRUMPYY_05_COM"]    = "pmm"     # continuous
my_methods["GRUMPYY_06_COM"]    = "pmm"     # continuous
my_methods["GRUMPYY_07_COM"]    = "pmm"     # continuous
my_methods["GRUMPYY_08_COM"]    = "pmm"     # continuous
my_methods["GRUMPYY_09_COM"]    = "pmm"     # continuous

# verify
my_methods
```


# 6 Run imputation               
```{r}
# Run multiple imputation
imp_data = mice(subset_data,
                method = my_methods,
                predictorMatrix = pred,
                m = 20,
                seed = 45,
                printFlag = TRUE)

# Save the imputed data object
saveRDS(imp_data, file = "imputed_data_CLSA.rds")

# Extract 20 completed datasets into a list
completed_list = lapply(1:20, function(i) complete(imp_data, action = i))

# Check NAs are gone in the variables we imputed
colSums(sapply(completed_list[[1]], is.na))
```

# 7 Run Diagnostic plots to check that the imputation looks reasonable 

```{r}
colSums(is.na(completed_list[[1]])) #check that there are no more NAs 

plot(imp_data)

stripplot(imp_data)






#Air pollution
densityplot(imp_data, ~pm25_1yr_com)
densityplot(imp_data, ~no2_1yr_com)
densityplot(imp_data, ~so2_1yr_com) #21% missing - most important to check 
densityplot(imp_data, ~o3_1yr_com)

# Greenspace
densityplot(imp_data, ~GRLANYY_26_COM)    # NDVI

# Tree canopy (all 5 buffers)
densityplot(imp_data, ~grtcc10_01_COM)
densityplot(imp_data, ~grtcc10_02_COM)
densityplot(imp_data, ~grtcc10_03_COM)
densityplot(imp_data, ~grtcc10_04_COM)
densityplot(imp_data, ~grtcc10_05_COM)

# Vegetation coverage (all 9 buffers)
densityplot(imp_data, ~GRUMPYY_01_COM)
densityplot(imp_data, ~GRUMPYY_02_COM)
densityplot(imp_data, ~GRUMPYY_03_COM)
densityplot(imp_data, ~GRUMPYY_04_COM)
densityplot(imp_data, ~GRUMPYY_05_COM)
densityplot(imp_data, ~GRUMPYY_06_COM)
densityplot(imp_data, ~GRUMPYY_07_COM)
densityplot(imp_data, ~GRUMPYY_08_COM)
densityplot(imp_data, ~GRUMPYY_09_COM)

# Lifestyle/demographic
densityplot(imp_data, ~PA2_DSCR2_MCQ)
stripplot(imp_data, ALC_FREQ_COM ~ .imp)
stripplot(imp_data, ICQ_SMOKE_COM ~ .imp)
stripplot(imp_data, INC_TOT_COM ~ .imp)
stripplot(imp_data, ENV_HMPRB_CON_MCQ ~ .imp)






```


# 8 Who is missing data with the air pollution measures? 

528 people out of 1478 (36%) would be dropped if I did complete case on all pollution variables. That is too many cases to drop. 

PM2.5
Province 3: 65% missing — MICE only has 55 observed values to learn the PM2.5 distribution for that province
Province 7: 35% missing — MICE has 200 observed values but 109 are missing

NO2:
The people missing NO2 are different people than those missing PM2.5.
Province 6: 58 missing out of 152 (~38% missing)
Province 9: 153 missing out of 295 (~52% missing)

```{r}
# look at characteristics of people missing vs not missing pm25
subset_data %>%
  mutate(missing_pm25 = is.na(pm25_1yr_com)) %>%
  group_by(missing_pm25) %>%
  summarise(
    n = n(),
    mean_age = mean(AGE_NMBR_COM, na.rm = TRUE),
    mean_income = mean(as.numeric(INC_TOT_COM), na.rm = TRUE))



subset_data %>%
  mutate(missing_pm25 = is.na(pm25_1yr_com)) %>%
  group_by(missing_pm25) %>%
  summarise(
    n = n(),
    mean_age = mean(AGE_NMBR_COM, na.rm = TRUE),
    mean_income = mean(as.numeric(INC_TOT_COM), na.rm = TRUE),
    pct_male = mean(SEX_ASK_COM == 1, na.rm = TRUE),  # adjust coding if needed
    mean_pm25 = mean(pm25_1yr_com, na.rm = TRUE),
    pct_race_white = mean(Race7 == 1, na.rm = TRUE))  # adjust coding if needed


#check which PM2.5 are missing (by province)
subset_data %>%
  mutate(missing_pm25 = is.na(pm25_1yr_com)) %>%
  group_by(WGHTS_PROV_COM, missing_pm25) %>%
  summarise(n = n()) %>%
  pivot_wider(names_from = missing_pm25, values_from = n)



#check which NO2 are missing (by province) - they are different than the PM2.5
subset_data %>%
  mutate(missing_pm25 = is.na(so2_1yr_com)) %>%
  group_by(WGHTS_PROV_COM, missing_pm25) %>%
  summarise(n = n()) %>%
  pivot_wider(names_from = missing_pm25, values_from = n)


#check how many total missing across all the air pollution variables (528!)
subset_data %>%
  mutate(missing_any = is.na(pm25_1yr_com) | is.na(no2_1yr_com) | 
           is.na(so2_1yr_com) | is.na(o3_1yr_com)) %>%
  count(missing_any)



#check if tree canopy coverage has the same provincial missingness pattern:
#province 5 has no tree canopy data! (112 missing) 
subset_data %>%
  mutate(missing_pm25 = is.na(grtcc10_01_COM)) %>%
  group_by(WGHTS_PROV_COM, missing_pm25) %>%
  summarise(n = n()) %>%
  pivot_wider(names_from = missing_pm25, values_from = n)

#check if %veg has same provincial missingness - no, it doesn't. it's pretty spread out. 
subset_data %>%
  mutate(missing_pm25 = is.na(GRUMPYY_04_COM)) %>%
  group_by(WGHTS_PROV_COM, missing_pm25) %>%
  summarise(n = n()) %>%
  pivot_wider(names_from = missing_pm25, values_from = n)


#check if condensation has same provincial missingness - no, it doesn't. it's pretty spread out. 
subset_data %>%
  mutate(missing_pm25 = is.na(ENV_HMPRB_CON_MCQ)) %>%
  group_by(WGHTS_PROV_COM, missing_pm25) %>%
  summarise(n = n()) %>%
  pivot_wider(names_from = missing_pm25, values_from = n)


```




















