---
title: "Descriptive Statistics, CLSA data, Methylation data set"
output:
  word_document: default
  pdf_document: default
  html_document: default
date: "2025-07-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,       # hides code
  message = FALSE,    # suppresses messages
  warning = FALSE     # suppresses warnings
)
```


### Categorical Variables: Descriptive statistics
Categorical variables were summarized using descriptive statistics. 

```{r}

library(dplyr)
library(ggplot2)
library(tidyverse)
library (haven)
library(table1)

# Load data  --------------------------------------------------------------

setwd("C:\\Users\\Amand\\OneDrive\\Documents\\GSAT\\Data") 
load("CLSA_baseline_and_clocks.rds")#load data
CLSA_methylation <- CLSA_baseline_and_clocks %>% #create subset with only DNAm data 
  filter(!is.na(DNAmAge_COM))

# Descriptive stats_predictors_categorical --------------------------------

table(CLSA_methylation$SEX_ASK_COM)#check what categories exist first
CLSA_methylation$SEX_ASK_COM_factored <- factor(CLSA_methylation$SEX_ASK_COM, #sex
                                                levels = c("F", "M"),
                                                labels = c("Female", "Male"))

table(CLSA_methylation$INC_TOT_COM)
CLSA_methylation$INC_TOT_COM_factored <- factor(CLSA_methylation$INC_TOT_COM,#personal income
                                                levels = c(1, 2, 3, 4, 5, 7, 8, 9),
                                                labels = c("<20K",
                                                           "20K–50K",
                                                           "50K–100K",
                                                           "100K–150K",
                                                           "150K+",
                                                           "Missing", 
                                                           "Don't know/No answer",
                                                           "Refused"))


table(CLSA_methylation$INC_PTOT_COM)
CLSA_methylation$INC_PTOT_COM_factored <- factor(CLSA_methylation$INC_PTOT_COM,#total household income 
                                                 levels = c(1, 2, 3, 4, 5, 7, 8, 9),
                                                 labels = c("<20K",
                                                            "20K–50K",
                                                            "50K–100K",
                                                            "100K–150K",
                                                            "150K+",
                                                            "Missing", 
                                                            "Don't know/No answer",
                                                            "Refused"))


table(CLSA_methylation$ED_HIGH_COM)
CLSA_methylation$ED_HIGH_COM_factored <- factor(CLSA_methylation$ED_HIGH_COM,#education
                                                levels = c(1, 2, 3, 4, 5, 6, 97, 98, 99),
                                                labels = c("01 = No post-sec",
                                                           "02 = Trade cert/diploma",
                                                           "03 = Non-univ college diploma",
                                                           "04 = Uni cert below bachelor’s",
                                                           "05 = Bachelor’s deg",
                                                           "06 = Above bachelor’s deg",
                                                           "97 = Other",
                                                           "98 = Don’t know/No ans",
                                                           "99 = Refused"))

table(CLSA_methylation$WEA_SVNGSVL_MCQ)
CLSA_methylation$WEA_SVNGSVL_MCQ_factored <- factor(CLSA_methylation$WEA_SVNGSVL_MCQ,#savings
                                                    levels = c(1, 2, 3, 4, 8, 9),
                                                    labels = c("<50K", 
                                                               "50K–100K",
                                                               "100K to <$1 million",
                                                               "$1 million +",
                                                               "Don't know/No answer",
                                                               "Refused"))
table(CLSA_methylation$WEA_DEBT_CCRD_MCQ)
CLSA_methylation$WEA_DEBT_CCRD_MCQ_factored <- factor(CLSA_methylation$WEA_DEBT_CCRD_MCQ,#debt
                                                      levels = c(0, 1),
                                                      labels = c("No", "Yes"))

table(CLSA_methylation$ENV_HMPRB_CON_MCQ)
CLSA_methylation$ENV_HMPRB_CON_MCQ_factored <- factor(CLSA_methylation$ENV_HMPRB_CON_MCQ,#condensation
                                                      levels = c(0, 1),
                                                      labels = c("No", "Yes"))

table(CLSA_methylation$ENV_HMPRB_NOI_MCQ)
CLSA_methylation$ENV_HMPRB_NOI_MCQ_factored <- factor(CLSA_methylation$ENV_HMPRB_NOI_MCQ,#noise
                                                      levels = c(0, 1),
                                                      labels = c("No", "Yes"))
table(CLSA_methylation$ICQ_SMOKE_COM)
CLSA_methylation$ICQ_SMOKE_COM_factored <- factor(CLSA_methylation$ICQ_SMOKE_COM,#smoking
                                                  levels = c(1, 2, 3, -8),
                                                  labels = c("Currently smoke",
                                                             "Never smoked",
                                                             "Former smoker",
                                                             "Missing"))
table(CLSA_methylation$ALC_FREQ_COM)
CLSA_methylation$ALC_FREQ_COM_factored <- factor(CLSA_methylation$ALC_FREQ_COM,
                                                 levels = c(1, 2, 3, 4, 5, 6, 7, 96, 98, 99, 77),
                                                 labels = c("Almost q day",
                                                            "4-5 times/wk",
                                                            "2-3 times/wk",
                                                            "Once/wk",
                                                            "2-3 times/month",
                                                            "About once/month",
                                                            "< once/month",
                                                            "Never",
                                                            "Don’t know/No answer",
                                                            "Refused",
                                                            "Missing"))

CLSA_methylation <- CLSA_methylation %>% #cultural background
  mutate(Race7 = case_when(
    SDC_CULT_WH_COM == 1 ~ "White",
    SDC_CULT_SA_COM == 1 ~ "South Asian",
    SDC_CULT_BL_COM == 1 ~ "Black",
    SDC_CULT_FP_COM == 1 | SDC_CULT_SE_COM == 1 | SDC_CULT_WA_COM == 1 |
      SDC_CULT_JA_COM == 1 | SDC_CULT_KO_COM == 1 | SDC_CULT_ZH_COM == 1 ~ "Asian",
    SDC_CULT_LA_COM == 1 ~ "Latino",
    SDC_CULT_AR_COM == 1 ~ "Arab",
    SDC_CULT_OT_COM == 1 ~ "Other",
  ))

label(CLSA_methylation$SEX_ASK_COM_factored)           <- "Sex (M/F)"#label categorical variables 
label(CLSA_methylation$INC_TOT_COM_factored)           <- "Total Personal Income (past 12 months)"
label(CLSA_methylation$INC_PTOT_COM_factored)          <- "Total Household Income (past 12 months)"
label(CLSA_methylation$ED_HIGH_COM_factored)           <- "Highest Education Obtained"
label(CLSA_methylation$WEA_SVNGSVL_MCQ_factored)       <- "Total Savings and Investments"
label(CLSA_methylation$WEA_DEBT_CCRD_MCQ_factored)     <- "Debt Type: Credit or Store Cards (Yes/No)"
label(CLSA_methylation$ENV_HMPRB_CON_MCQ_factored)     <- "Home Problems: Condensation"
label(CLSA_methylation$ENV_HMPRB_NOI_MCQ_factored)     <- "Home Problems: Noise"
label(CLSA_methylation$ICQ_SMOKE_COM_factored)         <- "Smoking Status"
label(CLSA_methylation$ALC_FREQ_COM_factored)          <- "Alcohol Consumption Frequency (past 12 months)"
label(CLSA_methylation$Race7)          <- "Cultural background"


table1(~ SEX_ASK_COM_factored + INC_TOT_COM_factored + INC_PTOT_COM_factored + ED_HIGH_COM_factored
       + WEA_SVNGSVL_MCQ_factored + WEA_DEBT_CCRD_MCQ_factored + ENV_HMPRB_CON_MCQ_factored + ENV_HMPRB_NOI_MCQ_factored
       + ICQ_SMOKE_COM_factored + ALC_FREQ_COM_factored + Race7
       , data = CLSA_methylation)


```


### Categorical Variables: Visualizations
Categorical variables were visualized using frequency tables and bar plots. 

```{r}

# Visualization_predictors_categorical ------------------------------------

barplot(table(CLSA_methylation$SEX_ASK_COM_factored),
        main = "Sex Distribution",
        col = "lightblue",
        ylab = "Count",
        las = 2,
        cex.names = 0.8)

barplot(table(CLSA_methylation$INC_TOT_COM_factored),
        main = "Total Personal Income",
        col = "lightblue",
        ylab = "Count",
        las = 2,
        cex.names = 0.6)


barplot(table(CLSA_methylation$INC_PTOT_COM_factored),
        main = "Total Household Income",
        col = "lightblue",
        ylab = "Count",
        las = 2,
        cex.names = 0.6)


barplot(table(CLSA_methylation$ED_HIGH_COM_factored),
        main = "Highest Education Obtained",
        col = "lightblue",
        ylab = "Count",
        las = 2,
        cex.names = 0.6)


barplot(table(CLSA_methylation$WEA_SVNGSVL_MCQ_factored),
        main = "Savings",
        col = "lightblue",
        ylab = "Count",
        las = 2,
        cex.names = 0.6)


barplot(table(CLSA_methylation$WEA_DEBT_CCRD_MCQ_factored),
        main = "Debt",
        col = "lightblue",
        ylab = "Count",
        las = 2,
        cex.names = 0.6)

barplot(table(CLSA_methylation$ENV_HMPRB_CON_MCQ_factored),
        main = "Condensation",
        col = "lightblue",
        ylab = "Count",
        las = 2,
        cex.names = 0.6)


barplot(table(CLSA_methylation$ENV_HMPRB_CON_MCQ_factored),
        main = "Noise",
        col = "lightblue",
        ylab = "Count",
        las = 2,
        cex.names = 0.6)


barplot(table(CLSA_methylation$ICQ_SMOKE_COM_factored),
        main = "Smoking",
        col = "lightblue",
        ylab = "Count",
        las = 2,
        cex.names = 0.6)

barplot(table(CLSA_methylation$ALC_FREQ_COM_factored),
        main = "Alcohol",
        col = "lightblue",
        ylab = "Count",
        las = 2,
        cex.names = 0.6)

barplot(table(CLSA_methylation$Race7),
        main = "Cultural background",
        col = "lightblue",
        ylab = "Count",
        las = 2,
        cex.names = 0.6)
```


### Continuous variables: Descriptive statistics 
Continuous variables were summarized using descriptive statistics. 



### Continuous variables: Visualizations 
Continuous variables were visualized with histograms, density plots, and boxplots.

```{r}

# Visualization_predictors_continuous -------------------------------------


#age
hist(CLSA_methylation$AGE_NMBR_COM,
     main = "Age in Years",
     xlab = "Age (years)",
     col = "skyblue",
     border = "white")


plot(density(CLSA_methylation$AGE_NMBR_COM, na.rm = TRUE),
     main = "Age in Years",
     xlab = "Age (years)",
     col = "darkblue",
     lwd = 2)

boxplot(CLSA_methylation$AGE_NMBR_COM,
        main = "Age in Years",
        ylab = "Age (years)",
        col = "lightgreen")



#PM2.5
hist(CLSA_methylation$pm25_1yr_com,
     main = "PM2.5 (1-Year Average)",
     xlab = "PM2.5 (µg/m³)",
     col = "skyblue",
     border = "white")

plot(density(CLSA_methylation$pm25_1yr_com, na.rm = TRUE),
     main = "PM2.5 (1-Year Average)",
     xlab = "PM2.5 (µg/m³)",
     col = "darkblue",
     lwd = 2)


boxplot(CLSA_methylation$pm25_1yr_com,
        main = "PM2.5 (1-Year Average)",
        ylab = "PM2.5 (µg/m³)",
        col = "lightgreen")

#NO₂ (1-year average)
hist(CLSA_methylation$no2_1yr_com,
     main = "NO2 (1-Year Average)",
     xlab = "NO2 (µg/m³)",
     col = "skyblue",
     border = "white")

plot(density(CLSA_methylation$no2_1yr_com, na.rm = TRUE),
     main = "NO2 (1-Year Average)",
     xlab = "NO2 (µg/m³)",
     col = "darkblue",
     lwd = 2)


boxplot(CLSA_methylation$no2_1yr_com,
        main = "NO2 (1-Year Average)",
        ylab = "NO₂ (µg/m³)",
        col = "lightgreen")



#ozone (1 y average)
hist(CLSA_methylation$o3_1yr_com,
     main = "Ozone (1-Year Average)",
     xlab = "Ozone (ppb)",
     col = "skyblue",
     border = "white")

plot(density(CLSA_methylation$o3_1yr_com, na.rm = TRUE),
     main = "Ozone (1-Year Average)",
     xlab = "Ozone (ppb)",
     col = "darkblue",
     lwd = 2)


boxplot(CLSA_methylation$o3_1yr_com,
        main = "Ozone (1-Year Average)",
        ylab = "Ozone (ppb)",
        col = "lightgreen")


#Greenness 
hist(CLSA_methylation$GRLANYY_26_COM,
     main = "Greenness",
     xlab = "Greenness (Max of annual mean NDVI within 250 m)",
     col = "skyblue",
     border = "white")

plot(density(CLSA_methylation$GRLANYY_26_COM, na.rm = TRUE),
     main = "Greenness",
     xlab = "Greenness",
     col = "darkblue",
     lwd = 2)


boxplot(CLSA_methylation$GRLANYY_26_COM,
        main = "Greenness",
        ylab = "Greenness",
        col = "lightgreen")

#temperature 
hist(CLSA_methylation$temp_1yr_com,
     main = "Temperature (1-Year Average)",
     xlab = "Temperature (°C)",
     col = "skyblue",
     border = "white")

plot(density(CLSA_methylation$temp_1yr_com, na.rm = TRUE),
     main = "Temperature (1-Year Average)",
     xlab = "Temperature (°C)",
     col = "darkblue",
     lwd = 2)


boxplot(CLSA_methylation$temp_1yr_com,
        main = "Temperature (1-Year Average)",
        ylab = "Temperature (°C)",
        col = "lightgreen")


#nighttime light brightness 
hist(CLSA_methylation$LGTNLTYY_01_COM,
     main = "Nighttime Light Brightness",
     xlab = "Nighttime Light Brightness",
     col = "skyblue",
     border = "white")

plot(density(CLSA_methylation$LGTNLTYY_01_COM, na.rm = TRUE),
     main = "Nighttime Light Brightness",
     xlab = "Nighttime Light Brightness",
     col = "darkblue",
     lwd = 2)


boxplot(CLSA_methylation$LGTNLTYY_01_COM,
        main = "Nighttime Light Brightness",
        ylab = "Nighttime Light Brightness",
        col = "lightgreen")


#nighttime brightness - shows that majority of participants have high level of nighttime brightness. keep this. 
library(ggplot2)
library(forcats)
library(dplyr)

CLSA_methylation %>%
  count(LGTNLTYY_01_COM) %>%
  arrange(desc(n)) %>%
  slice_head(n = 10) %>%                # keep top 10 categories
  ggplot(aes(x = reorder(LGTNLTYY_01_COM, n), y = n)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(x = "Category", y = "Count", 
       title = "Top 10 Categories of LGTNLTYY_01_COM")


CLSA_methylation %>%
  mutate(LGTNLTYY_01_COM = as.factor(LGTNLTYY_01_COM),
         LGTNLTYY_grouped = fct_lump_n(LGTNLTYY_01_COM, n = 10)) %>%
  count(LGTNLTYY_grouped) %>%
  ggplot(aes(x = reorder(LGTNLTYY_grouped, n), y = n)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(x = "Category", y = "Count", 
       title = "LGTNLTYY_01_COM (Top 10 + Other)")

#Age at Time of Immigration
hist(CLSA_methylation$SDC_DAIM_COM_clean,
     main = "Age at Time of Immigration",
     xlab = "Age at Time of Immigration (years)",
     col = "skyblue",
     border = "white")

plot(density(CLSA_methylation$SDC_DAIM_COM_clean, na.rm = TRUE),
     main = "Age at Time of Immigration (years)",
     xlab = "Age at Time of Immigration (years)",
     col = "darkblue",
     lwd = 2)


boxplot(CLSA_methylation$SDC_DAIM_COM_clean,
        main = "Age at Time of Immigration",
        ylab = "Age at Time of Immigration (years)",
        col = "lightgreen")

#Length of Time in Canada Since Immigration
hist(CLSA_methylation$SDC_DAIM_COM_clean,
     main = "Length of Time in Canada Since Immigration",
     xlab = "Length of Time in Canada Since Immigration (years)",
     col = "skyblue",
     border = "white")

plot(density(CLSA_methylation$SDC_DAIM_COM_clean, na.rm = TRUE),
     main = "Length of Time in Canada Since Immigration",
     xlab = "Length of Time in Canada Since Immigration (years)",
     col = "darkblue",
     lwd = 2)


boxplot(CLSA_methylation$SDC_DRES_COM_clean,
        main = "Length of Time in Canada Since Immigration",
        ylab = "Length of Time in Canada Since Immigration (years)",
        col = "lightgreen")

#Length of Time in Canada Since Immigration
hist(CLSA_methylation$SDC_YACA_YR_COM_clean,
     main = "Year of Arrival in Canada",
     xlab = "Year of Arrival in Canada",
     col = "skyblue",
     border = "white")

plot(density(CLSA_methylation$SDC_YACA_YR_COM_clean, na.rm = TRUE),
     main = "Year of Arrival in Canada",
     xlab = "Year of Arrival in Canada",
     col = "darkblue",
     lwd = 2)


boxplot(CLSA_methylation$SDC_YACA_YR_COM_clean,
        main = "Year of Arrival in Canada",
        ylab = "Year of Arrival in Canada",
        col = "lightgreen")

```




